---
title: "Decoy sugars hijack fungal carbon sensing and prevent plant disease"
author: "Guillaume Chesneau"
date: "26-05-2026"
output: html_document
---

# Before you start

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "") 
```

This script has been made on R 4.5.3 (and Rstudio 2026.01.2+418)
```{r R version}
R.version
#Load theme set
theme_set(theme_bw())
```

Pre-processing on QIIME2 and cutadapt :
- Demultiplexing
- DADA2 (16S parameters : trunc_F = 200, ITS : 200)
- Taxonomy assignation (silva database (2021) and UNITE (2022))

Start : You can directly start from point 2, by uploading the clean/filtered phyloseq objects

This script has been made on R version 4.5.3 (2026-03-11) and Rstudio 2026.01.2+418

```{r R version}
R.version
#Load theme set
library(ggplot2)
theme_set(theme_bw())
```

```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}
library(rstatix)  # For statistical analysis
library(ggpubr)   # For creating publication-ready plots using ggplot2
library(dunn.test)
library(ggthemes)
library(ggtext)
library(tidyr)
library(markdown)

# Prepare base theme
base_theme <- theme_bw() +
  theme(
    strip.background = element_rect(color = "white", fill = "white"),
    strip.text.x = element_markdown(size = 10, color = "black", hjust = 0.5, vjust = 0.5),
    legend.title = element_text(size = 11),
    axis.title.y = element_text(size =11),
    legend.text = element_text(size = 10),
    axis.text.x = element_text(size = 10, color = "black", angle = 90, hjust = 1),
    axis.text.y = element_text(size = 10, color = "black"),
    panel.border = element_rect(color = "black", fill = NA, size = 0.75))


#Vector for ordering 
Treatment_order <- c("Water", "Glucose", "3OMG")
Microbes_order <- c("MF", "FSC")

#Vector for color
treatment_colors <- c("Water"   = "#949494", "Glucose" = "#817E9F", "3OMG"    = "#7FC29B")
``` 


#A] Pre-processing.    
Don't need to run this part if you're just interested by making the plots.    
Directly go to B] Start of the script.

## 1.1 - Construct phyloseq objects

* 16S

```{r, echo=FALSE, results='hide', eval = FALSE}
library(phyloseq)
library(ggplot2)
library(dplyr)
library(tidyr)
library(tibble)
library(Biostrings)

# Load data
asv_table_16S <- read.table(file = "Final_16S/asv_table_16S.tsv", sep = "\t", dec = ".", header = TRUE)
tax_table_16S <- read.table(file = "Final_16S/taxonomy_16S.tsv", sep = "\t", dec = ".", header = TRUE)
metadata <- read.table(file = "Final_16S/metadata_16S.txt", sep = "\t", dec = ".", header = TRUE)
refseq <- readDNAStringSet("Final_16S/dna-sequences_16S.fasta")

# Clean taxonomy table
tax_table_16S <- tax_table_16S %>%
  mutate(Taxon = gsub("([dpcofgs]__)", "", Taxon)) %>%
  separate(Taxon, into = c("Kingdom","Phylum","Class","Order","Family","Genus","Species"), sep = ";") %>%
  mutate(Kingdom = "All")

# Rename ASVs consistently
asv_ids <- paste0("ASV", seq_len(nrow(asv_table_16S)))
rownames(asv_table_16S) <- asv_ids
rownames(tax_table_16S) <- asv_ids
names(refseq) <- asv_ids

# Remove old ID columns if present
asv_table_16S <- asv_table_16S[,-1]
tax_table_16S <- tax_table_16S[,-1]

# Prepare phyloseq components
OTU <- otu_table(as.matrix(asv_table_16S), taxa_are_rows = TRUE)
TAX <- tax_table(as.matrix(tax_table_16S))
rownames(metadata) <- metadata[,1]
metadata <- metadata[,-1]
SAM <- sample_data(metadata)
REFSEQ <- refseq

# Build phyloseq object
physeq_16S <- phyloseq(OTU, TAX, SAM, REFSEQ)
physeq_16S
```

* ITS

```{r, echo=FALSE, results='hide', eval = FALSE}
library(phyloseq)
library(ggplot2)
library(dplyr)
library(tidyr)
library(tibble)
library(Biostrings)

# Load data
asv_table_ITS <- read.table(file = "Final_ITS/asv_table-ITS.tsv", sep = "\t", dec = ".", header = TRUE)
tax_table_ITS <- read.table(file = "Final_ITS/taxonomy-ITS.tsv", sep = "\t", dec = ".", header = TRUE)
metadata <- read.table(file = "Final_ITS/metadata_ITS.txt", sep = "\t", dec = ".", header = TRUE)
refseq <- readDNAStringSet("Final_ITS/dna-sequences-ITS.fasta")

# Clean taxonomy table
tax_table_ITS <- tax_table_ITS %>%
  mutate(Taxon = gsub("([dpcofgs]__)", "", Taxon)) %>%
  separate(Taxon, into = c("Kingdom","Phylum","Class","Order","Family","Genus","Species"), sep = ";") %>%
  mutate(Kingdom = "All")

# Rename ASVs consistently
asv_ids <- paste0("FASV", seq_len(nrow(asv_table_ITS)))
rownames(asv_table_ITS) <- asv_ids
rownames(tax_table_ITS) <- asv_ids
names(refseq) <- asv_ids

# Remove old ID columns if present
asv_table_ITS <- asv_table_ITS[,-1]
tax_table_ITS <- tax_table_ITS[,-1]

# Prepare phyloseq components
OTU <- otu_table(as.matrix(asv_table_ITS), taxa_are_rows = TRUE)
TAX <- tax_table(as.matrix(tax_table_ITS))
rownames(metadata) <- metadata[,1]
metadata <- metadata[,-1]
SAM <- sample_data(metadata)
REFSEQ <- refseq

# Build phyloseq object
physeq_ITS <- phyloseq(OTU, TAX, SAM, REFSEQ)

physeq_ITS
```
## 1.2 - Filtering  phyloseq objects

### 1.2-1 16SrRNA
* Phyloseq object composition before filtering

phyloseq-class experiment-level object
otu_table()   OTU Table:         [ 8380 taxa and 25 samples ]
sample_data() Sample Data:       [ 25 samples by 5 sample variables ]
tax_table()   Taxonomy Table:    [ 8380 taxa by 8 taxonomic ranks ]
refseq()      DNAStringSet:      [ 8380 reference sequences ]


```{r, echo=FALSE, results='hide', eval = FALSE}
# Calculate the number of taxa before filtering
num_taxa_before <- ntaxa(physeq_16S)

#Filtered ASV based on taxonomy
physeq_16S_1a <- subset_taxa(physeq_16S,!is.na(tax_table(physeq_16S)[, "Phylum"]) & tax_table(physeq_16S)[, "Phylum"] != "Unassigned") #Before 8373, After  7846, don't use to keep everything
physeq_16S_1b <- subset_taxa(physeq_16S, !Order %in% c("Chloroplast"))  #Before 8379, After  8380

num_taxa_after1 <- data.frame(Before_filtering = ntaxa(physeq_16S), Remove_Unassigned = ntaxa(physeq_16S_1a), remove_chloroplast = ntaxa(physeq_16S_1b))

#Before after filtering
# Calculate the total number of reads per sample
total_reads <- colSums(otu_table(physeq_16S))
total_reads_Assigned <- colSums(otu_table(physeq_16S_1a))
total_reads_After_Chloro <- colSums(otu_table(physeq_16S_1b))

#Gather
total_filtering_reads <- cbind(row.names(total_reads), total_reads, total_reads_Assigned, total_reads_After_Chloro)
total_filtering_reads_f <- merge(total_filtering_reads, design, by = "row.names")
write.csv(total_filtering_reads, file = "total_filtering_reads.csv", row.names = FALSE)

# Create a data frame with sample IDs and total reads
reads_per_sample_table <- data.frame(SampleID = sample_names(physeq_16S_1b), TotalReads = total_reads,TotalReads_Assigned = total_reads_Assigned, TotalReads_After_Chloro =  total_reads_After_Chloro, Percent_after_filtering = 100*total_reads_After_Chloro/total_reads)

#Subset Samples (remove sample with less than 1000 counts)
physeq_16S_2 <- prune_samples(sample_sums(physeq_16S_1b)>=1000, physeq_16S_1b) #Before 25, After  24, remove the control water
physeq_16S_3 <- filter_taxa(physeq_16S_2, function(x) sum(x) > 0, TRUE) #Before 8379 After 8373

#Remove extremely rare ASVs
#Occurence, remove ASV present in a single sample
physeq_16S_4 <- filter_taxa(physeq_16S_3, function(x) sum(x > 0) > 1, TRUE) #Before 8373 After 2210

#Abundance filter (not applied)
#physeq_16S_5 <- transform_sample_counts(physeq_16S_4, function(x) ifelse((x / sum(x)) >= 0.001, x, 0))
#physeq_16S_5 <- filter_taxa(physeq_16S_5, function(x) sum(x) > 0, TRUE)

#Save RDS
saveRDS(physeq_16S_4, "physeq_16S_clean.rds") #2186 taxa and 24 samples
```



### 1.2-1-2 ITs data

* Phyloseq object composition before filtering

phyloseq-class experiment-level object
otu_table()   OTU Table:         [ 162 taxa and 768 samples ]
sample_data() Sample Data:       [ 768 samples by 7 sample variables ]
tax_table()   Taxonomy Table:    [ 162 taxa by 8 taxonomic ranks ]
refseq()      DNAStringSet:      [ 162 reference sequences ]


```{r, echo=FALSE, results='hide', eval = FALSE}
# Calculate the number of taxa before filtering
num_taxa_before <- ntaxa(physeq_ITS)

#Filtered ASV based on taxonomy
physeq_ITS_1a <- subset_taxa(physeq_ITS,!is.na(tax_table(physeq_ITS)[, "Phylum"]) & tax_table(physeq_ITS)[, "Phylum"] != "Unassigned") #Before 1994, After  1120, don't use to keep everything

#Subset Samples (remove sample with less than 1000 counts)
physeq_ITS_2 <- prune_samples(sample_sums(physeq_ITS)>=1000, physeq_ITS_1b) #Before 25, After  25
physeq_ITS_3 <- filter_taxa(physeq_ITS_2, function(x) sum(x) > 0, TRUE) #Before 1994 After 1994

#Remove extremely rare ASVs
#Occurence, remove ASV present in a single sample
physeq_ITS_4 <- filter_taxa(physeq_ITS_3, function(x) sum(x > 0) > 1, TRUE) #Before 1994 After 420

#Abundance filter (not applied)
#physeq_ITS_5 <- transform_sample_counts(physeq_ITS_4, function(x) ifelse((x / sum(x)) >= 0.001, x, 0))
#physeq_ITS_5 <- filter_taxa(physeq_ITS_5, function(x) sum(x) > 0, TRUE)

#Remove Control S25
physeq_ITS_5 <- subset_samples(physeq_ITS_4, Treatment != "Control") #Does not remove any taxa, Control is almost empty.

#Save RDS
saveRDS(physeq_ITS_5, "physeq_ITS_clean.rds") #2186 taxa and 24 samples
```

# 2 - Start community analysis   

**After filtering**   
16S dataset : 12469 taxa and 420 samples.   

```{r echo=FALSE, message=FALSE}
#Upload cleaned datasets
library(phyloseq); packageVersion("phyloseq")
d16S <- readRDS("physeq_16S_clean.rds") 
dITS <- readRDS("physeq_ITS_clean.rds") 
```

## 2.1 - Sequencing depth

We first assess the number of sequences per sample
```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}
#Assess sequencing depth
library(data.table)
sdt16S <- data.table(as(sample_data(d16S), "data.frame"),
                     TotalReads = sample_sums(d16S), keep.rownames = TRUE)

sdtITS <- data.table(as(sample_data(dITS), "data.frame"),
                     TotalReads = sample_sums(dITS), keep.rownames = TRUE)
```
 
## 2.2 - Rarefaction    
*(intercept x = 30000 reads, 999 steps)*.  
Just loose one MF glucose sample which as really low number of reads = 4539
```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}
library("vegan")
library(ggplot2)
source("~/Documents/1-CR_SAVE/5-Script/Amplicon_sequencing/Rscripts/richness.R")

#Order
sample_data(d16S)$Treatment <- factor(sample_data(d16S)$Treatment, levels = Treatment_order)
sample_data(d16S)$Microbes <- factor(sample_data(d16S)$Microbes, levels = Microbes_order)

#16S
rcurve16S <- ggrare(d16S,color = "Treatment",
                    step = 999, 
                    se = FALSE)  + labs (title = "16S rRNA gene") + geom_vline(xintercept=1000)

rcurve16Sf <- rcurve16S + theme(strip.text.x = element_text(size = 12, color = "white"),
        axis.title.x = element_text( size = 12, face = "bold"),
        axis.title.y = element_text( size = 12, face = "bold"),
        plot.title = element_text(size = 16),
        axis.text.x = element_text(face = "bold"), 
        axis.text.y = element_text(face = "bold"))+  
    ylab ("Species richness") + xlab ("Sample size") + ggtitle("16S rRNA gene") + facet_grid(~Microbes+Treatment, scale = "free") +
  scale_color_manual(values = treatment_colors) +
  base_theme
rcurve16Sf

#Order
sample_data(dITS)$Treatment <- factor(sample_data(dITS)$Treatment, levels = Treatment_order)
sample_data(dITS)$Microbes <- factor(sample_data(dITS)$Microbes, levels = Microbes_order)

#ITS
rcurveITS <- ggrare(dITS,color = "Treatment",
                    step = 999, 
                    se = FALSE)  + labs (title = "ITS rRNA gene") + geom_vline(xintercept=1000)

rcurveITSf <- rcurveITS + theme(strip.text.x = element_text(size = 12, color = "white"),
        axis.title.x = element_text( size = 12, face = "bold"),
        axis.title.y = element_text( size = 12, face = "bold"),
        plot.title = element_text(size = 16),
        axis.text.x = element_text(face = "bold"), 
        axis.text.y = element_text(face = "bold"))+  
    ylab ("Species richness") + xlab ("Sample size") + ggtitle("ITS rRNA gene") + facet_grid(~Microbes+Treatment, scale = "free") +
  scale_color_manual(values = treatment_colors) +
  base_theme
rcurveITSf
```

```{r, echo=FALSE, message=FALSE}
rcurve16Sf
rcurveITSf
```

## 2.3 - Alpha diversity analysis

**Normalization based on coverage:**         

 
* This will rarefy you so that all your samples have the same coverage (and it tells you the minimum coverage of your dataset)[^1].           

[^1]: Chao et al 2012, Chao et al 2014, Roswell et al 2021, doi: 10.1111/oik.07202. 

```{r, echo=FALSE, warning=FALSE, message=FALSE, results="hide"}
#Rarefy datasets
r16S <- rarefy_even_depth(d16S, sample.size = 31000, rngseed = 900) 
rITS <- rarefy_even_depth(dITS, sample.size = 46000, rngseed = 900) 
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results="hide"}
#16S
#Table of alpha-diversity estimators
table_r16S <- estimate_richness(r16S, split = TRUE, measures=c("Observed", "InvSimpson","Shannon"))

#Bind design + table of alpha_div
sdr16S <- sample_data(r16S)
datar16S <- cbind(sample_data(sdr16S), table_r16S) 

#ITS
#Table of alpha-diversity estimators
table_rITS <- estimate_richness(rITS, split = TRUE, measures=c("Observed", "InvSimpson","Shannon"))

#Bind design + table of alpha_div
sdrITS <- sample_data(rITS)
datarITS <- cbind(sample_data(sdrITS), table_rITS) 
```

**Observed richness and Shannon diversity index**. 

**16S**.   
```{r, echo=FALSE, message=FALSE, results="hide"}
#16S
library(ggplot2)
library(rstatix)
library(ggpubr)

#Calculate stats 
#Observed
stat_obs <- datar16S %>%
  group_by(Microbes) %>%
  wilcox_test(Observed ~ Treatment) %>%
  #filter(group1 == "Water") %>%
  adjust_pvalue(method = "BH") %>%
  add_significance()

stat_obs <- stat_obs %>%
  add_xy_position(x = "Treatment")

#Calculate stats 
#Shannon
stat_shannon <- datar16S %>%
  group_by(Microbes) %>%
  wilcox_test(Shannon ~ Treatment) %>%
 # filter(group1 == "Water") %>%
  adjust_pvalue(method = "BH") %>%
  add_significance()

stat_shannon <- stat_shannon %>%
  add_xy_position(x = "Treatment")

#Plot
Plot_16S_Observed <- ggplot(data = datar16S,
                            aes(x = Treatment, y = Observed, fill = Treatment)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(alpha = 1) +
  facet_wrap(~Microbes, scales = "free_y") +
  scale_fill_manual(values = treatment_colors) +
  ggpubr::stat_pvalue_manual(
    stat_obs,
    label = "p.adj.signif",
    tip.length = 0.01,
    hide.ns = TRUE
  ) +
  ylab("Observed ASV richness") +
  xlab("genotype") +
  ggtitle("16S rRNA") +
  base_theme
Plot_16S_Observed

Plot_16S_Shannon <- ggplot(data = datar16S,
                            aes(x = Treatment, y = Shannon, fill = Treatment)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(alpha = 1) +
  facet_wrap(~Microbes, scales = "free_y") +
  scale_fill_manual(values = treatment_colors) +
  ggpubr::stat_pvalue_manual(
    stat_shannon,
    label = "p.adj.signif",
    tip.length = 0.01,
    hide.ns = TRUE
  ) +
  ylab("Shannon Diversity") +
  xlab("genotype") +
  ggtitle("16S rRNA") +
  base_theme
Plot_16S_Shannon
```


**ITS**.     
```{r, echo=FALSE, message=FALSE, results="hide"}
#ITS
library(ggplot2)
library(rstatix)
library(ggpubr)

#Calculate stats 
#Observed
stat_obs <- datarITS %>%
  group_by(Microbes) %>%
  wilcox_test(Observed ~ Treatment) %>%
  #filter(group1 == "Water") %>%
  adjust_pvalue(method = "BH") %>%
  add_significance()

stat_obs <- stat_obs %>%
  add_xy_position(x = "Treatment")

#Calculate stats 
#Shannon
stat_shannon <- datarITS %>%
  group_by(Microbes) %>%
  wilcox_test(Shannon ~ Treatment) %>%
 # filter(group1 == "Water") %>%
  adjust_pvalue(method = "BH") %>%
  add_significance()

stat_shannon <- stat_shannon %>%
  add_xy_position(x = "Treatment")

#Plot
Plot_ITS_Observed <- ggplot(data = datarITS,
                            aes(x = Treatment, y = Observed, fill = Treatment)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(alpha = 1) +
  facet_wrap(~Microbes, scales = "free_y") +
  scale_fill_manual(values = treatment_colors) +
  ggpubr::stat_pvalue_manual(
    stat_obs,
    label = "p.adj.signif",
    tip.length = 0.01,
    hide.ns = TRUE
  ) +
  ylab("Observed ASV richness") +
  xlab("genotype") +
  ggtitle("ITS") +
  base_theme
Plot_ITS_Observed

Plot_ITS_Shannon <- ggplot(data = datarITS,
                            aes(x = Treatment, y = Shannon, fill = Treatment)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(alpha = 1) +
  facet_wrap(~Microbes, scales = "free_y") +
  scale_fill_manual(values = treatment_colors) +
  ggpubr::stat_pvalue_manual(
    stat_shannon,
    label = "p.adj.signif",
    tip.length = 0.01,
    hide.ns = TRUE
  ) +
  ylab("Shannon Diversity") +
  xlab("genotype") +
  ggtitle("ITS") +
  base_theme
Plot_ITS_Shannon
```

## 2.4  Permanova Habitats

**16S**.  
```{r, echo=FALSE, message=FALSE}
library(phyloseq)
library(vegan)
library(ggplot2)
library(dplyr)
library(pairwiseAdonis)
library(multcompView)
library(tidyr)
library(cowplot)


# Variability function
variability_table <- function(cca) {
  chi <- c(cca$tot.chi, cca$CCA$tot.chi, cca$CA$tot.chi)
  data.frame(
    inertia = chi,
    proportion = chi / chi[1],
    row.names = c("total", "constrained", "unconstrained")
  )
}

# Letters from pairwise PERMANOVA
get_letters <- function(bray, mdata){
  pw <- pairwise.adonis2(bray ~ Treatment, data=mdata, perm=999)
  pw_table <- do.call(rbind, lapply(names(pw)[-1], function(x){
    df <- pw[[x]]
    data.frame(contrast=x, p.value=df$`Pr(>F)`[1])
  }))
  pw_table <- pw_table %>% separate(contrast, into=c("group1","group2"), sep="_vs_")
  all_groups <- unique(c(pw_table$group1, pw_table$group2))
  pw_matrix <- matrix(1, nrow=length(all_groups), ncol=length(all_groups),
                      dimnames=list(all_groups, all_groups))
  for(i in 1:nrow(pw_table)){
    g1 <- pw_table$group1[i]; g2 <- pw_table$group2[i]
    pw_matrix[g1,g2] <- pw_table$p.value[i]
    pw_matrix[g2,g1] <- pw_table$p.value[i]
  }
  letters_df <- data.frame(Treatment=names(multcompLetters(pw_matrix<0.05)$Letters),
                           Letters=multcompLetters(pw_matrix<0.05)$Letters)
  return(letters_df)
}

# Centroids
get_centroids <- function(df, letters_df){
  df %>% group_by(Treatment, Microbes) %>%
    summarize(wCAP1=mean(wCAP1), wCAP2=mean(wCAP2)) %>%
    left_join(letters_df, by="Treatment")
}

# Loop over Microbes
plot_list <- list()
for(mb in c("MF","FSC")){
  df_CAP <- subset_samples(r16S, Microbes==mb)
  bray <- phyloseq::distance(df_CAP, method="bray")
  mdata <- as(sample_data(df_CAP), "data.frame")
  Cap <- capscale(bray ~ Treatment, data=mdata)
  var_tbl <- variability_table(Cap)
  eigenvalues <- Cap$CCA$eig / sum(Cap$CCA$eig)
  adonis_res <- adonis2(bray ~ Treatment, data=mdata, perm=999)

  # Ordination points
  points <- data.frame(
    uCAP1 = Cap$CCA$u[,1], uCAP2 = Cap$CCA$u[,2],
    wCAP1 = Cap$CCA$wa[,1], wCAP2 = Cap$CCA$wa[,2],
    CA1 = Cap$CA$u[,1], CA2 = Cap$CA$u[,2]
  )
  df_points <- cbind(points, mdata[match(rownames(points), rownames(mdata)),])

  # Letters & centroids
  letters_df <- get_letters(bray, mdata)
  centroids <- get_centroids(df_points, letters_df)
  centroids$Microbes <- mb

  # Plot
  p <- ggplot(df_points, aes(x=wCAP1, y=wCAP2, color=Treatment, fill=Treatment, shape=Microbes)) +
    geom_segment(data=df_points %>% left_join(centroids %>% select(Treatment,wCAP1,wCAP2), by="Treatment"),
                 aes(x=wCAP1.x, y=wCAP2.x, xend=wCAP1.y, yend=wCAP2.y),
                 color="gray70", alpha=0.5) +
    geom_point(alpha=0.7, size=1.5) +
    geom_point(data=centroids, aes(x=wCAP1, y=wCAP2, fill=Treatment),
               shape=21, color="black", size=4, stroke=1.2) +
    geom_text(data=centroids, aes(x=wCAP1, y=wCAP2, label=Letters, color=Treatment),
              fontface="bold", size=5, vjust=-1) +
    labs(x=sprintf("wCPCoA 1 (%.2f%%)",100*eigenvalues[1]),
         y=sprintf("wCPCoA 2 (%.2f%%)",100*eigenvalues[2]),
         title=sprintf("%.2f%% of variance; p=%.3f",100*var_tbl["constrained","proportion"],
                       adonis_res$`Pr(>F)`[1])) +
      scale_color_manual(values = treatment_colors) +
      scale_fill_manual(values = treatment_colors) +
    facet_grid(~Microbes, labeller = as_labeller(labels)) +
    base_theme
  plot_list[[mb]] <- p
}

# Combine plots
Final_plot_16S_Root <- plot_grid(plot_list$MF, plot_list$FSC, ncol=1, nrow=2, rel_heights=c(2,2))
Final_plot_16S_Root
```

**ITS**.  
```{r, echo=FALSE, message=FALSE}
library(phyloseq)
library(vegan)
library(ggplot2)
library(dplyr)
library(pairwiseAdonis)
library(multcompView)
library(tidyr)
library(cowplot)


# Variability function
variability_table <- function(cca) {
  chi <- c(cca$tot.chi, cca$CCA$tot.chi, cca$CA$tot.chi)
  data.frame(
    inertia = chi,
    proportion = chi / chi[1],
    row.names = c("total", "constrained", "unconstrained")
  )
}

# Letters from pairwise PERMANOVA
get_letters <- function(bray, mdata){
  pw <- pairwise.adonis2(bray ~ Treatment, data=mdata, perm=999)
  pw_table <- do.call(rbind, lapply(names(pw)[-1], function(x){
    df <- pw[[x]]
    data.frame(contrast=x, p.value=df$`Pr(>F)`[1])
  }))
  pw_table <- pw_table %>% separate(contrast, into=c("group1","group2"), sep="_vs_")
  all_groups <- unique(c(pw_table$group1, pw_table$group2))
  pw_matrix <- matrix(1, nrow=length(all_groups), ncol=length(all_groups),
                      dimnames=list(all_groups, all_groups))
  for(i in 1:nrow(pw_table)){
    g1 <- pw_table$group1[i]; g2 <- pw_table$group2[i]
    pw_matrix[g1,g2] <- pw_table$p.value[i]
    pw_matrix[g2,g1] <- pw_table$p.value[i]
  }
  letters_df <- data.frame(Treatment=names(multcompLetters(pw_matrix<0.05)$Letters),
                           Letters=multcompLetters(pw_matrix<0.05)$Letters)
  return(letters_df)
}

# Centroids
get_centroids <- function(df, letters_df){
  df %>% group_by(Treatment, Microbes) %>%
    summarize(wCAP1=mean(wCAP1), wCAP2=mean(wCAP2)) %>%
    left_join(letters_df, by="Treatment")
}

# Loop over Microbes
plot_list <- list()
for(mb in c("MF","FSC")){
  df_CAP <- subset_samples(rITS, Microbes==mb)
  bray <- phyloseq::distance(df_CAP, method="bray")
  mdata <- as(sample_data(df_CAP), "data.frame")
  Cap <- capscale(bray ~ Treatment, data=mdata)
  var_tbl <- variability_table(Cap)
  eigenvalues <- Cap$CCA$eig / sum(Cap$CCA$eig)
  adonis_res <- adonis2(bray ~ Treatment, data=mdata, perm=999)

  # Ordination points
  points <- data.frame(
    uCAP1 = Cap$CCA$u[,1], uCAP2 = Cap$CCA$u[,2],
    wCAP1 = Cap$CCA$wa[,1], wCAP2 = Cap$CCA$wa[,2],
    CA1 = Cap$CA$u[,1], CA2 = Cap$CA$u[,2]
  )
  df_points <- cbind(points, mdata[match(rownames(points), rownames(mdata)),])

  # Letters & centroids
  letters_df <- get_letters(bray, mdata)
  centroids <- get_centroids(df_points, letters_df)
  centroids$Microbes <- mb

  # Plot
  p <- ggplot(df_points, aes(x=wCAP1, y=wCAP2, color=Treatment, fill=Treatment, shape=Microbes)) +
    geom_segment(data=df_points %>% left_join(centroids %>% select(Treatment,wCAP1,wCAP2), by="Treatment"),
                 aes(x=wCAP1.x, y=wCAP2.x, xend=wCAP1.y, yend=wCAP2.y),
                 color="gray70", alpha=0.5) +
    geom_point(alpha=0.7, size=1.5) +
    geom_point(data=centroids, aes(x=wCAP1, y=wCAP2, fill=Treatment),
               shape=21, color="black", size=4, stroke=1.2) +
    geom_text(data=centroids, aes(x=wCAP1, y=wCAP2, label=Letters, color=Treatment),
              fontface="bold", size=5, vjust=-1) +
    labs(x=sprintf("wCPCoA 1 (%.2f%%)",100*eigenvalues[1]),
         y=sprintf("wCPCoA 2 (%.2f%%)",100*eigenvalues[2]),
         title=sprintf("%.2f%% of variance; p=%.3f",100*var_tbl["constrained","proportion"],
                       adonis_res$`Pr(>F)`[1])) +
      scale_color_manual(values = treatment_colors) +
      scale_fill_manual(values = treatment_colors) +
    facet_grid(~Microbes, labeller = as_labeller(labels)) +
    base_theme
  plot_list[[mb]] <- p
}

# Combine plots
Final_plot_ITS_Root <- plot_grid(plot_list$MF, plot_list$FSC, ncol=1, nrow=2, rel_heights=c(2,2))
Final_plot_ITS_Root
```

## 2.5 - Taxonomic composition

### 2.5.1 - Prevalence/Abundance
Evaluation of the taxonomic composition    

```{r, echo=FALSE, message=FALSE}

#16S
library(ggplot2)
#Compute prevalence of each feature, store as data.frame
prevdf16S <- apply(X = otu_table(d16S),
                   MARGIN = ifelse(taxa_are_rows(d16S), yes = 1, no = 2),
                   FUN = function(x){sum(x > 0)})

#Add taxonomy and total read counts to this data.frame
prevdf16S <- data.frame(Prevalence = prevdf16S, # Issue here with species name NA
                        TotalAbundance = taxa_sums(d16S),
                        tax_table(d16S))

#Taxa prevalence vs total count
plot16S <- ggplot(prevdf16S, aes(TotalAbundance, Prevalence / nsamples(d16S),color=Phylum)) +
  geom_hline(yintercept = 0.05, alpha = 0.5, linetype = 2) + geom_point(size = 2, alpha = 0.7) +
  scale_x_log10() +  xlab("Total Abundance") + ylab("Prevalence [Frac. Samples]") + ggtitle ("16S rRNA") +
  facet_wrap(~Phylum) + theme(strip.text.x = element_text(size = 10, color = "white"),
        axis.title.x = element_text(size = 11, face = "bold"),
        axis.title.y = element_text(size = 11, face = "bold"),
        plot.title = element_text(size = 12, face = "bold"),
        axis.text.x = element_text(angle = 70, hjust=1), 
        axis.text.y = element_text()) +
  base_theme
plot16S


#ITS
library(ggplot2)
#Compute prevalence of each feature, store as data.frame
prevdfITS <- apply(X = otu_table(dITS),
                   MARGIN = ifelse(taxa_are_rows(dITS), yes = 1, no = 2),
                   FUN = function(x){sum(x > 0)})

#Add taxonomy and total read counts to this data.frame
prevdfITS <- data.frame(Prevalence = prevdfITS, # Issue here with species name NA
                        TotalAbundance = taxa_sums(dITS),
                        tax_table(dITS))

#Taxa prevalence vs total count
plotITS <- ggplot(prevdfITS, aes(TotalAbundance, Prevalence / nsamples(dITS),color=Phylum)) +
  geom_hline(yintercept = 0.05, alpha = 0.5, linetype = 2) + geom_point(size = 2, alpha = 0.7) +
  scale_x_log10() +  xlab("Total Abundance") + ylab("Prevalence [Frac. Samples]") + ggtitle ("ITS rRNA") +
  facet_wrap(~Phylum) + theme(strip.text.x = element_text(size = 10, color = "white"),
        axis.title.x = element_text(size = 11, face = "bold"),
        axis.title.y = element_text(size = 11, face = "bold"),
        plot.title = element_text(size = 12, face = "bold"),
        axis.text.x = element_text(angle = 70, hjust=1), 
        axis.text.y = element_text()) +
  base_theme
plotITS

```

### 2.5.2 - Relative abundance per sample.   
Evaluation of the taxonomic composition     
```{r, echo=FALSE, message=FALSE}
library(reshape2); packageVersion("reshape2")
source("~/Documents/1-CR_SAVE/5-Script/Amplicon_sequencing/Rscripts/graphical_methods.R")

#Color
chill_palette_12 <- c(
  "#4C6A92", # muted blue
  "#6C9A8B", # desaturated teal
  "#A3B18A", # sage green
  "#DDBEA9", # sand
  "#D9A066", # warm brown
  "#8E7DBE", # soft purple
  "#E07A5F", # muted coral
  "#A8D8B9", # soft green
  "#6B6F8F", # deep indigo
  "#FAEED6", # pale amber
  "#A8DADC", # light aqua
  "#C9ADA7"  # dusty rose
)

#16S
library(reshape2)
#Plot
tax16S <- plot_composition(r16S, "Kingdom", "All", "Family", numberOfTaxa=10, fill="Family") + 
  facet_grid(~ Microbes + Treatment, scales="free_x") +
theme(strip.text.x = element_text(size = 12, color = "white"),
        axis.title.y = element_text( size = 12, face = "bold"),
        plot.title = element_text(size = 16),
        axis.text.y = element_text(face = "bold"),
        axis.text.x = element_blank())+  
  ylab ("Relative abundance (%)") + 
  ggtitle("Taxonomy - 16S") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
    scale_fill_manual(values = chill_palette_12) +
    scale_color_manual(values = chill_palette_12) +
  base_theme

tax16S

taxITS <- plot_composition(rITS, "Kingdom", "All", "Family", numberOfTaxa=10, fill="Family") + 
  facet_grid(~ Microbes + Treatment, scales="free_x") +
theme(strip.text.x = element_text(size = 12, color = "white"),
        axis.title.y = element_text( size = 12, face = "bold"),
        plot.title = element_text(size = 16),
        axis.text.y = element_text(face = "bold"),
        axis.text.x = element_blank())+  
  ylab ("Relative abundance (%)") + 
  ggtitle("Taxonomy - ITS") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
      scale_fill_manual(values = chill_palette_12) +
    scale_color_manual(values = chill_palette_12) +
  base_theme

taxITS
```
#Heat map + stats based on abundance

```{r, echo=FALSE, message=FALSE}
library(phyloseq)
library(tidyverse)
library(pheatmap)
library(FSA)
library(rcompanion)

# 1. EXTRACT OTU TABLE
otu <- as.data.frame(t(otu_table(rITS)))

# 2. ADD METADATA
meta <- as.data.frame(sample_data(rITS))
meta$SampleID <- rownames(meta)

otu$SampleID <- rownames(otu)

df_long <- otu %>%
  pivot_longer(-SampleID, names_to = "ASV", values_to = "Abundance") %>%
  left_join(meta, by = "SampleID")

# 3. ADD TAXONOMY (IMPORTANT)
Taxonomy_ITS <- as.data.frame(tax_table(rITS))
Taxonomy_ITS$ASV <- rownames(Taxonomy_ITS)

df_long <- df_long %>%
  left_join(Taxonomy_ITS, by = "ASV")

# CREATE LABEL
df_long$ASV_Species <- paste0(df_long$ASV, "_", df_long$Species)

# 4. KEEP ONLY FSC
df_long_FSC <- subset(df_long, Microbes %in% c("FSC"))

# 5. KRUSKAL-WALLIS FILTER
stats_asv <- df_long_FSC %>%
  group_by(ASV_Species) %>%
  summarise(
    pvalue = kruskal.test(Abundance ~ Treatment)$p.value,
    .groups = "drop"
  )

sig_ASVs <- stats_asv %>%
  filter(!is.na(pvalue), pvalue < 0.05) %>%
  pull(ASV_Species)

df_sig <- df_long_FSC %>%
  filter(ASV_Species %in% sig_ASVs)

# 6. MATRIX (MEAN ABUNDANCE)
mat <- df_sig %>%
  group_by(Treatment, ASV_Species) %>%
  summarise(Abundance = mean(Abundance), .groups = "drop") %>%
  pivot_wider(names_from = ASV_Species, values_from = Abundance) %>%
  column_to_rownames("Treatment") %>%
  as.matrix()

mat[is.na(mat)] <- 0
mat <- log1p(mat)

# 7. DUNN TEST + LETTERS
letters_df <- df_sig %>%
  group_by(ASV_Species) %>%
  summarise(
    model = list(
      tryCatch(
        dunnTest(Abundance ~ Treatment, method = "bh")$res,
        error = function(e) NULL
      )
    ),
    .groups = "drop"
  ) %>%
  mutate(
    letters = map(model, ~ {
      if (is.null(.x) || nrow(.x) == 0) return(NULL)

      rcompanion::cldList(
        P.adj ~ Comparison,
        data = .x,
        threshold = 0.05
      )
    })
  )

# 8. BUILD LETTER MATRIX
treatments <- unique(df_sig$Treatment)
asvs <- unique(df_sig$ASV_Species)

letter_mat <- matrix("",
                     nrow = length(treatments),
                     ncol = length(asvs))

rownames(letter_mat) <- treatments
colnames(letter_mat) <- asvs

for (i in seq_len(nrow(letters_df))) {

  asv <- letters_df$ASV_Species[i]
  letters <- letters_df$letters[[i]]

  if (!is.null(letters) && nrow(letters) > 0) {

    for (j in seq_len(nrow(letters))) {

      tr <- letters$Group[j]
      letter_mat[tr, asv] <- letters$Letter[j]

    }
  }
}




# 9. HEATMAP (STORE RESULT)
HM <- pheatmap(
  mat,
  scale = "column",
  color = colorRampPalette(c("darkblue", "white", "darkred"))(100),
  clustering_distance_rows = "euclidean",
  clustering_distance_cols = "euclidean",
  clustering_method = "complete",
  fontsize_row = 12,
  fontsize_col = 8,
  display_numbers = letter_mat,
  number_color = "black",
  main = "FSC ASV (Species) abundance + Dunn letters"
)

# 10. FIX LETTER ALIGNMENT AFTER CLUSTERING
row_order <- HM$tree_row$labels
col_order <- HM$tree_col$labels

mat_ordered <- mat[row_order, col_order]

letter_mat_ordered <- letter_mat[row_order, col_order]

# redraw correctly aligned heatmap
HM <- pheatmap(
  mat_ordered,
  scale = "column",
  color = colorRampPalette(c("#3B4CC0", "white", "#B40426"))(100),
    border_color = "black",
  clustering_distance_rows = "euclidean",
  clustering_distance_cols = "euclidean",
  clustering_method = "complete",
  fontsize_row = 12,
  fontsize_col = 8,
  display_numbers = letter_mat_ordered,
  number_color = "black",
  main = "FSC ASV (Species) abundance + Dunn letters"
)
HM
```

```{r, echo=FALSE, message=FALSE}
Plot_16S_Observed
Plot_16S_Shannon

Plot_ITS_Observed
Plot_ITS_Shannon

Final_plot_16S_Root
Final_plot_ITS_Root

tax16S
taxITS
```


---
END
---



