---
title: "Decoy sugars hijack fungal carbon sensing and prevent plant disease"
author: "Guillaume Chesneau"
date: "26-05-2026"
output: html_document
---

*Condition*.  

Biological parameters:    
- Plants grown in Agar-based system.     
- Harvested after ??
- Inoculated after ??
- Add informations....       
-....    

Pre-processing in Galaxy.     
- Quality checked with FastQC, Flatten, MultiFastQG.   
- Mapping using STAR.   
- Strandness checked with Infer Experiment (Reverse stranded library).     
- Feature counts (exons, fragments are counted).   
- DeSeq : All comparions made with Water-MF conditions:     
  - Water-MF_3OMG-MF.  
  - Water-MF_3OMG-F16.  
  - Water-MF_Water-F16.      
  

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "") 
```


# Before you start

This script has been made on R version 4.2.2 (2022-10-31)
```{r R version}
R.version
```

```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}
#Load Packages
library(ggplot2)
library(cowplot)
library(forcats)
library(ggpubr)
library(ggrepel)
library(tidyr)
library(DESeq2)
library(stringr)
library(vegan)
library(dplyr)
library(ggsignif)

# Prepare base theme
base_theme <- theme_bw() +
  theme(
    strip.background = element_rect(color = "white", fill = "white"),
    strip.text.x = element_markdown(size = 10, color = "black", hjust = 0.5, vjust = 0.5),
    legend.title = element_text(size = 11),
    axis.title.y = element_text(size =11),
    legend.text = element_text(size = 10),
    axis.text.x = element_text(size = 10, color = "black", angle = 90, hjust = 1),
    axis.text.y = element_text(size = 10, color = "black"),
    panel.border = element_rect(color = "black", fill = NA, size = 0.75))
```

# A] Root length

```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}

#Check quality
    #Open file
    df_length <- read.table(file="Transcripto_length.txt",sep="\t",dec=".",header=TRUE, row.names = 1)

#Order
df_length$Time <- factor(df_length$Time, levels = c ("72h", "168h"))
df_length$Supplementation <- factor(df_length$Supplementation, levels = c ("Water", "3OMG"))
df_length$Microbe <- factor(df_length$Microbe, levels = c ("M.free", "F16"))

# Assuming your dataframe is named df
Plot_length <- ggplot(df_length, aes(x = Microbe, y = Root_Length, fill = Supplementation)) +
     geom_jitter(alpha = 0.2, size = 0.5) +
  geom_boxplot(alpha=1,outlier.shape = NA) +
   facet_grid(~Time, space="free", scales="free")+
  labs(x = "", y = "Root Length (mm)", title = "Root Length by Microbe and Time Condition") +
        scale_fill_manual(values =c("#949494","#7FC29B")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + # To make x labels readable
  base_theme
Plot_length

# Perform Wilcoxon tests for each Time and Microbe combination
p_values <- df_length %>%
  group_by(Time, Microbe) %>%
  summarise(
    p_value = wilcox.test(Root_Length ~ Supplementation)$p.value,
    .groups = "drop"
  )

# Define significance stars based on p-values
p_values <- data.frame(p_values %>%
  mutate(
    significance = case_when(
      p_value < 0.001 ~ "***",
      p_value < 0.01 ~ "**",
      p_value < 0.05 ~ "*",
      TRUE ~ "ns"  # "ns" for no significance
    )
  ))

#Arragne talbe to match
p_values$Supplementation <- c("Water")

# Create the plot with significance stars
Plot_length <- ggplot(df_length, aes(x = Microbe, y = Root_Length, fill = Supplementation)) + 
geom_jitter(alpha = 0.2, size = 0.5, position = position_jitterdodge(dodge.width = 0.75, jitter.width = 0.2)) + 
  geom_boxplot(alpha = 1, outlier.shape = NA) + 
  facet_grid(~Time, space = "free", scales = "free") + 
  labs(x = "", y = "Root Length (mm)", title = "") + 
  scale_fill_manual(values = c("darkgray", "#7FC29B")) + 
  base_theme +
  # Add significance letters
    geom_text( data= p_values, mapping = aes(x = factor(Microbe), y = max(df_length$Root_Length), label = significance), size = 4)


# Print the plot
Plot_length
```

# B]  In Fungal Transcriptome

## 1) Normalize prepare data

```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}

    
#Open file
table_counts <- data.frame(read.table(file="Re Raw data transcriptome/count_table_P16.txt",sep="\t",dec=".",header=TRUE, row.names = 1))

#Remove Plant MF samples only
# Create a vector of column names to remove
cols_to_remove <- c("water_microbeFree_1",	"water_microbeFree_2",	"water_microbeFree_3",	"water_microbeFree_4","m3OMG_microbeFree_1",	"m3OMG_microbeFree_2",	"m3OMG_microbeFree_3",	"m3OMG_microbeFree_4")

# Remove these columns from your dataframe
table_counts2 <- table_counts[, !(names(table_counts) %in% cols_to_remove)]

#Normalization (rlog)
matrix_counts_Fungi <- as.matrix(table_counts2)
# Apply the vst transformation
#matrix_counts_Plant_rlog <- rlog(matrix_counts_Plant)
```



## 2) PcoA Bray-Curtis
```{r echo=FALSE, fig.show='hide', message=FALSE, results="hide"}
#Open files
df_transformed <- data.frame(matrix_counts_Fungi)

#Turn Table
df_transformed <- data.frame(t(df_transformed))

#Environements
env <- read.table(file="env_Fungi.txt",sep="\t",dec=".",header=TRUE, row.names = 1) #put env2 if filtered and env if not

data.div <- adonis2(df_transformed ~ Microbes/Molecule, data = env, permutations = 999, method="canberra") #Not significant between molecules will are looking therefore at slight differences
data.div


# Assuming your count matrix is 'count_data' and sample information 'sample_info'
dds <- DESeqDataSetFromMatrix(countData = matrix_counts_Fungi, colData = env, design = ~ Conditions)

# Variance-stabilizing transformation (VST)
#vsd <- vst(dds, blind = FALSE)

# Alternatively, you can use rlog transformation (rld)
rld <- rlog(dds, blind = FALSE)

# Perform PCA on the normalized data
pca_res <- prcomp(t(assay(rld)))

# Calculate the percentage of variance explained
percentVar <- round(100 * (pca_res$sdev^2 / sum(pca_res$sdev^2)), 2)

# Create a dataframe for the PCA plot
pca_df <- data.frame(PC1 = pca_res$x[, 1],
                     PC2 = pca_res$x[, 2],
                     Condition = env$Conditions)

# Extract molecule and microbes information from rownames (assuming you have this encoded in your sample names)
pca_df <- pca_df %>%
  mutate(Environment = str_split_fixed(row.names(pca_df), '_', 3)[, 1],
         Media = str_split_fixed(row.names(pca_df), '_', 3)[, 2],
         Environment_Media = paste(Environment, Media, sep = " - "))
pca_df


# Calculate the mean PCA coordinates for each Media/Microbe combination
mean_pca <- pca_df %>%
  group_by(Environment_Media) %>%
  summarise(PC1 = mean(PC1), PC2 = mean(PC2)) %>%
  separate(Environment_Media, into = c("Environment", "Media"))

# Prepare the segment data to link each replicate to the mean point
segments_pca <- pca_df %>%
  group_by(Environment, Media) %>%
  mutate(meanX = mean(PC1), meanY = mean(PC2))

#Add facet
pca_df$Habitat <- c("Fungal Transcriptome 72h")


# Plot the PCA results with mean points and segments
PcoA_Fung <- ggplot() +
  # Add jittered points for the individual replicates
  geom_point(data = pca_df, aes(x = PC1, y = PC2, color = Media, shape = Environment), size = 2, alpha = 1) +
  
  # Add mean points for each Microbe/Media combination
  geom_point(data = mean_pca, aes(x = PC1, y = PC2, color = Media, shape = Environment), size = 5) +
  
  # Add segments linking replicates to their corresponding mean points
  geom_segment(data = segments_pca, aes(x = meanX, y = meanY, xend = PC1, yend = PC2, color = Media), size = 0.7, alpha = 0.6) +
   facet_grid(~Habitat, space="free", scales="free")+
  # Customize colors and shapes as before
   scale_color_manual('Media',  values = c("#7FC29B", "#949494")) +
   scale_shape_manual('Environment', labels = c("Membrane", "Plant"), values = c(16, 17)) +
  # Add titles and labels with the explained variance
  xlab(paste("PC1 (", percentVar[1], "%)", sep = "")) +
  ylab(paste("PC2 (", percentVar[2], "%)", sep = "")) +
   base_theme
PcoA_Fung
```
## 3) Differential expressed gene analysis 

## a) Vulcano Plot

```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}
# Assuming `counts` is your count matrix and `colData` is your sample information
dds <- DESeqDataSetFromMatrix(countData = matrix_counts_Fungi, colData = env, design = ~ Conditions)


#Pre-filtering
smallestGroupSize <- 3
keep <- rowSums(counts(dds) >= 10) >= smallestGroupSize
dds <- dds[keep,]

#Run DeSeq2
dds <- DESeq(dds)

dds$Conditions
# Get results for m3OMG_F16 vs water_F16
results_Control <- results(dds, contrast = c("Conditions", "3OMG_Control", "Water_Control"))
results_Control <- as.data.frame(results_Control)

# Get results for m3OMG_MF vs water_MF
results_Plant <- results(dds, contrast = c("Conditions", "3OMG_Plant", "Water_Plant"))
results_Plant <- as.data.frame(results_Plant)

#Preapre plot
# Calculate -log10(p-value)
results_Control <- results_Control %>%
  mutate(logP = -log10(padj))
results_Plant <- results_Plant %>%
  mutate(logP = -log10(padj))

# Create a new column for color based on conditions
results_Control$Color <- ifelse(results_Control$log2FoldChange > 0.5 & results_Control$padj
 < 0.05, "#7FC29B",
                   ifelse(results_Control$log2FoldChange < -0.5 & results_Control$padj
 < 0.05, "#949494", "lightgray"))

results_Plant$Color <- ifelse(results_Plant$log2FoldChange > 0.5 & results_Plant$padj
 < 0.05, "#7FC29B",
                   ifelse(results_Plant$log2FoldChange < -0.5 & results_Plant$padj
 < 0.05, "#949494", "lightgray"))


#Add geom_text
#gene_annot_short <- read.table(file="Re Raw data transcriptome/Plecto_annot_Caz.txt",sep="\t",header=TRUE)
#gene_annot_short_des <- read.table(file="Re Raw data transcriptome/Plecto_annot_des.txt",sep="\t",header=TRUE)
gene_annot_short_full <- read.table(file="Re Raw data transcriptome/Plecto_annot_Caz_des.txt",sep="\t",header=TRUE)

results_Control$locus_name <- row.names(results_Control)
results_Controlf <- merge(results_Control, gene_annot_short_full, by = "locus_name", all.x = FALSE)
# Create the new column 'symbol_2'
#results_Controlf$symbol_2 <- ifelse(results_Controlf$Color != "lightgray", results_Controlf$symbol, NA)
# Remove duplicates from the "gene" column
#results_Controlf <- results_Controlf %>% distinct(locus_name, .keep_all = TRUE)

#Add geom_text
results_Plant$locus_name <- row.names(results_Plant)
results_Plantf <- merge(results_Plant, gene_annot_short_full, by = "locus_name")
# Create the new column 'symbol_2'
#results_Plantf$symbol_2 <- ifelse(results_Plantf$Color != "lightgray", results_MFf$symbol, NA)
# Remove duplicates from the "gene" column
#results_Plantf <- results_Plantf %>% distinct(locus_name, .keep_all = TRUE)

#Control
# Get the top 20 most significant positive Log2FC
    top_pos_control <- results_Controlf %>%
      filter(log2FoldChange > 0.5) %>%  # Positive Log2FC
      arrange(desc(logP)) %>%
      head(15)

    # Get the top 20 most significant negative Log2FC
    top_neg_control <- results_Controlf %>%
      filter(log2FoldChange < -0.5) %>%  # Negative Log2FC
      arrange(desc(logP)) %>%
      head(15)

    # Combine the top positive and negative loci
    top_loci_Control <- bind_rows(top_pos_control, top_neg_control)
    
#Plant
    # Replace the CMB20 to show it, p-value abve limit
results_Plantf$logP[is.infinite(results_Plantf$logP)] <- max((results_Plantf$logP+1)[!is.infinite(results_Plantf$logP)])

# Get the top 20 most significant positive Log2FC
    top_pos_plant <- results_Plantf %>%
      filter(log2FoldChange > 0.5) %>%  # Positive Log2FC
      arrange(desc(logP)) %>%
      head(50)

    # Get the top 20 most significant negative Log2FC
    top_neg_plant <- results_Plantf %>%
      filter(log2FoldChange < -0.5) %>%  # Negative Log2FC
      arrange(desc(logP)) %>%
      head(50)

    # Combine the top positive and negative loci
    top_loci_Plant <- bind_rows(top_pos_plant, top_neg_plant)
    
# Create the volcano plot

# Remove rows where log2FC is greater than -5
#results_F16f <- results_F16f[results_F16f$log2FoldChange >= -5, ]

# Add Facet
results_Controlf$Habitat <- c("3OMG Membrane vs  Water Membrane")
    
Vulc_Control <- ggplot(results_Controlf, aes(x = log2FoldChange, y = logP)) +
   geom_point(aes(color = Color), size = 2)+  # Color points based on p-value
#geom_text(vjust = 1.5, hjust = 0.5, size = 3) +     # Add gene labels
    scale_color_identity() +
  labs(title = "",
       x = "Log2 Fold Change",
       y = "-Log10(p-value)") +
#ylim(0,250) +
        facet_grid(~Habitat, space="free", scales="free")+
geom_text_repel(
        data = top_loci_Control,  # Only label the top 15 positive and negative Log2FC loci
        aes(label = CAZYName),
        size = 3,
        max.overlaps = 50,  # Control text overlap
        box.padding = 0.5,  # Increase this to add more space between labels and points
        point.padding = 0.5,  # Increase space between points and labels
        segment.size = 0.2,  # Adjust the line size connecting points to labels
        segment.color = "grey50"  # Change color of connecting lines
      ) +
  base_theme +
   geom_vline(xintercept = c(-0.5, 0.5), linetype = "dashed", color = "darkgray", size = 0.5) +
      geom_hline(yintercept = -log10(0.05), linetype = "dashed", color = "darkgray", size = 0.5) +
      scale_color_identity() 

Vulc_Control



# Remove rows where log2FoldChange is greater than 10
#results_MFf <- results_MFf[results_MFf$log2FoldChange <= 10, ]

# Add Facet
results_Plantf$Habitat <- c("3OMG Plant vs  Water Plant")
    

Vulc_Plant <- ggplot(results_Plantf, aes(x = log2FoldChange, y = logP)) +
   geom_point(aes(color = Color), size = 2)+  # Color points based on p-value
#geom_text(vjust = 1.5, hjust = 0.5, size = 3) +     # Add gene labels
    scale_color_identity() +
  labs(title = "",
       x = "Log2 Fold Change",
       y = "-Log10(p-value)") +
  # ylim(0,250) +
       facet_grid(~Habitat, space="free", scales="free")+
geom_text_repel(
        data = top_loci_Plant,  # Only label the top 15 positive and negative Log2FC loci
        aes(label = CAZYName),
        size = 3,
        max.overlaps = 50,  # Control text overlap
        box.padding = 0.5,  # Increase this to add more space between labels and points
        point.padding = 0.5,  # Increase space between points and labels
        segment.size = 0.2,  # Adjust the line size connecting points to labels
        segment.color = "grey50"  # Change color of connecting lines
      ) +
  base_theme +
   geom_vline(xintercept = c(-0.5, 0.5), linetype = "dashed", color = "darkgray", size = 0.5) +
      geom_hline(yintercept = -log10(0.05), linetype = "dashed", color = "darkgray", size = 0.5) +
      scale_color_identity() 

Vulc_Plant
  


#final Plot
Plot_Comb1 <- plot_grid(Vulc_Control, Vulc_Plant, rel_heights = c(1,1), rel_widths = c(1.1,1), nrow=1,   align="hv",  axis = "tb")
Plot_Comb1



#extract tables
#write.csv(results_Controlf, file = "results_Fungi_Control.csv", row.names = TRUE)
#write.csv(results_Plantf, file = "results_Fungi_Plant.csv", row.names = TRUE)
```
## b) Count CAZymes

```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}

# Count NA and non-NA entries for each 'Color'
table_Cazymes_Control <- results_Controlf %>%
  group_by(Color) %>%
  summarise(
    NA_count = sum(is.na(CAZYName)),
    Non_NA_count = sum(!is.na(CAZYName)),
    Mean_Log2FC = mean((log2FoldChange), na.rm = TRUE)  # Calculate the mean of Log2FC, excluding NAs
  )
table_Cazymes_Control$Condition <- c("Control")

table_Cazymes_Plant <- results_Plantf %>%
  group_by(Color) %>%
  summarise(
    NA_count = sum(is.na(CAZYName)),
    Non_NA_count = sum(!is.na(CAZYName)),
    Mean_Log2FC = mean((log2FoldChange), na.rm = TRUE)  # Calculate the mean of Log2FC, excluding NAs
  )
table_Cazymes_Plant$Condition <- c("Plant")

#Merge tables
table_Cazymes <- rbind(table_Cazymes_Control, table_Cazymes_Plant)

#Check CAzynmaes Log2FC
# Keep only rows where CAZYName is not NA
results_Controlf_f <- results_Controlf %>%
  filter(!is.na(CAZYName))
results_Controlf_f$Condition <- c("Control")

results_Plantf_f <- results_Plantf %>%
  filter(!is.na(CAZYName))
results_Plantf_f$Condition <- c("Plant")

#Merge tables
table_Cazymes_Tot <- rbind(results_Controlf_f, results_Plantf_f)

#remove non signifciant
table_Cazymes_Tot_sig <- subset(table_Cazymes_Tot, !Color %in% c("lightgray"))

# Create a new column based on log2FoldChange values and the corresponding condition
table_Cazymes_Tot_sig <- table_Cazymes_Tot_sig %>%
 mutate(Group = case_when(
     log2FoldChange >= 0 ~ "OV-3OMG",  # Above 0, label as OV-3OMG
    log2FoldChange < 0 ~ "OC-Water",  # Below 0, label as OC-Water
    TRUE ~ NA_character_  # Assign NA for all other cases
  ))

#put absolut Log2FC
table_Cazymes_Tot_sig$log2FoldChange <- abs(table_Cazymes_Tot_sig$log2FoldChange)


# Create a boxplot
# Calculate the number of replicates for each group and condition
count_data <- table_Cazymes_Tot_sig %>%
  group_by(Group, Condition) %>%
  summarize(Count = n(), .groups = 'drop')  # Count the number of rows for each group and condition
count_data$log2FoldChange <- c(0)

# Check normality per Group x Condition
shapiro <- table_Cazymes_Tot_sig %>%
  group_by(Condition, Group) %>%
  summarise(shapiro_p = shapiro.test(log2FoldChange)$p.value) #Not normal Wilcoxon
shapiro

# Prepare comparisons for bracket plotting
comparisons <- list(c("OC-Water", "OV-3OMG"))

# Define comparisons
comparisons <- list(c("OC-Water", "OV-3OMG"))

# Ensure Group is a factor with correct order
table_Cazymes_Tot_sig$Group <- factor(table_Cazymes_Tot_sig$Group, levels = c("OC-Water", "OV-3OMG"))

# Compute Wilcoxon test per Condition
stat_test <- table_Cazymes_Tot_sig %>%
  group_by(Condition) %>%
  wilcox_test(log2FoldChange ~ Group) %>%
  adjust_pvalue(method = "BH") %>%       # adjust p-values if needed
  add_significance("p.adj") %>%          # optional significance annotation
  mutate(
    y.position = 7.5,
    label_text = ifelse(p.adj <= 0.05, paste0("P = ", signif(p.adj, 2)), NA)  # only show significant
  )

# Create a column with "P = ..." formatting
stat_test <- stat_test %>%
  mutate(label_text = paste0("P = ", signif(p.adj, 3)))

#Add column for plottig
stat_test$Group <- c("OV-3OMG")

# Summarize number of samples per Group × Condition
n_labels <- table_Cazymes_Tot_sig %>%
  group_by(Group, Condition) %>%
  summarise(n = n(), .groups = "drop") %>%
  # Position slightly below the minimum of y-axis
  mutate(y = min(abs(table_Cazymes_Tot_sig$log2FoldChange)) - 1)

# Main plot
Plot_FC_CAZ <- ggplot(table_Cazymes_Tot_sig, aes(x = Group, y = abs(log2FoldChange), fill = Group)) +
  geom_boxplot(alpha = 1, outlier.shape = NA) +
  geom_jitter(width = 0.2, size = 1, alpha = 0.7) +
  facet_grid(~ Condition, labeller = labeller(Condition = c("Control" = "Polyester Membrane", "Plant" = "In Planta"))) +
  labs(
    x = "",
    y = "Cumulative Log2 Fold Change CAZymes (|Log2FC|)",
    title = ""
  ) +
 scale_fill_manual(
  name = "Condition",    # Legend title
  values = c("#949494", "#7FC29B"),
  labels = c("OC-Water" = "Water",
             "OV-3OMG" = "3OMG")
) +
  theme_bw(base_size = 13) +
  theme(
    strip.background = element_rect(color = "black", fill = "white"),
    strip.text.x = element_text(size = 12),
    axis.text.x = element_blank()
  ) +
  # Add p-values with brackets
  stat_pvalue_manual(
    stat_test,
    label = "label_text",
    tip.length = 0.02,
    size = 4
  ) +
  # Add sample size labels
  geom_text(
    data = n_labels,
    aes(x = Group, y = y, label = paste0("n = ", n)),
    inherit.aes = FALSE,
    size = 4,
    vjust = 1
  ) +
  ylim(-1,8) +
  base_theme

Plot_FC_CAZ
```

## c) Caz families COunt Families

```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}
#Open file
Caz_Sub <- read.table(file="CaZ_Subtrate.txt",sep="\t",dec=".",header=TRUE, row.names = 1)

unique(Caz_Sub$Substrates2)

#Vector dependning of categories of Cazymes
Plant_colonization <- c("cellulose", "pectin", "xylan", "beta-mannan", "arabinan", "xyloglucan", 
  "arabinogalactan protein", "lignin", "beta-glucan", "cutin", "polyphenol")

Component_plant <- c("cellulose", "pectin", "xylan", "beta-mannan", "arabinan", "xyloglucan", 
  "arabinogalactan protein", "lignin", "cutin", "starch", "sucrose", 
  "beta-glucan", "raffinose", "beta-galactan")

Not_relevant <- c("alpha-galactan", "alpha-glucan", "glycogen", "trehalose", "host glycan", 
  "chitin", "chitosan", "exo-polysaccharide", "human milk polysaccharide", 
  "fructan", "glucosylglycerate", "glucosylglycerol", "chitooligosaccharide", 
  "sialic acid", "uric acid", "ulvan", "alpha-rhamnoside", "beta-fucosides", 
  "alkaloid", "beta-glucuronan")


        
Caz_Sub_cleaned <- Caz_Sub %>%
  filter(Substrates2 %in% Component_plant)


#Merge tables
Caz_tot_Sub <- merge(table_Cazymes_Tot_sig, Caz_Sub_cleaned, by= c("CAZYName"))
Caz_tot_Sub


# Assuming you have a dataframe called df with 'Substrates', 'Condition', and 'Group'

# 1. Count the number of occurrences of each substrate by group and condition
count_data <- Caz_tot_Sub %>%
  group_by(Group, Condition, Substrates2) %>%
  tally() %>%
  ungroup()

# 2. Normalize to percentages (0-100)
total_count <- count_data %>%
  group_by(Group, Condition) %>%
  summarize(total = sum(n)) %>%
  ungroup()

count_data <- count_data %>%
  left_join(total_count, by = c("Group", "Condition")) %>%
  mutate(percentage = (n / total) * 100)

# Create a color gradient with 18 colors
gradient_colors <- colorRampPalette(c("#FF0000", "#FF7F00", "#FFFF00", "#00FF00", "#0000FF", "#4B0082", "#8A2BE2", "#FF00FF", "#FF1493", "#FF6347", "#FFD700", "#B0E0E6", "#90EE90", "#ADFF2F", "#8B0000", "#C71585", "#000080", "#32CD32", "pink"))(19)


# 3. Create the stacked bar plot
Plot_Subsrates <- ggplot(count_data, aes(x = Group, y = percentage, fill = Substrates2)) +
  geom_bar(stat = "identity", position = "stack") +
  facet_grid(~ Condition, labeller = labeller(Condition = c("Control" = "Membrane"))) +  #
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  labs(y = "Percentage (%)", x = "Condition", fill = "Substrates") +
  scale_fill_manual(values =c(gradient_colors)) +
  theme(
    strip.background = element_rect(color = "black", fill = "white", size = 1, linetype = "solid"),
    strip.text.x = element_text(size = 12, color = "black"),
    strip.text.y = element_text(size = 12, color = "black"),
    axis.text.x = element_text(angle = 90, hjust = 1)
  ) 
Plot_Subsrates
```




## d) Caz families COunt Families

```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}
#Open file
table_Cazymes_Tot_sig

#Merge tables
Caz_tot_Sub <- merge(table_Cazymes_Tot_sig, Caz_Sub_cleaned, by= c("CAZYName")) #Caz_Sub_cleaned
Caz_tot_Sub
Caz_tot_Sub$concat <- paste(Caz_tot_Sub$Substrates2, "_", Caz_tot_Sub$Group, "_", Caz_tot_Sub$Condition)

# Assuming your data is in a dataframe called 'df'
# Summarize and calculate cumulative log2FC by Substrate, Group, and Condition
df_cumulative <- Caz_tot_Sub %>%
  group_by(Substrates2, Group, Condition) %>%
  summarise(
    CumulativeLog2FC = sum(log2FoldChange),       # Summing log2FoldChange
    meanpvalue = sum(logP, na.rm = TRUE),        # Calculating the mean p-value
    .groups = "drop"
  )



# Modify the CumulativeLog2FC values for the "OC-Water" group to be negative
df_cumulative <- df_cumulative %>%
  mutate(CumulativeLog2FC = ifelse(Group == "OC-Water", -abs(CumulativeLog2FC), CumulativeLog2FC))
```

```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}
#Try to merge with count data
count_data <- data.frame(count_data)
count_data$merge <- paste0(count_data$Group, "_", count_data$Condition, "_", count_data$Substrates2)

# Remove the individual columns used to create 'merge'
count_data <- count_data[, !(names(count_data) %in% c("Group", "Condition", "Substrates2"))]
df_cumulative$merge <- paste0(df_cumulative$Group, "_", df_cumulative$Condition, "_", df_cumulative$Substrates2)


Final_table <- merge(df_cumulative, count_data, by = c("merge"))


# Reorder Substrates2 based on n
Final_table <- Final_table %>%
  mutate(Substrates2 = fct_reorder(Substrates2, n, .desc = FALSE))

#Order manually
unique(Final_table$Substrates2 )
Final_table$Substrates2 <- factor(Final_table$Substrates2, levels = c ("raffinose", "cutin", "starch", "beta-mannan", "xyloglucan", "arabinan", "lignin", "arabinogalactan protein", "beta-galactan", "beta-glucan", "xylan", "cellulose", "pectin"))

Plot_Cum_log2FCf <- ggplot(Final_table, aes(x = CumulativeLog2FC, y = Substrates2, color = Group, group = Group)) + 
  geom_segment(aes(xend = 0, yend = Substrates2), size = 0.5) +  # Draw vertical lines from each point to the x-axis
  geom_point(aes(size = n)) +  # Add points for each substrate
  facet_grid(~ Condition, labeller = labeller(Condition = c("Control" = "Membrane"))) +  
  labs(
    x = "Cumulative Log2FoldChange CAZymes", 
    y = "Predicted Substrates CAZymes") + 
  theme(
    strip.background = element_rect(color = "black", fill = "white", size = 1, linetype = "solid"),
    strip.text.x = element_text(size = 12, color = "black"),
    strip.text.y = element_text(size = 12, color = "black"),
    axis.text.x = element_text(angle = 90, hjust = 1)
  ) +
  scale_color_manual('Media', values = c("#949494", "#7FC29B"), labels = c("Water", "3OMG")) +
  scale_size_continuous(name = "Counts CAZymes") +
  geom_text(aes(
    label = n,
    vjust = ifelse(Group == "OC-Water", 0.5, 0.5),
    hjust = ifelse(Group == "OC-Water", 2.2, -1.5)
  ), size = 4) +
  xlim(-30, 20) +
    scale_size_continuous(name = "Counts CAZymes", range = c(0.01, 6)) +
   geom_vline(xintercept = 0, color = "black", size = 0.5)

Plot_Cum_log2FCf

#Plot for counts

Final_table$n2 <- ifelse(Final_table$Group == "OC-Water", -Final_table$n, Final_table$n)

Plot_Cum_log2FCf2_Dot <- ggplot(Final_table, aes(x = n2, y = Substrates2, color = Group, group = Group)) + 
  geom_segment(aes(xend = 0, yend = Substrates2), size = 0.5) +  # Draw vertical lines from each point to the x-axis
  geom_point(aes(size = abs(CumulativeLog2FC)))+  # Add points for each substrate
  facet_grid(~ Condition, labeller = labeller(Condition = c("Control" = "Membrane"))) +  
  labs(
    x = "Counts CAZymes", 
    y = "Predicted Substrates CAZymes"
  ) + 
  base_theme +
  scale_color_manual('Media', values = c("#949494", "#7FC29B"), labels = c("Water", "3OMG")) +
  geom_text(aes(
    label = n,
    vjust = ifelse(Group == "OC-Water", 0.5, 0.5),
    hjust = ifelse(Group == "OC-Water", 2.2, -1.5)
  ), size = 3) + 
  xlim(-20, 20) +
   scale_size_continuous(name = "Mean Log2FC", range = c(0.01, 6))   # Increase the size range here

Plot_Cum_log2FCf2_Dot
```

```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}
# Ensure all combinations of Substrates2, Group, and Condition are present
Final_table_complete <- Final_table %>%
  complete(Substrates2, Group, Condition, fill = list(n = 0))  # Fill missing combinations with n = 0

Plot_Cum_log2FCf <- ggplot(Final_table_complete, aes(x =Substrates2 , y = n , fill = Group)) +
  geom_col(position = position_dodge()) +  # Create bar plot
  facet_grid(~ Condition, labeller = labeller(Condition = c("Control" = "Membrane"))) +
  labs(
    y = "Counts CAZymes", 
    x = "Predicted Substrates CAZymes"
  ) +
  base_theme +
  scale_fill_manual('Media', values = c("#949494", "#7FC29B"), labels = c("Water", "3OMG")) +
  geom_text(aes(
    label = n,
    vjust = ifelse(Group == "OC-Water", -0.5, -0.5)
  ), size = 3, position = position_dodge(width = 0.9)) +
  ylim(0,16)

Plot_Cum_log2FCf

Plot_Cum_log2FCf2 <- ggplot(Final_table_complete, aes(x =Substrates2 , y = CumulativeLog2FC , fill = Group)) +
  geom_col(position = position_dodge()) +  # Create bar plot
  facet_grid(~ Condition, labeller = labeller(Condition = c("Control" = "Membrane"))) +
  labs(
    y = "Counts CAZymes", 
    x = "Predicted Substrates CAZymes"
  ) +
  base_theme +
  scale_fill_manual('Media', values = c("#949494", "#7FC29B"), labels = c("Water", "3OMG")) +
  geom_text(aes(
    label = round(CumulativeLog2FC,2),
  vjust = ifelse(Group == "OC-Water", -0.5, -0.5)
  ), size = 3, position = position_dodge(width = 0.9)) +
  ylim(-6,6)

Plot_Cum_log2FCf2

```

# C] All plots fungi

```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}
PcoA_Fung

Vulc_Control
Vulc_Plant

Plot_FC_CAZ
Plot_Cum_log2FCf2_Dot

# Empty placeholder plot
empty_plot <- ggplot() + 
  theme_void()  # completely blank

Plot_Comb1 <- plot_grid(empty_plot, Vulc_Control, Vulc_Plant,empty_plot,  frel_heights = c(1,1), rel_widths = c(0.3,1,1, 0.3), nrow=1, ncol=4)
Plot_Comb1

#Export 8X9 Landscape
Plot_Comb2 <- plot_grid(
  Plot_Comb1, Plot_Cum_log2FCf2_Dot,
  ncol = 1,               # stack vertically
  rel_heights = c(1, 0.7) # top plot bigger than bottom
)
Plot_Comb2
```

---
END
---