---
title: "Decoy sugars hijack fungal carbon sensing and prevent plant disease"
author: "Guillaume Chesneau"
date: "26-05-2026"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/Post_doc/Postdoc/Scripts/2-Biolog/Biolog_Plant_FSC") 
```


# Before you start

This script has been made on R 4.5.3 (and Rstudio 2023.09.1+494)

```{r R version}
R.version
```

```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}
#Load Packages
library(rstatix)  # For statistical analysis
library(ggpubr)   # For creating publication-ready plots using ggplot2
library(dunn.test)
library(ggthemes)
library(ggtext)
library(tidyr)
library(markdown)
library(cowplot)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(ggrepel)


# Prepare base theme
base_theme <- theme_bw() +
  theme(
    strip.background = element_rect(color = "white", fill = "white"),
    strip.text.x = element_markdown(size = 10, color = "black", hjust = 0.5, vjust = 0.5),
    legend.title = element_text(size = 11),
    axis.title.y = element_text(size =11),
    legend.text = element_text(size = 10),
    axis.text.x = element_text(size = 10, color = "black", angle = 90, hjust = 1),
    panel.border = element_rect(color = "black", fill = NA))
```


# A] Screen Biolog

## 1 - Prepare data

### a) Load raw_data
```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}
Data <- read.table(file="Table_Df.txt",sep="\t",dec = ".",header=TRUE, row.names=1)
```

### b) Integrate Sources
```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}
Sources_PM1 <- read.csv("PM01.csv", header= TRUE)
colnames(Sources_PM1)[2] <- c("Well")
Data_PM1 <- merge(Data, Sources_PM1, by = c("Well", "Plate"))  

Sources_PM2 <- read.csv("PM02A.csv", header= TRUE)
colnames(Sources_PM2)[2] <- c("Well")
Data_PM2 <- merge(Data, Sources_PM2, by = c("Well", "Plate"))  

#Bind datasets
Data_final <- rbind(Data_PM1,Data_PM2)
```

### c) Veen diagramm carbon sources
```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}
#Bind files
Sources_PM_tot <- rbind(Sources_PM1, Sources_PM2)

# Summarize counts per family
family_counts <- Sources_PM_tot %>%
  group_by(Family_compounds) %>%
  summarise(count = n()) %>%
  mutate(label = paste0(count))  # optional: add count as label

# Simple horizontal bar plot
bar_plot <- ggplot(family_counts, aes(x = reorder(Family_compounds, count), y = count, fill = Family_compounds)) +
  geom_bar(stat = "identity", color = "black", fill = "white") +  # white bars with black borders
  geom_text(aes(label = label), hjust = -0.1, size = 4) +        # labels outside the bar
  coord_flip() +                                                  # horizontal bars
  theme_minimal() +
  theme(legend.position = "none") +                               # hide legend
  labs(x = "Carbon Source Families", y = "Number of Carbon Source", title = "Number of Compounds per Family")
bar_plot
```

### c) Germination Rate

I calculate the germination rate for each condition for each carbon source

```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}
Data_final <- Data_final %>%
  group_by( Microbe,Compound) %>%
  mutate(
    Germination_rate = mean(Germination, na.rm = TRUE) * 100
  ) %>%
  ungroup()
```


## 2 - Fungal dataset

### a) Calculate Ratio

I calculate the ratio of the FSC phennotype in "Negative Control" / FSC Phenotype in "Carbon Sources".
The objective is to look at the impact of the Carbon Sources in protection against Fungal Pathogens
I'm using the MiniSynCom compoased of 4 fungal root pathogens.

```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}

# Filter to only FSC rows
FSC_df <- Data_final %>%
  filter(Microbe == "FSC")

#Ratio Relative Shoot Area (+F) C vs NC
# Step 1: Compute median White.pixels for Negative Control within FSC
neg_control_median_FSC <- FSC_df %>%
  filter(Microbe == "FSC", Compound == "Negative Control") %>%
  summarise(neg_median = median(White.pixels, na.rm = TRUE)) %>%
  pull(neg_median)


# Step 2: Compute per-replicate log2 normalized values for FSC only,
# normalized by Negative Control median in FSC
FSC_df <- FSC_df %>%
  mutate(log2_norm = ifelse(
    Microbe == "FSC",
    log2(White.pixels / neg_control_median_FSC),
    NA
  ))
```

### b) Plot Fungi

```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}
# Test for normality per group
normality_results <- FSC_df %>%
  group_by(Compound) %>%
  shapiro_test(log2_norm) %>%   # this returns a tibble with 'stat' and 'p'
  mutate(
    normal = p > 0.05           # TRUE if p > 0.05, FALSE otherwise
  )

normality_results # 11% not normal I perform therefore t-test

# Pairwise t-test vs Negative Control
t_results <- FSC_df %>%
  t_test(log2_norm ~ Compound, ref.group = "Negative Control", p.adjust.method = "bonferroni", alternative = "two.sided"  ) %>%
  add_significance("p.adj") %>%
  mutate(
    stars = case_when(
      p <= 0.001 ~ "***",
      p <= 0.01  ~ "**",
      p <= 0.05  ~ "*",
      TRUE ~ ""
    )
  )

# Rename columns to match previous format
colnames(t_results) <- c("y", "Control", "Compound", "n1", "n2", 
                         "statistic", "df", "p", "p.adj", "p.adj.signif", "stars")

# Summarize log2-normalized values
summary_df_FSC <- FSC_df %>%
  group_by(Compound) %>%
  summarise(
    median_log2 = median(log2_norm, na.rm = TRUE),
    se_log2 = sd(log2_norm, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  ) %>%
  mutate(
    Compound_reordered = reorder(Compound, median_log2)
  )

# Merge significance into summary_df_FSC
summary_df_FSC <- summary_df_FSC %>%
  left_join(t_results %>% dplyr::select(Compound, stars), by = "Compound")

# Reorder FSC_df for jitter
FSC_df <- FSC_df %>%
  mutate(Compound_reordered = reorder(Compound, log2_norm))

# Add horizontal line for Negative Control
neg_se <- summary_df_FSC %>%
  filter(Compound == "Negative Control") %>%
  summarise(
    median_val = median_log2,
    ymin = median_log2 - se_log2,
    ymax = median_log2 + se_log2
  )

# Add colors
summary_df_FSC <- summary_df_FSC %>%
  mutate(
    bar_color = case_when(
      stars != "" ~ "#7FC29B",                       # highlight significant bars
      median_log2 < neg_se$ymin ~ "#817E9F",        # downregulated
      median_log2 > neg_se$ymax ~ "#D3E8DA",        # upregulated
      TRUE ~ "darkgray"                              # neutral
    )
  )

# Plot
Plot_FSC <- ggplot() +
  geom_bar(data = summary_df_FSC,
           aes(x = Compound_reordered, y = median_log2, fill = bar_color),
           stat = "identity", width = 0.7, show.legend = FALSE, size = 0.7) +
  geom_errorbar(data = summary_df_FSC,
                aes(x = Compound_reordered, ymin = median_log2 - se_log2, ymax = median_log2 + se_log2),
                width = 0.2) +
  geom_jitter(data = FSC_df,
              aes(x = Compound_reordered, y = log2_norm),
              width = 0.15, alpha = 0.5, size = 1, color = "black") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray40") +
  geom_rect(aes(xmin = -Inf, xmax = Inf, ymin = neg_se$ymin, ymax = neg_se$ymax),
            fill = "black", alpha = 0.1) +
  scale_fill_identity() +
  scale_color_identity() +
  scale_x_discrete(position = "top") +
  labs(x = "",
       y = expression( ~ log[2]("Shoot Area + FSC (Carbon / No Carbon)")),
       title = "") +
  theme(
    axis.text.x = element_text(
      angle = 90,
      hjust = 0,
      vjust = 0.5,
      size = 5,
      margin = margin(b = 5)
    )
  ) +
  geom_text(data = summary_df_FSC,
            aes(x = Compound_reordered, y = 4.2, label = stars),
            vjust = 0, size = 6, angle =90) +
  geom_segment(
    data = summary_df_FSC[summary_df_FSC$Compound_reordered == "Negative Control", ],
    aes(x = Compound_reordered, xend = Compound_reordered, y = max(FSC_df$log2_norm), yend = min(FSC_df$log2_norm)),
    color = "black",
    linetype = "dashed",
    size = 0.5
  ) +
  ylim(-2,4.4) +
  base_theme +
  theme(
    axis.text.x = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank()
  )

Plot_FSC

```

### c) Plot Fungi Germination

```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}
FSC_df$Germination_rate

# Reorder FSC_df for jitter
FSC_df <- FSC_df %>%
  mutate(Compound_reordered = reorder(Compound, log2_norm))

GR_FSC <- ggplot(FSC_df, aes(x = Compound_reordered, y = "", fill = Germination_rate)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "#A88B6D", mid = "#A88B6D", high = "darkgreen", midpoint = 0, name = "GR FSC")+
  labs(
    title = "",
    x = "Compound",
    y = "GR FSC",
    fill = "%"
  ) + 
  base_theme +
  theme(axis.text.x = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
        axis.ticks.y = element_blank())

GR_FSC
```

## 3 - MF dataset

### a) Calculate Ratio

I calculate the ratio of the MF phennotype in "Negative Control" / MF Phenotype in "Carbon Sources".
The objective is to look at the impact of the Carbon Sources in protection against Fungal Pathogens
I'm using the MiniSynCom compoased of 4 fungal root pathogens.

```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}
# Filter to only MF rows
MF_df <- Data_final %>%
  filter(Microbe == "MF")

# Step 1: Compute median White.pixels for Negative Control within MF
neg_control_median_MF <- MF_df %>%
  filter(Microbe == "MF", Compound == "Negative Control") %>%
  summarise(neg_median = median(White.pixels, na.rm = TRUE)) %>%
  pull(neg_median)

# Step 2: Compute per-replicate log2 normalized values for MF only,
# normalized by Negative Control median in MF
MF_df <- MF_df %>%
  mutate(log2_norm = ifelse(
    Microbe == "MF",
    log2(White.pixels / neg_control_median_MF),
    NA
  ))
```

### b) Plot MF 
```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}
# Test for normality per compound
normality_results <- MF_df %>%
  group_by(Compound) %>%
  shapiro_test(log2_norm) %>%
  mutate(normal = p > 0.05)

normality_results  # inspect normality; TRUE = approximately normal

# Since most are normal, perform t-test vs Negative Control
t_results <- MF_df %>%
  t_test(log2_norm ~ Compound, ref.group = "Negative Control", p.adjust.method = "bonferroni") %>%
  add_significance("p.adj") %>%
  mutate(
    stars = case_when(
      p <= 0.001 ~ "***",
      p <= 0.01  ~ "**",
      p <= 0.05  ~ "*",
      TRUE ~ ""     # non-significant
    )
  )

# Rename columns to match previous format
colnames(t_results) <- c("y", "Control", "Compound", "n1", "n2",
                         "statistic", "df", "p", "p.adj", "p.adj.signif", "stars")

# Summarize log2-normalized values
summary_df <- MF_df %>%
  group_by(Compound) %>%
  summarise(
    median_log2 = median(log2_norm, na.rm = TRUE),
    se_log2 = sd(log2_norm, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  ) %>%
  mutate(
    Compound_reordered = reorder(Compound, median_log2)
  )

# Merge significance into summary_df
summary_df <- summary_df %>%
  left_join(t_results %>% dplyr::select(Compound, stars), by = "Compound")

# Reorder MF_df for jitter
MF_df <- MF_df %>%
  mutate(Compound_reordered = reorder(Compound, log2_norm))

# Add horizontal line for Negative Control
neg_se <- summary_df %>%
  filter(Compound == "Negative Control") %>%
  summarise(
    median_val = median_log2,
    ymin = median_log2 - se_log2,
    ymax = median_log2 + se_log2
  )

# Add colors: only highlight significant bars
summary_df <- summary_df %>%
  mutate(
    bar_color = case_when(
      stars != "" ~ "#817E9F",          # significant
      TRUE ~ "darkgray"                  # non-significant neutral
    )
  )

# Order compounds based on FSC condition (if needed)
compound_order <- summary_df_FSC %>% arrange(median_log2) %>% pull(Compound)
MF_df$Compound <- factor(MF_df$Compound, levels = compound_order)
summary_df$Compound <- factor(summary_df$Compound, levels = compound_order)

# Plot
Plot_MF <- ggplot() +
  geom_bar(data = summary_df,
           aes(x = Compound, y = median_log2, fill = bar_color),
           stat = "identity", width = 0.7, show.legend = FALSE, size = 0.7) +
  geom_errorbar(data = summary_df,
                aes(x = Compound, ymin = median_log2 - se_log2, ymax = median_log2 + se_log2),
                width = 0.2) +
  geom_jitter(data = MF_df,
              aes(x = Compound, y = log2_norm),
              width = 0.15, alpha = 0.5, size = 1, color = "black") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray40") +
  geom_rect(aes(xmin = -Inf, xmax = Inf, ymin = neg_se$ymin, ymax = neg_se$ymax),
            fill = "black", alpha = 0.1) +
  scale_fill_identity() +
  scale_color_identity() +
  scale_x_discrete(position = "top") +
  labs(x = "",
       y = expression(log[2]~"Shoot Area + MF (Carbon / No Carbon)"),
       title = "") +
  geom_text(data = summary_df %>% filter(stars != ""),  # only plot significant stars
            aes(x = Compound_reordered, y = 1.8, label = stars),
            vjust = 0, size = 6, angle =90) +
  geom_segment(
    data = summary_df[summary_df$Compound_reordered == "Negative Control", ],
    aes(x = Compound_reordered, xend = Compound_reordered,
        y = max(MF_df$log2_norm), yend = min(MF_df$log2_norm)),
    color = "black",
    linetype = "dashed",
    size = 0.5
  ) +
  ylim(-4.2,2) +
  base_theme +
  theme(axis.text.x = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank())

Plot_MF
```

### c) Plot MF Germination

```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}
# Order compounds based only on the FSC condition
compound_order <- summary_df %>%
  arrange(median_log2) %>%
  pull(Compound)
MF_df$Compound <- factor(MF_df$Compound, levels = compound_order)

#Plot
GR_MF <- ggplot(MF_df, aes(x = Compound, y = "", fill = Germination_rate)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "#A88B6D", mid = "#A88B6D", high = "darkgreen", midpoint = 0, name = "GR MF")+
  labs(
    title = "",
    x = "Compound",
    y = "GR MF",
    fill = "%"
  )  + 
  base_theme +
  theme(axis.text.x = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
        axis.ticks.y = element_blank())

GR_MF
```

# B] Final Plot
```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}
#Plot Biolog
Plot_FSC
Plot_MF

#Plot Biolog with legend
Plot_FSC_Legend <- Plot_FSC  + theme( axis.text.x = element_text(size = 6, color = "black", angle =90, hjust= 0))
Plot_MF_Legend <- Plot_MF + theme( axis.text.x = element_text(size = 6, color = "black", angle =90, hjust= 0))

#Plot Biolog zoom
zoom_FSC <- Plot_FSC +coord_cartesian(xlim = c(171, 191))  + theme( axis.text.x = element_text(size = 10, color = "black", angle =90, hjust= 0))
zoom_MF <- Plot_MF +coord_cartesian(xlim = c(171, 191))  + theme( axis.text.x = element_text(size = 10, color = "black", angle =90, hjust= 0))

#Plot Germination Rate
GR_MF
GR_FSC

#Plot Germination Rate zoom
zoom_FSC_GR <- GR_FSC +coord_cartesian(xlim = c(171, 191)) 
zoom_MF_GR <- GR_MF +coord_cartesian(xlim = c(171, 191)) 

#Plot figure S1
Figure_full <- plot_grid(
  Plot_FSC_Legend,
  Plot_MF_Legend,
  GR_FSC,
  GR_MF,
  ncol = 2,
  nrow=2,
  rel_heights = c(1,0.6),
  rel_widths = c(1,1),
    align = "hv", 
 axis = "tblr"
)
Figure_full

#Plot Zoom
Figure_zoom <- plot_grid(
  zoom_MF,
  zoom_FSC,
  zoom_MF_GR,
  zoom_FSC_GR,
  ncol = 2,
  nrow=2,
  rel_heights = c(1,0.6),
  rel_widths = c(1,1),
    align = "hv", 
 axis = "tblr"
)
Figure_zoom
```

---
END
---