---
title: "Decoy sugars hijack fungal carbon sensing and prevent plant disease"
author: "Guillaume Chesneau"
date: "26-05-2026"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "") 
```

# Before you start

This script has been made on R 4.0.0 (and Rstudio 1.2.5019)

```{r R version}
R.version
```
```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}
#Load libraries
library(markdown)
library(ggthemes)
library(ggpubr) 
library(rstatix)
library(dunn.test)
library(ggtext)
library(tidyr)
library(markdown)
library(cowplot)
library(ggtree)
library(ape)
library(ggplot2)
library(ggnewscale)
library(dplyr)
library(tidytree)
library(readxl)
library(phytools)
library(tidyverse)

# Prepare base theme
base_theme <- theme_bw() +
  theme(
    strip.background = element_rect(color = "white", fill = "white"),
    strip.text.x = element_markdown(size = 10, color = "black", hjust = 0.5, vjust = 0.5),
    legend.title = element_text(size = 11),
    axis.title.y = element_text(size =11),
    legend.text = element_text(size = 10),
    axis.text.x = element_text(size = 10, color = "black", angle = 90, hjust = 1),
    axis.text.y = element_text(size = 10, color = "black"),
    panel.border = element_rect(color = "black", fill = NA, size = 0.75))

#Colors
color_palette <- c("Water" = "#949494", "3OMG" = "#7FC29B", "Glucose" = "#817E9F", "lightgray" = "lightgray")

```


# A] Bacteria screening

## 1)  Pre-processing

```{r}
#Load dataframe
df <- read.csv(file = "df.csv",header = TRUE, row.names = c("X"), sep=",")
df

#Check water controls sterility
water_df <- df %>% 
  filter(Strain == "Water")

water_check <- water_df %>%
  select(Plate, Well, Media,Time, OD600) %>%
  distinct() %>%   # ensure no duplicates
  pivot_wider(
    names_from = Time,
    values_from = OD600
  ) %>%
  mutate(delta_OD = T120 - T0)

#Keep only T120 or T1W
df120_1 <- subset(df, Time %in% c("T0", "T120"))
df1W_1 <- subset(df, Time %in% c("T0", "T1W"))

#Normalize OD by inoc D0
df120_1 <- df120_1 %>%
  group_by(Plate, Well) %>%
  mutate(norm_OD = OD600[Time == "T120"] - OD600[Time == "T0"]) %>%
  ungroup()

df1W_1 <- df1W_1 %>%
  group_by(Plate, Well) %>%
  mutate(norm_OD = OD600[Time == "T1W"] - OD600[Time == "T0"]) %>%
  ungroup()
```

## 2)  Phylogenetic tree

### a- Prepare tree

```{r}
#Open tree
tree <- read.newick("treeNJ.16S.newick")

#Write tree as nexus
write.nexus(tree, file = "treeNJ.16S.nexus")

# Read the tree (NEXUS may contain multiple trees)
tree_file <- "treeNJ.16S.nexus"
stopifnot(file.exists(tree_file))

# Read the tree (NEXUS may contain multiple trees)
phy <- ape::read.nexus(tree_file)
if (inherits(phy, "multiPhylo")) {
  message("NEXUS contained ", length(phy), " trees; using the first.")
  phy <- phy[[1]]
}


#most basic tree visualization 
tree_plot <- ggplot(phy, aes(x, y)) + geom_tree() + theme_tree()
tree_plot

#Get tips order
tip_order <- phy$tip.label
tip_order

#Add manually R78 and R61
tip_order

#Add water 
tip_order <- c(tip_order, "Water")
```

### b- Trimmed tree

```{r}
#Keep only strain I've screend
to_keep <- unique(df$Strain2)

# Trim the tree to keep only the tips in 'to_keep'
phy_trimmed <- drop.tip(phy, setdiff(phy$tip.label, to_keep))

#Root tree midpoint
#phy_trimmed <- midpoint.root(phy_trimmed)
phy_trimmed <- root(phy_trimmed, outgroup = "Root420", resolve.root = TRUE)

#most basic tree visualization 
tree_plot <- ggplot(phy_trimmed, aes(x, y)) + geom_tree() + theme_tree()
tree_plot

# Your tree
phylo_tree <- phy_trimmed

# Your metadata
metadata <- df
  # Keep only the first row for each unique label
metadata_unique <- metadata %>%
  distinct(Strain2, .keep_all = TRUE)

#Remove Water 
metadata_unique <- subset(metadata_unique, !Strain %in% c( "Water"))

# Make sure metadata has a 'label' column that matches tip labels
metadata_unique <- metadata_unique %>%
  rename(label = Strain2)  # or whatever column matches tip labels in your tree

metadata_unique <- data.frame(metadata_unique, row.names = 7)
metadata_unique$label <- row.names(metadata_unique)
# Check that all tip labels are in the metadata_unique
all(phylo_tree$tip.label %in% row.names(metadata_unique))
# If FALSE, some labels are missing → must fix names

#Merge tree
x <- full_join(as_tibble(phylo_tree), metadata_unique, by = "label") 
tree2 <- as.treedata(x)

# Define phyla colors
col_phyla <- c(
  "Actinobacteria" = "#cc505a",
  "Alphaproteobacteria" = "#84ba66",
  "Bacteroidetes" = "#5d7db9",
  "Betaproteobacteria" = "#adca5c",
  "Firmicutes" = "#e19944",
  "Gammaproteobacteria" = "#5f9353",
  "#N/A" = "gray"
)


# rectangular tree
tree <- ggtree(tree2, layout = "rectangular", branch.length = "none", size = 0.15) +
  geom_tiplab(aes(label = label), size = 3, offset = 0.01) +
  geom_tippoint(aes(color = Phylum_alternative)) +
  scale_color_manual(values = col_phyla)+
  theme(legend.position = "left")  # move legend to the left
tree

tree_V <- ggtree(tree2, layout = "rectangular", branch.length = "none", size = 0.15) +
  geom_tiplab(aes(label = label), size = 3, offset = 0.01, angle = 90, hjust = 1) +
  geom_tippoint(aes(color = Phylum_alternative)) +
  scale_color_manual(values = col_phyla) +
  theme(legend.position = "bottom") +
  coord_flip()+
  scale_x_reverse()
tree_V

#Extract tps order
# Get the tree data
tree_data <- tree$data

# Filter only tips
tip_order <- tree_data %>%
  filter(isTip) %>%
  arrange(y) %>%        # 'y' is the vertical position in rectangular layout
  pull(label)

tip_order

#Add water 
tip_order <- c(tip_order, "Water")
```

## 3)  BarPlot T120

```{r}

#Keep just TimePoint T120
df_T120 <- subset(df120_1, Time %in% c("T120"))

#Make a summary
df_summary <- df_T120 %>%
  group_by(Strain2, Media) %>%
  summarise(
    median_OD = median(norm_OD, na.rm = TRUE),
    sd_OD = sd(norm_OD, na.rm = TRUE),       # standard deviation for error bar
    .groups = "drop"
  )

# Run Wilcoxon test for each Media
results <- df_T120 %>%
  group_by(Media, Strain2) %>%
  summarise(values = list(norm_OD), .groups = "drop") %>%
  group_by(Media) %>%
  filter(Strain2 != "Water") %>%
  mutate(
    p_value_vs_Water = sapply(values, function(x) {
      control <- df_T120 %>% filter(Media == first(Media), Strain2 == "Water") %>% pull(norm_OD)
      wilcox.test(x, control)$p.value
    }),
    p_adj_vs_Water = p.adjust(p_value_vs_Water, method = "BH")  # Benjamini-Hochberg adjustment
  ) %>%
  select(Media, Strain2, p_value_vs_Water, p_adj_vs_Water)

print(results)

#Calculate water median
median_water <- df_T120 %>%
  filter(Strain2 == "Water") %>%          # keep only Water
  group_by(Media) %>%                 # compute per strain
  summarise(median_norm_OD = median(norm_OD, na.rm = TRUE))
median_water

# Add a fill column based on whether each value is below Water median
df_summary$fill_color <- mapply(function(value, media) {
  water_med <- median_water$median_norm_OD[median_water$Media == media]
  if (value <= water_med) {
    "lightgray"             # below Water median → white
  } else {
    media               # otherwise use Media
  }
}, df_summary$median_OD, df_summary$Media)

#Add column based on significance also 
df_summary <- df_summary %>%
  left_join(results %>% select(Media, Strain2, p_adj_vs_Water), 
            by = c("Media", "Strain2"))

# Add fill_color column
df_summary$fill_color <- mapply(function(value, media, p_val) {
  water_med <- median_water$median_norm_OD[median_water$Media == media]
  if (!is.na(p_val) && p_val > 0.05) {
    "lightgray"               # Not significant → gray
  } else if (value <= water_med) {
    "lightgray"               # Below Water median → gray
  } else {
    media                     # Otherwise use Media color
  }
}, df_summary$median_OD, df_summary$Media, df_summary$p_adj_vs_Water)


# Keep only significant strains (not lightgray) and relevant Media

test <- subset(df_summary, !fill_color %in% c("lightgray"))

sig_medians <- df_summary %>%
  filter(fill_color != "lightgray", Media %in% c("3OMG", "Glucose")) %>%
  group_by(Media) %>%
  summarise(sig_median = median(median_OD, na.rm = TRUE)) %>%
  ungroup()


#Order
df_summary$Strain2 <- factor(df_summary$Strain2, levels = c (tip_order))

#Plot
Bar_plot <- ggplot(df_summary, aes(y = Strain2, x = log(median_OD + 1), fill = fill_color)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7) +
  geom_errorbarh(aes(xmin = log(median_OD - sd_OD + 1),
                      xmax = log(median_OD + sd_OD + 1)),
                 position = position_dodge(width = 0.8),
                 height = 0.2) +
  geom_vline(data = median_water,
             aes(xintercept = log(median_norm_OD + 1)),
             color = "black",
             linetype = "dashed",
             linewidth = 0.2) +
  facet_grid(~ Media, scales = "fixed") +
  scale_fill_manual(values = color_palette) +
  base_theme +
  theme(axis.text.y = element_text(),
        plot.title = element_text(hjust = 0.5)) +
  labs(title = "",
       x = "Bacteria growth (log(Median OD600 + 1))",
       y = "Strain")
Bar_plot

# Vertical bar plot
Bar_plot_V <- ggplot(df_summary, aes(
  x = Strain2,
  y = log(median_OD + 1),
  fill = fill_color)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7) +
  geom_errorbar(aes(ymin = log(median_OD - sd_OD + 1),
        ymax = log(median_OD + sd_OD + 1)),
    position = position_dodge(width = 0.8),
    width = 0.2) +
  geom_hline(
    data = median_water,
    aes(yintercept = log(median_norm_OD + 1)),
    color = "black",
    linetype = "dashed",
    linewidth = 0.2) +
   geom_hline(
  data = sig_medians,
  aes(yintercept = log(sig_median + 1), color = Media),
  linetype = "dotted",
  linewidth = 0.7,
  inherit.aes = FALSE
) +
   # Add text for median values just above the line (since y-axis is reversed)
  geom_text(data = sig_medians,
    aes(x = 18,  # place at left of panel; adjust if needed
      y = log(sig_median + 1),
      label = paste0("median = ", round(sig_median, 3))
    ),
    inherit.aes = FALSE,
    hjust = 0,  # left-align
    color = "black",
    size = 3
  ) +
  facet_grid(Media ~ ., scales = "fixed") +
  scale_fill_manual(values = color_palette) +
  scale_color_manual(values = color_palette) +
   base_theme +
  scale_x_discrete(position = "top") +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 1, hjust = 0),
    plot.title = element_text(hjust = 0.5),
    legend.position = "bottom"
  ) +
  labs(
    title = "",
    y = "Bacteria growth (log(Median OD600 + 1))",
    x = ""
  ) +
  scale_y_reverse() 

Bar_plot_V
```

## 3)  BarPlot T1W

```{r}

#Keep just TimePoint T120
df_T1W <- subset(df1W_1, Time %in% c("T1W"))

#Make a summary
df_summary <- df_T1W %>%
  group_by(Strain2, Media) %>%
  summarise(
    median_OD = median(norm_OD, na.rm = TRUE),
    sd_OD = sd(norm_OD, na.rm = TRUE),       # standard deviation for error bar
    .groups = "drop"
  )

# Run Wilcoxon test for each Media
results <- df_T1W %>%
  group_by(Media, Strain2) %>%
  summarise(values = list(norm_OD), .groups = "drop") %>%
  group_by(Media) %>%
  filter(Strain2 != "Water") %>%
  mutate(
    p_value_vs_Water = sapply(values, function(x) {
      control <- df_T1W %>% filter(Media == first(Media), Strain2 == "Water") %>% pull(norm_OD)
      wilcox.test(x, control)$p.value
    }),
    p_adj_vs_Water = p.adjust(p_value_vs_Water, method = "BH")  # Benjamini-Hochberg adjustment
  ) %>%
  select(Media, Strain2, p_value_vs_Water, p_adj_vs_Water)

print(results)

#Calculate water median
median_water <- df_T1W %>%
  filter(Strain2 == "Water") %>%          # keep only Water
  group_by(Media) %>%                 # compute per strain
  summarise(median_norm_OD = median(norm_OD, na.rm = TRUE))
median_water

# Add a fill column based on whether each value is below Water median
df_summary$fill_color <- mapply(function(value, media) {
  water_med <- median_water$median_norm_OD[median_water$Media == media]
  if (value <= water_med) {
    "lightgray"             # below Water median → white
  } else {
    media               # otherwise use Media
  }
}, df_summary$median_OD, df_summary$Media)

#Add column based on significance also 
df_summary <- df_summary %>%
  left_join(results %>% select(Media, Strain2, p_adj_vs_Water), 
            by = c("Media", "Strain2"))

# Add fill_color column
df_summary$fill_color <- mapply(function(value, media, p_val) {
  water_med <- median_water$median_norm_OD[median_water$Media == media]
  if (!is.na(p_val) && p_val > 0.05) {
    "lightgray"               # Not significant → gray
  } else if (value <= water_med) {
    "lightgray"               # Below Water median → gray
  } else {
    media                     # Otherwise use Media color
  }
}, df_summary$median_OD, df_summary$Media, df_summary$p_adj_vs_Water)


# Keep only significant strains (not lightgray) and relevant Media

test <- subset(df_summary, !fill_color %in% c("lightgray"))

sig_medians <- df_summary %>%
  filter(fill_color != "lightgray", Media %in% c("3OMG", "Glucose")) %>%
  group_by(Media) %>%
  summarise(sig_median = median(median_OD, na.rm = TRUE)) %>%
  ungroup()


#Order
df_summary$Strain2 <- factor(df_summary$Strain2, levels = c (tip_order))

#Plot
Bar_plot_1W <- ggplot(df_summary, aes(y = Strain2, x = log(median_OD + 1), fill = fill_color)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7) +
  geom_errorbarh(aes(xmin = log(median_OD - sd_OD + 1),
                      xmax = log(median_OD + sd_OD + 1)),
                 position = position_dodge(width = 0.8),
                 height = 0.2) +
  geom_vline(data = median_water,
             aes(xintercept = log(median_norm_OD + 1)),
             color = "black",
             linetype = "dashed",
             linewidth = 0.2) +
  facet_grid(~ Media, scales = "fixed") +
  scale_fill_manual(values = color_palette) +
  base_theme +
  theme(axis.text.y = element_text(),
        plot.title = element_text(hjust = 0.5)) +
  labs(title = "",
       x = "Bacteria growth (log(Median OD600 + 1))",
       y = "Strain")
Bar_plot_1W

# Vertical bar plot
Bar_plot_1W_V <- ggplot(df_summary, aes(
  x = Strain2,
  y = log(median_OD + 1),
  fill = fill_color)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7) +
  geom_errorbar(aes(ymin = log(median_OD - sd_OD + 1),
        ymax = log(median_OD + sd_OD + 1)),
    position = position_dodge(width = 0.8),
    width = 0.2) +
  geom_hline(
    data = median_water,
    aes(yintercept = log(median_norm_OD + 1)),
    color = "black",
    linetype = "dashed",
    linewidth = 0.2) +
   geom_hline(
  data = sig_medians,
  aes(yintercept = log(sig_median + 1), color = Media),
  linetype = "dotted",
  linewidth = 0.7,
  inherit.aes = FALSE
) +
   # Add text for median values just above the line (since y-axis is reversed)
  geom_text(data = sig_medians,
    aes(x = 18,  # place at left of panel; adjust if needed
      y = log(sig_median + 1),
      label = paste0("median = ", round(sig_median, 3))
    ),
    inherit.aes = FALSE,
    hjust = 0,  # left-align
    color = "black",
    size = 3
  ) +
  facet_grid(Media ~ ., scales = "fixed") +
  scale_fill_manual(values = color_palette) +
  scale_color_manual(values = color_palette) +
   base_theme +
  scale_x_discrete(position = "top") +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 1, hjust = 0),
    plot.title = element_text(hjust = 0.5),
    legend.position = "bottom"
  ) +
  labs(
    title = "",
    y = "Bacteria growth (log(Median OD600 + 1))",
    x = ""
  ) +
  ylim(-0.01,0.205) #I lose one outlier bottom, but better display

Bar_plot_1W_V
```

## 5)   Heat Map plot

```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}
glucose_palette <- c(
  "#c9c7d9",
  "#9a96b6",
  "#817E9F",
  "#4a4866"
)

omg_palette <- c(
  "#cfe8dc",
  "#a6d3be",
  "#7FC29B",
  "#3f7f61"
)

df_summary$fill_value <- mapply(function(value, media, p_val) {
  water_med <- median_water$median_norm_OD[median_water$Media == media]

  if (!is.na(p_val) && p_val > 0.05) {
    NA
  } else if (value <= water_med) {
    NA
  } else {
    log(value + 1)
  }
}, df_summary$median_OD, df_summary$Media, df_summary$p_adj_vs_Water)



#Order
df_summary$Strain2 <- factor(df_summary$Strain2, levels = c (tip_order))

plot <- ggplot() +

  geom_tile(
    data = df_summary[df_summary$Media == "Glucose", ],
    aes(x = Strain2, y = 1, fill = fill_value),
    color = "white",
    linewidth = 0.2
  ) +
  scale_fill_gradientn(
    colors = glucose_palette,
    na.value = "white",
    name = "Glucose\nlog(Median OD600 + 1)",
    limits = range(df_summary$fill_value[df_summary$Media == "Glucose"], na.rm = TRUE)
  ) +

  ggnewscale::new_scale_fill() +

  geom_tile(
    data = df_summary[df_summary$Media == "3OMG", ],
    aes(x = Strain2, y = 1, fill = fill_value),
    color = "white",
    linewidth = 0.2
  ) +
  scale_fill_gradientn(
    colors = omg_palette,
    na.value = "white",
    name = "3-OMG\nlog(Median OD600 + 1)",
    limits = range(df_summary$fill_value[df_summary$Media == "3OMG"], na.rm = TRUE)
  ) +

  facet_grid(Media ~ .) +
  scale_x_discrete(position = "top") +
  base_theme +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 1, hjust = 0),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom"
  ) +
  labs(x = "", y = "")
plot
```


## 4)   All plots Plant

```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}
Plot_Comb1 <- plot_grid(tree_V, Bar_plot_V, rel_heights = c(1,1), rel_widths = c(1,1), nrow=2,   align="hv",  axis = "tb")
Plot_Comb1

Plot_Comb2 <- plot_grid(tree_V, Bar_plot_1W_V, rel_heights = c(1,1), rel_widths = c(1,1), nrow=2,   align="hv",  axis = "tb")
Plot_Comb2
```




# B] Fungal screening

## 1)  Pre-processing

```{r}
#Load dataframe
df <- read.csv(file = "all_data_3OMG.csv",header = TRUE, row.names = c("X"), sep=",")
df

#Normalize OD by water control
df2 <- df %>%
  group_by(Strain2) %>%
  mutate(
    max_water = mean(OD600[Media == "Water"], na.rm = TRUE),
    OD600_corrected = OD600 - max_water
  )
```

## 2)  Plot

```{r}
#Summarize for bars
df_summary <- df2 %>%
  group_by(Strain2, Media, Host) %>%
  summarise(
    median_OD = median(OD600_corrected, na.rm = TRUE),
    sd_OD = sd(OD600_corrected, na.rm = TRUE),
    .groups = "drop"
  )

# Vérifier la normalité par Time × Media
normality_results <- df %>%
  group_by(Strain2) %>%
  shapiro_test(OD600) 
normality_results # Not normal for more than half --> non-parametric

#Stats
wilcox_results <- df %>%
  filter(Media %in% c("Water", "3OMG", "Glucose")) %>%
  group_by(Strain2) %>%
  summarise(
    # Extract vectors
    water_vals   = list(OD600[Media == "Water"]),
    omg_vals     = list(OD600[Media == "3OMG"]),
    glucose_vals = list(OD600[Media == "Glucose"]),
    
    # Do tests only if both groups exist
    p_Water_vs_3OMG = if(length(unlist(water_vals)) > 0 && length(unlist(omg_vals)) > 0)
                        wilcox.test(unlist(water_vals), unlist(omg_vals))$p.value else NA,

    p_Water_vs_Glucose = if(length(unlist(water_vals)) > 0 && length(unlist(glucose_vals)) > 0)
                           wilcox.test(unlist(water_vals), unlist(glucose_vals))$p.value else NA
  ) %>%
  ungroup() %>%
  mutate(
    p_adj_Water_vs_3OMG    = p.adjust(p_Water_vs_3OMG, method = "BH"),
    p_adj_Water_vs_Glucose = p.adjust(p_Water_vs_Glucose, method = "BH")
  )


# Assign strain-level color based on significance
wilcox_expanded <- wilcox_results %>%
  tidyr::pivot_longer(
    cols = c(p_Water_vs_3OMG, p_Water_vs_Glucose),
    names_to = "Media",
    values_to = "p_value"
  ) %>%
  mutate(
    Media = dplyr::recode(Media,
                          "p_Water_vs_3OMG"    = "3OMG",
                          "p_Water_vs_Glucose" = "Glucose")
  ) %>%
  mutate(
    color_by_sig = case_when(
      !is.na(p_value) & p_value <= 0.05 & Media == "3OMG"    ~ "3OMG",
      !is.na(p_value) & p_value <= 0.05 & Media == "Glucose" ~ "Glucose",
      TRUE ~ "lightgray"
    )
  )

#Join colors to orignial table
df_summary <- df_summary %>%
  left_join(wilcox_expanded %>% dplyr::select(Strain2, Media, color_by_sig),
            by = c("Strain2", "Media")) %>%
  mutate(
    color_by_sig = case_when(
      median_OD < 0 ~ "lightgray",  # NEW RULE
      TRUE ~ color_by_sig           # fall back to original color
    )
  )

#Calculate overall median value
global_water_median <- df_summary%>%
  filter(Media == "Water") %>%
  summarise(global_med = median(median_OD, na.rm = TRUE)) %>%
  pull(global_med)

global_water_median

#Remove water
df_summary_f <- subset(df_summary, !Media %in% c("Water"))

#BarPLot
Bar_plot <- ggplot(df_summary_f, aes(y = Strain2, x = log(median_OD + 1), fill = color_by_sig)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7) +
  geom_errorbarh(
    aes(xmin = log(median_OD - sd_OD + 1),
        xmax = log(median_OD + sd_OD + 1)),
    height = 0.2,
    position = position_dodge(width = 0.8)
  ) +
  geom_vline(
  xintercept = log(0 + 1),
  linetype = "dashed",
  color = "black",
  linewidth = 0.4
) +
  facet_grid(Host ~ Media, scales = "free_y") +
  scale_fill_manual(values = color_palette) +
  theme_bw() +
  theme(
    axis.text.y = element_text(),
    plot.title = element_text(hjust = 0.5)
  ) +
  labs(
    title = "Median Corrected OD600 by Strain × Media",
    x = " Fungal growth (log(Median Corrected-OD600 + 1)",
    y = "Strain"
  ) +
    scale_fill_manual(values = color_palette) +
   base_theme

Bar_plot
```

## 3)  Compare growth capacity between isolates Glucose / 3OMG

```{r}
#Subset dataset 
df_Growth <- subset(df_summary_f, !color_by_sig %in% c("lightgray"))

#Calcualte median per Treatment and Host
median_summary <- df_Growth %>%
  group_by(Host, Media) %>%
  summarise(
    median_median_OD = median(median_OD),
    n = n(),
    .groups = "drop"
  )

median_summary

#Stats Glucose VS 3OMG per Treatment
wilcox_results <- df_Growth %>%
  group_by(Host) %>%
  filter(n_distinct(Media) == 2) %>%   # only Hosts with both media
  group_modify(~{
    test <- wilcox.test(median_OD ~ Media, data = .x, exact = TRUE)
    broom::tidy(test)
  }) %>%
  ungroup()

print(wilcox_results)
```

## 4)  Plot Arabidopsis thaliana

```{r}
#Load raw data
df <- read.csv(file = "all_data_3OMG.csv",header = TRUE, row.names = c("X"), sep=",")
df

#Keep only Arabdiopsis
df <- subset(df, Host %in% c("A_thaliana"))
df_summary_f_At <- subset(df_summary_f, Host %in% c("A_thaliana"))
```


### a- Phylogenetic tree

#### i- Prepare tree

```{r}
#Open tree
tree_ITS <- read.newick("tree41_goodnames.nwk")

#Write tree as nexus
write.nexus(tree_ITS, file = "tree41_goodnames.nwk.nexus")

# Read the tree (NEXUS may contain multiple trees)
tree_file_ITS <- "tree41_goodnames.nwk.nexus"
stopifnot(file.exists(tree_file_ITS))

# Read the tree (NEXUS may contain multiple trees)
phy <- ape::read.nexus(tree_file_ITS)
if (inherits(phy, "multiPhylo")) {
  message("NEXUS contained ", length(phy), " trees; using the first.")
  phy <- phy[[1]]
}


#most basic tree visualization 
tree_plot <- ggplot(phy, aes(x, y)) + geom_tree() + theme_tree()
tree_plot

#Get tips order
tip_order <- phy$tip.label
tip_order

#Add manually R78 and R61
tip_order

#Add water 
tip_order <- c(tip_order, "Water")
```


#### ii-  Trimmed tree At

```{r}
#Keep only strain I've screend
to_keep <- unique(df_summary_f_At$Strain2)

# Trim the tree to keep only the tips in 'to_keep'
phy_trimmed <- drop.tip(phy, setdiff(phy$tip.label, to_keep))

#Root tree midpoint
#phy_trimmed <- midpoint.root(phy_trimmed)
phy_trimmed <- root(phy_trimmed, outgroup = "Mortierella_elongata_MPI-CAGE-AA-0104", resolve.root = TRUE)

#most basic tree visualization 
tree_plot <- ggplot(phy_trimmed, aes(x, y)) + geom_tree() + theme_tree()
tree_plot

# Your tree
phylo_tree <- phy_trimmed

# Your metadata
metadata <- df
  # Keep only the first row for each unique label
metadata_unique <- metadata %>%
  distinct(Strain2, .keep_all = TRUE)


# Make sure metadata has a 'label' column that matches tip labels
metadata_unique <- metadata_unique %>%
  rename(label = Strain2)  # or whatever column matches tip labels in your tree

metadata_unique <- data.frame(metadata_unique, row.names = 2)
metadata_unique$label <- row.names(metadata_unique)
# Check that all tip labels are in the metadata_unique
all(phylo_tree$tip.label %in% row.names(metadata_unique))
# If FALSE, some labels are missing → must fix names

#Merge tree
x <- full_join(as_tibble(phylo_tree), metadata_unique, by = "label") 
tree2 <- as.treedata(x)

# Define phyla colors
col_class <- c(
  "Mortierellomycetes" = "#2166AC",
  "Sordariomycetes" = "#EF8A62",
  "Dothideomycetes" = "#FDDBC7",
  "Leotiomycetes" = "#67A9CF",
  "#N/A" = "gray"
)

# rectangular tree
tree <- ggtree(tree2, layout = "rectangular", branch.length = "none", size = 0.15) +
  geom_tiplab(aes(label = label), size = 3, offset = 0.01) +
  geom_tippoint(aes(color = Class)) +
  scale_color_manual(values = col_class)+
  theme(legend.position = "left")  # move legend to the left
tree

tree_V <- ggtree(tree2, layout = "rectangular",branch.length = "none", size = 0.15) +
  geom_tiplab(aes(label = label), size = 3, offset = 0.01, angle = 90, hjust = 1) +
  geom_tippoint(aes(color = Class)) +
  scale_color_manual(values = col_class) +
  theme(legend.position = "right") +
  coord_flip()+
  scale_x_reverse()+
    theme(legend.position = "right") 
tree_V


#Extract tps order
# Get the tree data
tree_data <- tree$data

# Filter only tips
tip_order <- tree_data %>%
  filter(isTip) %>%
  arrange(y) %>%        # 'y' is the vertical position in rectangular layout
  pull(label)

tip_order

#Add water 
tip_order <- c(tip_order, "Water")
```

### b- BarPlot

```{r}
# Keep only significant strains (not lightgray) and relevant Media
sig_medians <- df_summary_f_At %>%
  filter(color_by_sig != "lightgray", Media %in% c("3OMG", "Glucose")) %>%
  group_by(Media) %>%
  summarise(sig_median = median(median_OD, na.rm = TRUE)) %>%
  ungroup()

# Add a small offset to place text just below the line
offset <- 0.1  # adjust as needed

#Order
df_summary_f_At$Strain2 <- factor(df_summary_f_At$Strain2, levels = c (tip_order))

#BarPLot
# Vertical bar plot 
Bar_plot_At <- ggplot(df_summary_f_At, aes(
  x = Strain2,
  y = log(median_OD + 1),
  fill = color_by_sig
)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7) +
  geom_errorbar(
    aes(ymin = log(median_OD - sd_OD + 1),
        ymax = log(median_OD + sd_OD + 1)),
    width = 0.2,
    position = position_dodge(width = 0.8)
  ) +
  geom_hline(
    yintercept = log(0 + 1),
    linetype = "dashed",
    color = "black",
    linewidth = 0.4
  ) +
   geom_hline(
    data = sig_medians,
    aes(yintercept = log(sig_median + 1), color = Media),
    linetype = "dotted",
    linewidth = 0.7
  ) +
   # Add text for median values just above the line (since y-axis is reversed)
  geom_text(
    data = sig_medians,
    aes(x = 18,  # place at left of panel; adjust if needed
      y = log(sig_median + 1) + offset,
      label = paste0("median = ", round(sig_median, 2))
    ),
    inherit.aes = FALSE,
    hjust = 0,  # left-align
    color = "black",
    size = 3
  ) +
  facet_grid(Media  ~ ., scales = "free_x") +
  scale_fill_manual(values = color_palette) +
  scale_color_manual(values = color_palette) +
  base_theme +
  scale_x_discrete(position = "top") +
  theme(
    axis.text.x = element_blank(),
    plot.title = element_text(hjust = 0.5)
  ) +
  labs(
    title = "",
    y = "Fungal growth (log(Median OD600 + 1))",
    x = ""
  ) +
  scale_y_reverse() +
    theme(legend.position = "bottom") 

Bar_plot_At
```

### c- HeatMap

```{r}
# Define palettes
glucose_palette <- c("#c9c7d9", "#9a96b6", "#817E9F", "#4a4866")
omg_palette <- c("#cfe8dc", "#a6d3be", "#7FC29B", "#3f7f61")
  
# Compute fill_value: log + 1, NA if non-significant
df_summary_f_At$fill_value <- mapply(function(value, media, sig) {
  if (sig == "lightgray") {
    NA
  } else {
    log(value + 1)
  }
}, df_summary_f_At$median_OD, df_summary_f_At$Media, df_summary_f_At$color_by_sig)

#add better names for strains
df_summary_f_At <- df_summary_f_At %>%
  mutate(
    Strain2_F = paste0("F", str_extract(Strain2, "\\d{3}$"))
  )

# Order strains
#add better names for strains
tip_order_F <- paste0("F", str_extract(tip_order, "\\d{3}$"))
df_summary_f_At$Strain2_F <- factor(df_summary_f_At$Strain2_F, levels = tip_order_F)


# Compute fill_value for heatmap
df_summary_f_At$fill_value <- mapply(function(value, sig) {
  if (sig == "lightgray") NA else log(value + 1)
}, df_summary_f_At$median_OD, df_summary_f_At$color_by_sig)

# Define fixed scale limits for both media (optional, makes them comparable)
common_limits <- range(df_summary_f_At$fill_value, na.rm = TRUE)

# Heatmap plot
heatmap_plot <- ggplot() +

  # Glucose tiles
  geom_tile(
    data = df_summary_f_At %>% filter(Media == "Glucose"),
    aes(x = Strain2_F, y = 1, fill = fill_value),
    color = "white", linewidth = 0.2
  ) +
  scale_fill_gradientn(
    colors = glucose_palette,
    na.value = "white",
    name = "Glucose\nlog(Median OD600 + 1)",
    limits = common_limits
  ) +

  ggnewscale::new_scale_fill() +

  # 3OMG tiles
  geom_tile(
    data = df_summary_f_At %>% filter(Media == "3OMG"),
    aes(x = Strain2_F, y = 1, fill = fill_value),
    color = "white", linewidth = 0.2
  ) +
  scale_fill_gradientn(
    colors = omg_palette,
    na.value = "white",
    name = "3-OMG\nlog(Median OD600 + 1)",
    limits = common_limits
  ) +

  facet_grid(Media ~ .) +
  scale_x_discrete(position = "top") +
  base_theme +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 1, hjust = 0),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom"
  ) +
  labs(x = "", y = "")
heatmap_plot
```

## 5)  Plot Brassica Napus

OD is normalized by T0 --> (OD600.T120) - (OD600 at.T0).    
OD is also normalized by water --> OD600.Fungi - OD600 at.water

```{r}
#Load raw data
df <- read.csv(file = "all_data_3OMG.csv",header = TRUE, row.names = c("X"), sep=",")
df

#Keep only Arabdiopsis
df <- subset(df, Host %in% c("B_napus"))
df_summary_f_Bn <- subset(df_summary_f, Host %in% c("B_napus"))
```

### a- Phylogenetic tree

#### i- Prepare tree

```{r}
#Open tree
tree_ITS <- read.newick("treeNJ.ITS.BN.newick")

#Write tree as nexus
write.nexus(tree_ITS, file = "treeNJ.ITS.BN.newick.nexus")

# Read the tree (NEXUS may contain multiple trees)
tree_file_ITS <- "treeNJ.ITS.BN.newick.nexus"
stopifnot(file.exists(tree_file_ITS))

# Read the tree (NEXUS may contain multiple trees)
phy <- ape::read.nexus(tree_file_ITS)
if (inherits(phy, "multiPhylo")) {
  message("NEXUS contained ", length(phy), " trees; using the first.")
  phy <- phy[[1]]
}


#most basic tree visualization 
tree_plot <- ggplot(phy, aes(x, y)) + geom_tree() + theme_tree()
tree_plot

#Get tips order
tip_order <- phy$tip.label
tip_order

#Add manually R78 and R61
tip_order

#Add water 
tip_order <- c(tip_order, "Water")
```

#### ii-  Trimmed tree At

```{r}
#Keep only strain I've screend
to_keep <- unique(df_summary_f_Bn$Strain2)

# Trim the tree to keep only the tips in 'to_keep'
phy_trimmed <- drop.tip(phy, setdiff(phy$tip.label, to_keep))

#Root tree midpoint
phy_trimmed <- midpoint.root(phy_trimmed)
#phy_trimmed <- root(phy_trimmed, outgroup = "Mortierella_elongata_MPI-CAGE-AA-0104", resolve.root = TRUE)

#most basic tree visualization 
tree_plot <- ggplot(phy_trimmed, aes(x, y)) + geom_tree() + theme_tree()
tree_plot

# Your tree
phylo_tree <- phy_trimmed

# Your metadata
metadata <- df
  # Keep only the first row for each unique label
metadata_unique <- metadata %>%
  distinct(Strain2, .keep_all = TRUE)

# Make sure metadata has a 'label' column that matches tip labels
#metadata_unique <- metadata_unique %>%
#  rename(label = Strain2)  # or whatever column matches tip labels in your tree

metadata_unique <- data.frame(metadata_unique, row.names = 2)
metadata_unique$label <- row.names(metadata_unique)
# Check that all tip labels are in the metadata_unique
all(phylo_tree$tip.label %in% row.names(metadata_unique))
# If FALSE, some labels are missing → must fix names

#Merge tree
x <- full_join(as_tibble(phylo_tree), metadata_unique, by = "label") 
tree2 <- as.treedata(x)

# Define phyla colors
col_class <- c(
  "Mortierellomycetes" = "#2166AC",
  "Sordariomycetes" = "#EF8A62",
  "Dothideomycetes" = "#FDDBC7",
  "Leotiomycetes" = "#67A9CF",
  "Mucoromycetes" = "#1B9E77",
  "Agaricomycetes" = "#A6761D",
  "Eurotiomycetes" = "#FFD92F",
  "#N/A" = "gray"
)


# rectangular tree
tree <- ggtree(tree2, layout = "rectangular", branch.length = "none", size = 0.15) +
  geom_tiplab(aes(label = label), size = 3, offset = 0.01) +
  geom_tippoint(aes(color = Class)) +
  scale_color_manual(values = col_class)+
  theme(legend.position = "left")  # move legend to the left
tree

tree_V <- ggtree(tree2, layout = "rectangular",branch.length = "none", size = 0.15) +
  geom_tiplab(aes(label = label), size = 3, offset = 0.01, angle = 90, hjust = 1) +
  geom_tippoint(aes(color = Class)) +
  scale_color_manual(values = col_class) +
  theme(legend.position = "right") +
  coord_flip()+
  scale_x_reverse()+
    theme(legend.position = "right") 
tree_V


#Extract tps order
# Get the tree data
tree_data <- tree$data

# Filter only tips
tip_order <- tree_data %>%
  filter(isTip) %>%
  arrange(y) %>%        # 'y' is the vertical position in rectangular layout
  pull(label)
tip_order

#Add water 
tip_order <- c(tip_order, "Water")
```
### b- HeatMap

```{r}
# Define palettes
glucose_palette <- c("#c9c7d9", "#9a96b6", "#817E9F", "#4a4866")
omg_palette <- c("#cfe8dc", "#a6d3be", "#7FC29B", "#3f7f61")
  
# Compute fill_value: log + 1, NA if non-significant
df_summary_f_Bn$fill_value <- mapply(function(value, media, sig) {
  if (sig == "lightgray") {
    NA
  } else {
    log(value + 1)
  }
}, df_summary_f_Bn$median_OD, df_summary_f_Bn$Media, df_summary_f_Bn$color_by_sig)

# Order strains
df_unique <- df[!duplicated(df$Strain2), ]
df_summary_f_Bn <- merge(df_summary_f_Bn, df_unique[, c("Strain2", "Class")], 
                   by = "Strain2", 
                   all.x = TRUE)  # keep all rows from df1

df_summary_f_Bn$Class <- factor(df_summary_f_Bn$Class, levels = c("Mortierellomycetes","Dothideomycetes", "Leotiomycetes", "Sordariomycetes", "Mucoromycetes", "Eurotiomycetes", "Agaricomycetes", "Pezizomycotina",  "NA"))

# Compute fill_value for heatmap
df_summary_f_Bn$fill_value <- mapply(function(value, sig) {
  if (sig == "lightgray") NA else log(value + 1)
}, df_summary_f_Bn$median_OD, df_summary_f_Bn$color_by_sig)

# Define fixed scale limits for both media (optional, makes them comparable)
common_limits <- range(df_summary_f_Bn$fill_value, na.rm = TRUE)

#Calculate media growth per Carbon
df_summary_f_Bn_sig <- subset(df_summary_f_Bn, !color_by_sig %in% c("lightgray"))
df_Median <- df_summary_f_Bn_sig %>%
  group_by(Media) %>%
  summarise(median_OD = mean(median_OD, na.rm = TRUE))


# Heatmap plot
heatmap_plot_Bn <- ggplot() +

  # Glucose tiles
  geom_tile(
    data = df_summary_f_Bn %>% filter(Media == "Glucose"),
    aes(x = Strain2, y = 1, fill = fill_value),
    color = "white", linewidth = 0.2
  ) +
  scale_fill_gradientn(
    colors = glucose_palette,
    na.value = "white",
    name = "Glucose\nlog(Median OD600 + 1)",
    limits = common_limits
  ) +
  ggnewscale::new_scale_fill() +
  # 3OMG tiles
  geom_tile(
    data = df_summary_f_Bn %>% filter(Media == "3OMG"),
    aes(x = Strain2, y = 1, fill = fill_value),
    color = "white", linewidth = 0.2
  ) +
  scale_fill_gradientn(
    colors = omg_palette,
    na.value = "white",
    name = "3-OMG\nlog(Median OD600 + 1)",
    limits = common_limits
  ) +

ggh4x::facet_grid2(Media ~ Class , space = "free", scales = "free") +
  scale_x_discrete(position = "top") +
  base_theme +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 1, hjust = 0),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom"
  ) +
  labs(x = "", y = "")
heatmap_plot_Bn
```

## 6)  Plot Triticum aestivum

OD is normalized by T0 --> (OD600.T120) - (OD600 at.T0).    
OD is also normalized by water --> OD600.Fungi - OD600 at.water

```{r}
#Load raw data
df <- read.csv(file = "all_data_3OMG.csv",header = TRUE, row.names = c("X"), sep=",")
df

#Keep only Arabdiopsis
df <- subset(df, Host %in% c("T_aestivum"))
df_summary_f_Ta <- subset(df_summary_f, Host %in% c("T_aestivum"))
```

### a- Phylogenetic tree

#### i- Prepare tree

```{r}
#Open tree
tree_ITS <- read.newick("treeNJ.ITS.TA.newick")

#Write tree as nexus
write.nexus(tree_ITS, file = "treeNJ.ITS.TA.newick.nexus")

# Read the tree (NEXUS may contain multiple trees)
tree_file_ITS <- "treeNJ.ITS.TA.newick.nexus"
stopifnot(file.exists(tree_file_ITS))

# Read the tree (NEXUS may contain multiple trees)
phy <- ape::read.nexus(tree_file_ITS)
if (inherits(phy, "multiPhylo")) {
  message("NEXUS contained ", length(phy), " trees; using the first.")
  phy <- phy[[1]]
}


#most basic tree visualization 
tree_plot <- ggplot(phy, aes(x, y)) + geom_tree() + theme_tree()
tree_plot

#Get tips order
tip_order <- phy$tip.label
tip_order

#Add manually R78 and R61
tip_order

#Add water 
tip_order <- c(tip_order, "Water")
```
#### ii-  Trimmed tree Ta

```{r}
#Keep only strain I've screend
to_keep <- unique(df_summary_f_Ta$Strain2)

# Trim the tree to keep only the tips in 'to_keep'
phy_trimmed <- drop.tip(phy, setdiff(phy$tip.label, to_keep))

#Root tree midpoint
phy_trimmed <- midpoint.root(phy_trimmed)
#phy_trimmed <- root(phy_trimmed, outgroup = "Mortierella_elongata_MPI-CAGE-AA-0104", resolve.root = TRUE)

#most basic tree visualization 
tree_plot <- ggplot(phy_trimmed, aes(x, y)) + geom_tree() + theme_tree()
tree_plot

# Your tree
phylo_tree <- phy_trimmed

# Your metadata
metadata <- df
  # Keep only the first row for each unique label
metadata_unique <- metadata %>%
  distinct(Strain2, .keep_all = TRUE)


# Make sure metadata has a 'label' column that matches tip labels
#metadata_unique <- metadata_unique %>%
#  rename(label = Strain2)  # or whatever column matches tip labels in your tree

metadata_unique <- data.frame(metadata_unique, row.names = 2)
metadata_unique$label <- row.names(metadata_unique)
# Check that all tip labels are in the metadata_unique
all(phylo_tree$tip.label %in% row.names(metadata_unique))
# If FALSE, some labels are missing → must fix names

#Merge tree
x <- full_join(as_tibble(phylo_tree), metadata_unique, by = "label") 
tree2 <- as.treedata(x)

# Define phyla colors
col_class <- c(
  "Mortierellomycetes" = "#2166AC",
  "Sordariomycetes" = "#EF8A62",
  "Dothideomycetes" = "#FDDBC7",
  "Leotiomycetes" = "#67A9CF",
  "Mucoromycetes" = "#1B9E77",
  "Agaricomycetes" = "#A6761D",
  "Eurotiomycetes" = "#FFD92F",
  "#N/A" = "gray"
)

# rectangular tree
tree <- ggtree(tree2, layout = "rectangular", branch.length = "none", size = 0.15) +
  geom_tiplab(aes(label = label), size = 3, offset = 0.01) +
  geom_tippoint(aes(color = Class)) +
  scale_color_manual(values = col_class)+
  theme(legend.position = "left")  # move legend to the left
tree

tree_V <- ggtree(tree2, layout = "rectangular",branch.length = "none", size = 0.15) +
  geom_tiplab(aes(label = label), size = 3, offset = 0.01, angle = 90, hjust = 1) +
  geom_tippoint(aes(color = Class)) +
  scale_color_manual(values = col_class) +
  theme(legend.position = "right") +
  coord_flip()+
  scale_x_reverse()+
    theme(legend.position = "right") 
tree_V

#Extract tps order
# Get the tree data
tree_data <- tree$data

# Filter only tips
tip_order <- tree_data %>%
  filter(isTip) %>%
  arrange(y) %>%        # 'y' is the vertical position in rectangular layout
  pull(label)

tip_order

#Add water 
tip_order <- c(tip_order, "Water")
```

### b- HeatMap

```{r}
# Define palettes
glucose_palette <- c("#c9c7d9", "#9a96b6", "#817E9F", "#4a4866")
omg_palette <- c("#cfe8dc", "#a6d3be", "#7FC29B", "#3f7f61")
  
# Compute fill_value: log + 1, NA if non-significant
df_summary_f_Ta$fill_value <- mapply(function(value, media, sig) {
  if (sig == "lightgray") {
    NA
  } else {
    log(value + 1)
  }
}, df_summary_f_Ta$median_OD, df_summary_f_Ta$Media, df_summary_f_Ta$color_by_sig)

# Order strains
df_unique <- df[!duplicated(df$Strain2), ]
df_summary_f_Ta <- merge(df_summary_f_Ta, df_unique[, c("Strain2", "Class")], 
                   by = "Strain2", 
                   all.x = TRUE)  # keep all rows from df1

df_summary_f_Ta$Class <- factor(df_summary_f_Ta$Class, levels = c("Mortierellomycetes","Dothideomycetes", "Leotiomycetes", "Sordariomycetes", "Mucoromycetes", "Eurotiomycetes", "Agaricomycetes", "Pezizomycotina",  "NA"))


# Compute fill_value for heatmap
df_summary_f_Ta$fill_value <- mapply(function(value, sig) {
  if (sig == "lightgray") NA else log(value + 1)
}, df_summary_f_Ta$median_OD, df_summary_f_Ta$color_by_sig)

# Define fixed scale limits for both media (optional, makes them comparable)
common_limits <- range(df_summary_f_Ta$fill_value, na.rm = TRUE)

#Calculate media growth per Carbon
df_summary_f_Ta_sig <- subset(df_summary_f_Ta, !color_by_sig %in% c("lightgray"))
df_Median <- df_summary_f_Ta_sig %>%
  group_by(Media) %>%
  summarise(median_OD = mean(median_OD, na.rm = TRUE))

# Heatmap plot
heatmap_plot_Ta <- ggplot() +
  # Glucose tiles
  geom_tile(
    data = df_summary_f_Ta %>% filter(Media == "Glucose"),
    aes(x = Strain2, y = 1, fill = fill_value),
    color = "white", linewidth = 0.2
  ) +
  scale_fill_gradientn(
    colors = glucose_palette,
    na.value = "white",
    name = "Glucose\nlog(Median OD600 + 1)",
    limits = common_limits
  ) +
  ggnewscale::new_scale_fill() +
  # 3OMG tiles
  geom_tile(
    data = df_summary_f_Ta %>% filter(Media == "3OMG"),
    aes(x = Strain2, y = 1, fill = fill_value),
    color = "white", linewidth = 0.2
  ) +
  scale_fill_gradientn(
    colors = omg_palette,
    na.value = "white",
    name = "3-OMG\nlog(Median OD600 + 1)",
    limits = common_limits
  ) +
ggh4x::facet_grid2(Media ~ Class , space = "free", scales = "free") +
  scale_x_discrete(position = "top") +
  base_theme +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 1, hjust = 0),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom"
  ) +
  labs(x = "", y = "") 
heatmap_plot_Ta
```

# C] All plots Plant

```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}
Plot_Comb1 <- plot_grid(tree_V, heatmap_plot, rel_heights = c(1,1), rel_widths = c(1,1), nrow=2,   align="hv",  axis = "tb")
Plot_Comb1
```


---
END
---