---
title: "Decoy sugars hijack fungal carbon sensing and prevent plant disease"
author: "Guillaume Chesneau"
date: "26-05-2026"
output: html_document
---

Pre-processing in Galaxy.     
- Quality checked with FastQC, Flatten, MultiFastQG.   
- Mapping using STAR.   
- Strandness checked with Infer Experiment (Reverse stranded library).     
- Feature counts (exons, fragments are counted).   
- DeSeq : All comparions made with Water-MF conditions:     
  - Water-MF_3OMG-MF.  
  - Water-MF_3OMG-F16.  
  - Water-MF_Water-F16.      
  

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "") 
```


```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}
# Load libraries
library(ggplot2)
library(dplyr)
library(tidyr)
library(tibble)
library(stringr)
library(rstatix)
library(ggpubr)
library(dunn.test)
library(ggthemes)
library(ggtext)
library(markdown)
library(DESeq2)
library(vegan)
library(ggrepel)
library(clusterProfiler)
library(org.At.tair.db)
library(cowplot)

# Prepare base theme
base_theme <- theme_bw() +
  theme(
    strip.background = element_rect(color = "white", fill = "white"),
    strip.text.x = element_markdown(size = 10, color = "black", hjust = 0.5, vjust = 0.5),
    legend.title = element_text(size = 11),
    axis.title.y = element_text(size =11),
    legend.text = element_text(size = 10),
    axis.text.x = element_text(size = 10, color = "black", angle = 90, hjust = 1),
    axis.text.y = element_text(size = 10, color = "black"),
    panel.border = element_rect(color = "black", fill = NA, size = 0.75))


# Colors
color_palette <- c("Water" = "#949494", "3OMG" = "#7FC29B",  "Glucose" = "#817E9F")
```

# Before you start

This script has been made on R 4.5.3 (and Rstudio 2026.01.2+418)
```{r R version}
R.version
theme_set(theme_bw())
```
# A] Root length

```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}

df_length <- read.table("Transcripto_length.txt", sep="\t", dec=".", header=TRUE, row.names=1)

df_length$Time <- factor(df_length$Time, levels = c("72h", "1week"))
df_length$Supplementation <- factor(df_length$Supplementation, levels = c("Water", "3OMG"))
df_length$Microbe <- factor(df_length$Microbe, levels = c("M.free", "F16"))

# Test de normalité (Shapiro)
normality_results <- df_length %>%
  group_by(Time, Microbe, Supplementation) %>%
  shapiro_test(Root_Length)
print(normality_results)

# Tests Wilcoxon (Time × Microbe)
Stats_length <- df_length %>%
  group_by(Time, Microbe) %>%
  wilcox_test(Root_Length ~ Supplementation) %>%
  adjust_pvalue(method = "fdr") %>%
  mutate(label_text = paste0("P=", formatC(p.adj, format="e", digits=2)))

# Positions pour les brackets
Stats_length <- Stats_length %>%
  left_join(
    df_length %>%
      group_by(Time, Microbe, Supplementation) %>%
      summarise(ymax = max(Root_Length, na.rm=TRUE), .groups = "drop") %>%
      pivot_wider(names_from = Supplementation, values_from = ymax),
    by = c("Time", "Microbe")
  ) %>%
  mutate(
    y_position = max(df_length$Root_Length, na.rm=TRUE) * 1.05,
    x1 = as.numeric(factor(Microbe)) - 0.2,
    x2 = as.numeric(factor(Microbe)) + 0.2,
    Supplementation = "3OMG"
  )

Stats_length <- Stats_length %>% filter(p.adj < 0.05)

# Comptage des réplicats
rep_counts <- df_length %>%
  group_by(Time, Microbe, Supplementation) %>%
  summarise(n = n(), .groups = "drop") %>%
  left_join(
    df_length %>%
      group_by(Time) %>%
      summarise(y_pos = min(Root_Length, na.rm=TRUE) - 0.05 * diff(range(Root_Length))),
    by = "Time"
  )

# Plot
Plot_length <- ggplot(df_length, aes(x = Microbe, y = Root_Length, fill = Supplementation)) +
  geom_boxplot(alpha = 1, outlier.shape = NA, position = position_dodge(width = 0.75)) +
  geom_jitter(aes(shape = Supplementation),
              alpha = 0.2, size = 1,
              position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.75)) +
  geom_segment(data = Stats_length,
               aes(x = x1, xend = x2, y = y_position, yend = y_position),
               inherit.aes = FALSE) +
  geom_text(data = Stats_length,
            aes(x = (x1 + x2)/2, y = y_position * 1.03, label = label_text),
            size = 3) +
  geom_text(data = rep_counts,
            aes(x = Microbe, y = -0.5, label = paste0("n=", n)),
            position = position_dodge(width = 0.75),
            vjust = 1, size = 3) +
  facet_grid(~Time, space = "free", scales = "free") +
  scale_fill_manual(values = c("#949494", "#7FC29B"), name = "Supplementation") +
  labs(x = "", y = "Root Length (mm)", title = "") +
  theme_bw(base_size = 12) +
  theme(strip.background = element_rect(color = "black", fill = "white"),
        axis.text.x = element_text(angle = 45, hjust = 1))

Plot_length

```

# B] In Planta Transcriptome

## 1) Normalize prepare data

```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}

QC_Sum <- read.table(file="summary_table_salmon_mapping.txt",sep="\t",dec=".",header=TRUE, row.names = 1)
    
QC_Sum_long <- QC_Sum %>%
  gather(key = "read_type", value = "proportion", prop.plant, prop.fungi)

plot_prop <- ggplot(QC_Sum_long, aes(x = sampleID, y = proportion, fill = read_type)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(values = c("prop.plant" = "#521A10", "prop.fungi" = "#F9B244")) +
  labs(x = "Sample ID", y = "Proportion", fill = "Read Type") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

plot_prop

table_counts <- read.table(file="count_table_arath.txt",sep="\t",dec=".",header=TRUE, row.names = 1)

# Merge transcript isoforms by gene ID.
table_counts <- table_counts %>%
  mutate(GeneID = gsub("\\..*", "", TranscriptID)) %>%
  group_by(GeneID) %>%
  summarise(across(where(is.numeric), sum, na.rm = TRUE))


# Remove fungi-only control samples.
cols_to_remove <- c("Control_3OMG_F16_1", "Control_3OMG_F16_2", "Control_3OMG_F16_3", "Control_3OMG_F16_4",
                    "Control_Water_F16_1", "Control_Water_F16_2", "Control_Water_F16_3")

table_counts2 <- table_counts %>% dplyr::select(-all_of(cols_to_remove))


# Remove two samples with low assigned reads.
table_counts_filtered <- data.frame(table_counts2[, !names(table_counts2) %in% c("m3OMG_microbeFree_2", "water_microbeFree_3")])


row.names(table_counts_filtered) <- table_counts_filtered$GeneID
table_counts_filtered$GeneID <- NULL

matrix_counts_Plant <- as.matrix(table_counts_filtered)
```

## 2) PcoA Bray-Curtis
```{r echo=FALSE, fig.show='hide', message=FALSE, results="hide"}
df_transformed <- data.frame(matrix_counts_Plant)

# Transpose count table for sample-level distance analysis.
df_transformed <- data.frame(t(df_transformed))

env <- read.table(file="env2.txt",sep="\t",dec=".",header=TRUE, row.names = 1)

#Adonis
data.div <- adonis2(df_transformed ~ Microbes/Molecule, data = env, permutations = 999, method="canberra")
data.div


dds <- DESeqDataSetFromMatrix(countData = matrix_counts_Plant, colData = env, design = ~ Conditions)

rld <- rlog(dds, blind = FALSE)

pca_res <- prcomp(t(assay(rld)))

percentVar <- round(100 * (pca_res$sdev^2 / sum(pca_res$sdev^2)), 2)

pca_df <- data.frame(PC1 = pca_res$x[, 1],
                     PC2 = pca_res$x[, 2],
                     Condition = env$Conditions)

# Extract molecule and microbe information from sample names.
pca_df <- pca_df %>%
  mutate(Molecule = str_split_fixed(row.names(pca_df), '_', 3)[, 1],
         Microbes = str_split_fixed(row.names(pca_df), '_', 3)[, 2],
         Microbes_Molecule = paste(Microbes, Molecule, sep = " - "))
pca_df


mean_pca <- pca_df %>%
  group_by(Microbes_Molecule) %>%
  summarise(PC1 = mean(PC1), PC2 = mean(PC2)) %>%
  separate(Microbes_Molecule, into = c("Microbes", "Molecule"))

segments_pca <- pca_df %>%
  group_by(Microbes, Molecule) %>%
  mutate(meanX = mean(PC1), meanY = mean(PC2))


pca_df$Environement <- c("Plant Transcriptome")

PcoA_Plant <- ggplot() +
  geom_point(data = pca_df, aes(x = PC1, y = PC2, color = Molecule , shape = Microbes), size = 2, alpha = 1) +
  geom_point(data = mean_pca, aes(x = PC1, y = PC2, color = Molecule, shape = Microbes), size = 5) +
  geom_segment(data = segments_pca, aes(x = meanX, y = meanY, xend = PC1, yend = PC2, color = Molecule), size = 0.7, alpha = 0.5) +
  facet_grid(~Environement, space="free", scales="free")+
  scale_color_manual('Microbes', labels = c("3OMG", "Water"), values = c("#7FC29B", "#949494")) +
 scale_shape_manual('Microbes', labels = c("P. cucumerina", "Microbes Free"), values = c(16, 17)) +
  xlab(paste("PC1 (", percentVar[1], "%)", sep = "")) +
  ylab(paste("PC2 (", percentVar[2], "%)", sep = "")) +
 base_theme
PcoA_Plant
```
## 3) Differential expressed gene analysis 

### a) Run DeSeq2

```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}
dds <- DESeqDataSetFromMatrix(countData = matrix_counts_Plant, colData = env, design = ~ Conditions)


# Pre-filter low-count genes.
smallestGroupSize <- 3
keep <- rowSums(counts(dds) >= 10) >= smallestGroupSize
dds <- dds[keep,]

dds <- DESeq(dds)

conditions <- levels(dds$Conditions)
all_comparisons <- combn(conditions, 2)

results_list <- list()

for (i in 1:ncol(all_comparisons)) {
  contrast <- all_comparisons[, i]
  result_name <- paste(contrast[1], "_vs_", contrast[2], sep = "")
  
  res <- results(dds, contrast = c("Conditions", contrast[1], contrast[2]))
  results_list[[result_name]] <- res
}

names(results_list) <- apply(all_comparisons, 2, function(x) paste(x, collapse = "_vs_"))

# Combine pairwise DESeq2 results.
combined_results <- do.call(rbind, lapply(names(results_list), function(name) {
  res <- as.data.frame(results_list[[name]])
  res$Comparison <- name
  res$GeneID <- row.names(res)
  row.names(res) <- NULL
  res
}))


filtered_results <- combined_results[
  (combined_results$log2FoldChange >= 0.5 | combined_results$log2FoldChange <= -0.5) & 
  combined_results$padj < 0.05, 
]

filtered_results <- na.omit(filtered_results)

count_MF <- subset(filtered_results, Comparison %in% c("3OMG_MF_vs_Water-MF"))
count_MF_df <- table(Direction = ifelse(count_MF$log2FoldChange > 0, "Above 0", "Below 0"))

```

### b) Vulcano Plots

#### i) Loop all conditions

```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}
gene_annot_short <- read.table(file = "At_Gene_Short.txt", sep = "\t", header = TRUE)

condition_levels <- levels(dds$Conditions)

volcano_plots <- list()

for (i in 1:(length(condition_levels) - 1)) {
  for (j in (i + 1):length(condition_levels)) {
    condition1 <- condition_levels[i]
    condition2 <- condition_levels[j]
    
    results_df <- results(dds, contrast = c("Conditions", condition1, condition2)) %>%
      as.data.frame() %>%
      mutate(
        logP = -log10(padj),
        Color = case_when(
          log2FoldChange > 0.5 & padj < 0.05 ~ "#7FC29B",
          log2FoldChange < -0.5 & padj < 0.05 ~ "#949494",
          TRUE ~ "lightgray"
        )
      ) %>%
      rownames_to_column("locus_name") %>%
      inner_join(gene_annot_short, by = "locus_name") %>%
      mutate(
        symbol_2 = ifelse(Color != "lightgray", symbol, NA)
      ) %>%
      distinct(locus_name, .keep_all = TRUE)

    top_pos_loci <- results_df %>%
      filter(log2FoldChange > 0) %>%
      arrange(desc(logP)) %>%
      head(30)

    top_neg_loci <- results_df %>%
      filter(log2FoldChange < 0) %>%
      arrange(desc(logP)) %>%
      head(30)

    top_loci <- bind_rows(top_pos_loci, top_neg_loci)

    volcano_plot <- ggplot(results_df, aes(x = log2FoldChange, y = logP, label = symbol_2, color = Color)) +
      geom_point(size = 2) +
      geom_vline(xintercept = c(-0.5, 0.5), linetype = "dashed", color = "darkgray", size = 0.5) +
      geom_hline(yintercept = -log10(0.05), linetype = "dashed", color = "darkgray", size = 0.5) +
      geom_text_repel(
        data = top_loci,
        aes(label = symbol),
        size = 3,
        max.overlaps = 50,
        box.padding = 0.5,
        point.padding = 0.5,
        segment.size = 0.2,
        segment.color = "grey50"
      ) +
     base_theme +
      scale_color_identity() +
      labs(
        title = paste("Volcano Plot:", condition1, "vs", condition2),
        x = "Log2 Fold Change",
        y = "-Log10(p-value)"
      ) +
      theme(legend.position = "none")

    plot_name <- paste(condition1, "vs", condition2, sep = "_")
    volcano_plots[[plot_name]] <- volcano_plot
  }
}

volcano_plots[[1]]

output_dir <- "volcano_plots/"
dir.create(output_dir, showWarnings = FALSE)
for (name in names(volcano_plots)) {
  ggsave(filename = paste0(output_dir, name, ".png"), plot = volcano_plots[[name]], width = 8, height = 6)
}
```
#### ii) 3OMG_MF VS Water-MF

```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}

gene_annot_short <- read.table(file = "At_Gene_Short.txt", sep = "\t", header = TRUE)

condition1 <- c("3OMG_MF")
condition2 <- c("Water-MF")
  
results1 <- results(dds, contrast = c("Conditions", condition1, condition2)) %>%
  as.data.frame() %>%
  mutate(
    logP = -log10(padj),
    Color = case_when(
      log2FoldChange > 0.5 & padj < 0.05 ~ "#7FC29B",
      log2FoldChange < -0.5 & padj < 0.05 ~ "#949494",
      TRUE ~ "lightgray"
    )
  )

results1 <- results1 %>%
  rownames_to_column("locus_name") %>%
  left_join(gene_annot_short, by = "locus_name") %>%
  mutate(
    symbol_2 = ifelse(Color != "lightgray", symbol, NA)
  ) %>%
  distinct(locus_name, .keep_all = TRUE)

    top_pos_loci <- results1 %>%
      filter(log2FoldChange > 0.5) %>%
      arrange(desc(logP))

    top_neg_loci <- results1 %>%
      filter(log2FoldChange < -0.5) %>%
      arrange(desc(logP))

    top_loci <- bind_rows(top_pos_loci, top_neg_loci)

genes_to_highlight <- c("AT1G59750","AT1G11260", "AT5G01770", "AT1G50030", "AT1G10760", "AT5G26570", "AT1G22770", "AT1G69830", "AT3G46970", "AT1G01060", "AT4G36880", "AT3G23390", "AT3G05600")

top_loci_f <- top_loci %>%
  filter(locus_name %in% genes_to_highlight)
  
    results1$Comp <- paste(condition1, "vs", condition2)

df_results_MF <- results1
  
    volcano_results1 <- ggplot(results1, aes(x = log2FoldChange, y = logP, label = symbol)) +
      geom_point(aes(color = Color),size = 2) +
      geom_vline(xintercept = c(-0.5, 0.5), linetype = "dashed", color = "darkgray", size = 0.5) +
      geom_hline(yintercept = -log10(0.05), linetype = "dashed", color = "darkgray", size = 0.5) +
      facet_grid(~Comp, space="free", scales="free")+
      geom_text_repel(
        data = top_loci_f,
        aes(label = symbol_2),
        size = 2.5,
        max.overlaps = 50,
        box.padding = 0.5,
        point.padding = 0.5,
        segment.size = 0.2,
        segment.color = "grey50"
      ) +
   base_theme +
      scale_color_identity() +
      labs(
        x = "Log2 Fold Change",
        y = "-Log10(p-value)"
      )  + ylim(0, 22) + xlim(-4,4)

volcano_results1

volcano_results1_MF <- volcano_results1
```


#### iii) 3OMG_F16 VS Water_F16

```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}

gene_annot_short <- read.table(file = "gene_annotations_Selec_f.txt", sep = "\t", header = TRUE)

condition1 <- c("3OMG_F16")
condition2 <- c("Water_F16")

gene_annot_short <- gene_annot_short[gene_annot_short$X3OMGF16_WaterF16 != "NULL", ]

  
results1 <- results(dds, contrast = c("Conditions", condition1, condition2)) %>%
  as.data.frame() %>%
  mutate(
    logP = -log10(padj),
    Color = case_when(
      log2FoldChange > 0.5 & padj < 0.05 ~ "#7FC29B",
      log2FoldChange < -0.5 & padj < 0.05 ~ "#949494",
      TRUE ~ "lightgray"
    )
  )

results1 <- results1 %>%
  rownames_to_column("locus_name") %>%
  left_join(gene_annot_short, by = "locus_name") %>%
  mutate(
    symbol_2 = ifelse(Color != "lightgray", symbol, NA)
  ) %>%
  distinct(locus_name, .keep_all = TRUE)

    top_pos_loci <- results1 %>%
      filter(log2FoldChange > 0.5) %>%
      arrange(desc(logP)) %>%
      head(50)

    top_neg_loci <- results1 %>%
      filter(log2FoldChange < -0.5) %>%
      arrange(desc(logP)) %>%
      head(50)

    top_loci <- bind_rows(top_pos_loci, top_neg_loci)

    results1$Comp <- paste(condition1, "vs", condition2)
    
    
df_results_F16 <- results1

    volcano_results1 <- ggplot(results1, aes(x = log2FoldChange, y = logP, label = symbol)) +
      geom_point(aes(color = Color),size = 2) +
      geom_vline(xintercept = c(-0.5, 0.5), linetype = "dashed", color = "darkgray", size = 0.5) +
      geom_hline(yintercept = -log10(0.05), linetype = "dashed", color = "darkgray", size = 0.5) +
      facet_grid(~Comp, space="free", scales="free")+
      geom_text_repel(
        data = top_loci,
        aes(label = symbol),
        size = 2.5,
        max.overlaps = 50,
        box.padding = 0.5,
        point.padding = 0.5,
        segment.size = 0.2,
        segment.color = "grey50"
      ) +
base_theme +
      scale_color_identity() +
      labs(
        x = "Log2 Fold Change",
        y = "-Log10(p-value)"
      )  + ylim(0, 22) + xlim(-4,4)

volcano_results1

volcano_results1_F16 <- volcano_results1
```

## 4) Go Term enrichment analysis

### a) 3OMG_MF VS Water-MF

```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}
condition1 <- c("3OMG MF")
condition2 <- c("Water MF")

gene_list_red <- df_results_MF$locus_name[df_results_MF$Color == "#7FC29B"]
gene_list_blue <- df_results_MF$locus_name[df_results_MF$Color == "#949494"]

gene_list_red <- gene_list_red[!is.na(gene_list_red)]
gene_list_blue <- gene_list_blue[!is.na(gene_list_blue)]

converted_genes_red <- bitr(gene_list_red, fromType = "TAIR", 
                            toType = c("ENTREZID", "SYMBOL"), 
                            OrgDb = org.At.tair.db)
converted_genes_blue <- bitr(gene_list_blue, fromType = "TAIR", 
                             toType = c("ENTREZID", "SYMBOL"), 
                             OrgDb = org.At.tair.db)


ego_red <- enrichGO(gene         = converted_genes_red$ENTREZID,
                    OrgDb        = org.At.tair.db,
                    keyType      = "ENTREZID",
                    ont          = "ALL",
                    pAdjustMethod = "BH",
                    pvalueCutoff  = 0.05,
                    qvalueCutoff  = 0.05)

ego_blue <- enrichGO(gene         = converted_genes_blue$ENTREZID,
                     OrgDb        = org.At.tair.db,
                     keyType      = "ENTREZID",
                     ont          = "ALL",
                     pAdjustMethod = "BH",
                     pvalueCutoff  = 0.05,
                     qvalueCutoff  = 0.05)

data_red <- data.frame(ego_red)
data_blue <- data.frame(ego_blue)

data_red <- data_red %>%
  mutate(GeneRatioPercentage = (as.numeric(sapply(strsplit(as.character(GeneRatio), "/"), "[", 1)) / 
                               as.numeric(sapply(strsplit(as.character(GeneRatio), "/"), "[", 2))) * 100)

data_blue <- data_blue %>%
  mutate(GeneRatioPercentage = (as.numeric(sapply(strsplit(as.character(GeneRatio), "/"), "[", 1)) / 
                               as.numeric(sapply(strsplit(as.character(GeneRatio), "/"), "[", 2))) * 100)

data_top10_red <- data_red %>%
  arrange(p.adjust) %>%
  slice_head(n = 15)

data_top10_blue <- data_blue %>%
  arrange(p.adjust) %>%
  slice_head(n = 15)

data_red$Comp <- paste("GO Term Up", condition1, "vs", condition2)
data_blue$Comp <- paste("GO Term Down", condition1, "vs", condition2)

data_top10_red$Comp <- paste("GO Term Up", condition1, "vs", condition2)
data_top10_blue$Comp <- paste("GO Term Down", condition1, "vs", condition2)

data_top10_red$Description <- data_top10_red$Description %>%
  gsub("metabolic process", "metabolism", .) %>%
  gsub("biosynthetic process", "biosynthesis", .) %>%
  gsub("catabolic process", "catabolism", .) %>%
  gsub("transporter activity", "transport", .)

data_top10_blue$Description <- data_top10_blue$Description %>%
  gsub("metabolic process", "metabolism", .) %>%
  gsub("biosynthetic process", "biosynthesis", .) %>%
  gsub("catabolic process", "catabolism", .) %>%
  gsub("transporter activity", "transport", .)

plot_dot_MF_up <- ggplot(data_top10_red, aes(x = -log10(p.adjust), y = reorder(Description, -log10(p.adjust)), fill = -log10(p.adjust))) +
  geom_point(aes(size = Count), shape = 21) +
  facet_grid(~Comp, space="free", scales="free")+
  scale_fill_gradient(low = "#D1E8D3", high = "#4F8C6B") +
  labs(x = "-log10 p-value", y = "") +
 base_theme +
  theme(legend.position = "bottom") +
   scale_y_discrete(position = "right")
plot_dot_MF_up

plot_dot_MF_down <- ggplot(data_top10_blue, aes(x = -log10(p.adjust), y = reorder(Description, -log10(p.adjust)), fill =  -log10(p.adjust))) +
  geom_point(aes(size = Count), shape = 21) +
  facet_grid(~Comp, space="free", scales="free")+
  scale_fill_gradient(low = "lightgray", high = "black") +
  labs(x = "-log10 p-value", y = "") +
 base_theme +
  theme(legend.position = "bottom")
plot_dot_MF_down

data_top10_blue$Color <- c("darkblue")
data_top10_red$Color <- c("darkred")

data_GoTerm <- rbind(data_top10_blue, data_top10_red)
```

### b) 3OMG_F16 VS Water_F16

```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}
condition1 <- c("3OMG F16")
condition2 <- c("Water F16")

gene_list_red <- df_results_F16$locus_name[df_results_F16$Color == "#7FC29B"]
gene_list_blue <- df_results_F16$locus_name[df_results_F16$Color == "#949494"]

gene_list_red <- gene_list_red[!is.na(gene_list_red)]
gene_list_blue <- gene_list_blue[!is.na(gene_list_blue)]

converted_genes_red <- bitr(gene_list_red, fromType = "TAIR", 
                            toType = c("ENTREZID", "SYMBOL"), 
                            OrgDb = org.At.tair.db)
converted_genes_blue <- bitr(gene_list_blue, fromType = "TAIR", 
                             toType = c("ENTREZID", "SYMBOL"), 
                             OrgDb = org.At.tair.db)

ego_red <- enrichGO(gene         = converted_genes_red$ENTREZID,
                    OrgDb        = org.At.tair.db,
                    keyType      = "ENTREZID",
                    ont          = "ALL",
                    pAdjustMethod = "BH",
                    pvalueCutoff  = 0.05,
                    qvalueCutoff  = 0.05)

ego_blue <- enrichGO(gene         = converted_genes_blue$ENTREZID,
                     OrgDb        = org.At.tair.db,
                     keyType      = "ENTREZID",
                     ont          = "ALL",
                     pAdjustMethod = "BH",
                     pvalueCutoff  = 0.05,
                     qvalueCutoff  = 0.05)

data_red <- data.frame(ego_red)
data_blue <- data.frame(ego_blue)

data_red <- data_red %>%
  mutate(GeneRatioPercentage = (as.numeric(sapply(strsplit(as.character(GeneRatio), "/"), "[", 1)) / 
                               as.numeric(sapply(strsplit(as.character(GeneRatio), "/"), "[", 2))) * 100)

data_blue <- data_blue %>%
  mutate(GeneRatioPercentage = (as.numeric(sapply(strsplit(as.character(GeneRatio), "/"), "[", 1)) / 
                               as.numeric(sapply(strsplit(as.character(GeneRatio), "/"), "[", 2))) * 100)

data_top10_red <- data_red %>%
  arrange(p.adjust) %>%
  slice_head(n = 15)

data_top10_blue <- data_blue %>%
  arrange(p.adjust) %>%
  slice_head(n = 15)

data_red$Comp <- paste("GO Term Up", condition1, "vs", condition2)
data_blue$Comp <- paste("GO Term Down", condition1, "vs", condition2)

data_top10_red$Comp <- paste("GO Term Up", condition1, "vs", condition2)
data_top10_blue$Comp <- paste("GO Term Down", condition1, "vs", condition2)

data_top10_red$Description <- data_top10_red$Description %>%
  gsub("metabolic process", "metabolism", .) %>%
  gsub("biosynthetic process", "biosynthesis", .) %>%
  gsub("catabolic process", "catabolism", .) %>%
  gsub("transporter activity", "transport", .)

data_top10_blue$Description <- data_top10_blue$Description %>%
  gsub("metabolic process", "metabolism", .) %>%
  gsub("biosynthetic process", "biosynthesis", .) %>%
  gsub("catabolic process", "catabolism", .) %>%
  gsub("transporter activity", "transport", .)


plot_dot_F16_up <- ggplot(data_top10_red, aes(x = -log10(p.adjust), y = reorder(Description, -log10(p.adjust)), fill = -log10(p.adjust))) +
  geom_point(aes(size = Count), shape = 21) +
  facet_grid(~Comp, space="free", scales="free")+
  scale_fill_gradient(low = "#D1E8D3", high = "#4F8C6B") +
  labs(x = "-log10 p-value", y = "") +
 base_theme +
  theme(legend.position = "bottom") +
   scale_y_discrete(position = "right")
plot_dot_F16_up

plot_dot_F16_down <- ggplot(data_top10_blue, aes(x = -log10(p.adjust), y = reorder(Description, -log10(p.adjust)), fill =  -log10(p.adjust))) +
  geom_point(aes(size = Count), shape = 21) +
  facet_grid(~Comp, space="free", scales="free")+
  scale_fill_gradient(low = "lightgray", high = "black") +
  labs(x = "-log10 p-value", y = "") +
 base_theme +
  theme(legend.position = "bottom")
plot_dot_F16_down

data_top10_blue$Color <- c("darkblue")
data_top10_red$Color <- c("darkred")

data_GoTerm <- rbind(data_top10_blue, data_top10_red)
```



## 5) All plots Plant

```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}

# Create an empty plot
empty_plot <- ggplot() + 
  theme_void()  # completely blank

Plot_Comb1 <- plot_grid(volcano_results1_MF, volcano_results1_F16, rel_heights = c(1,1), rel_widths = c(1,1), nrow=1,   align="hv",  axis = "tb")
Plot_Comb1

Plot_Comb_up <- plot_grid(plot_dot_MF_up, plot_dot_F16_up, rel_heights = c(1,1), rel_widths = c(1,1), nrow=2,   align="hv",  axis = "tb")
Plot_Comb_up

Plot_Comb_down <- plot_grid(plot_dot_MF_down,plot_dot_F16_down, rel_heights = c(1,1), rel_widths = c(1,1), nrow=2,   align="hv",  axis = "tb")
Plot_Comb_down

#Export 15x14
Plot_Comb_MF <- plot_grid(
  plot_dot_MF_down, volcano_results1_MF, plot_dot_MF_up, rel_heights = c(1,1,1), rel_widths = c(0.6,0.5,0.75), nrow=1,   align="h",  axis = "tb")
Plot_Comb_MF

Plot_Comb_F16 <- plot_grid(
  plot_dot_F16_down, volcano_results1_F16, plot_dot_F16_up, rel_heights = c(1,1,1), rel_widths = c(0.7,0.5,0.65), nrow=1,   align="h",  axis = "tb")
Plot_Comb_F16

#Plot verti
Plot_Comb_Vulc <- plot_grid( volcano_results1_MF,empty_plot,  volcano_results1_F16, empty_plot, rel_heights = c(1,0.25,1,0.25), rel_widths = c(1), nrow=4,ncol=1,   align="vh",  axis = "tb")
Plot_Comb_Vulc

Plot_Comb_Down <- plot_grid( plot_dot_MF_down, plot_dot_F16_down, rel_heights = c(1,1), rel_widths = c(1,1), nrow=2,   align="vh",  axis = "tb")
Plot_Comb_Down

Plot_Comb_Up <- plot_grid( plot_dot_MF_up, plot_dot_F16_up, rel_heights = c(1,1), rel_widths = c(1,1), nrow=2,   align="vh",  axis = "tb")
Plot_Comb_Up

Plot_Comb_Final <- plot_grid( Plot_Comb_Down, Plot_Comb_Vulc,Plot_Comb_Up,  rel_heights = c(1,1), rel_widths = c(0.85,0.6,1), nrow=1,   align="vh",  axis = "tb")
Plot_Comb_Final
```

---
END
---
