---
title: "Decoy sugars hijack fungal carbon sensing and prevent plant disease"
author: "Guillaume Chesneau"
date: "26-05-2026"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "") 
```


# Before you start

This script has been made on R 4.5.3 (and Rstudio 2026.01.2+418)
```{r R version}
R.version
```

```{r}
#load libraries
library(rstatix) 
library(ggpubr)  
library(dunn.test)
library(ggthemes)
library(ggtext)
library(tidyr)
library(markdown)
library(ggplot2)
library(dplyr)
library(ggrepel)
library(Biostrings)
library(DECIPHER)
library(scales)
library(ape)
library(ggtree)
library(cowplot)

#Colors
color_palette <- c("Water" = "#949494", "3OMG" = "#7FC29B")
theme_set(theme_bw())

# Prepare base theme
base_theme <- theme_bw() +
  theme(
    strip.background = element_rect(color = "white", fill = "white"),
    strip.text.x = element_markdown(size = 10, color = "black", hjust = 0.5, vjust = 0.5),
    legend.title = element_text(size = 11),
    axis.title.y = element_text(size =11),
    legend.text = element_text(size = 10),
    axis.text.x = element_text(size = 10, color = "black", angle = 90, hjust = 1),
    axis.text.y = element_text(size = 10, color = "black"),
    panel.border = element_rect(color = "black", fill = NA, size = 0.75))

#Set the levels of the experiment
Microbes_levels<-c("MF","F16")

Treatment_levels<-c ("Water","3OMG")
labels <- c(
  'MF' = "Microbe free",
  'F16' = "*Plectosphaerella*<br>*cucumerina*",
  'MS_agar' = "Agar 1/2 MS",
  'Arabidopsis_thaliana' = "A. thaliana"
)
```

# A] Analysis growth qPCR

## 1) Pre-processing

```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}
df_qPCR <- read.table(file="qPCR_Drop_out_vf.txt",sep="\t",dec=".",header=TRUE)
df_qPCR <- subset(df_qPCR, !Expe %in% c("Control", "SynCom"))
df_qPCR_SC <- subset(df_qPCR, Expe %in% c("SynCom"))


#Put non detected Microbial primers as a value of max
maxITS <- max(df_qPCR$CtITS, na.rm = TRUE)
df_qPCR$CtITS[is.na(df_qPCR$CtITS)] <- maxITS
df_qPCR$CtITS <- as.numeric (df_qPCR$CtITS)

max16S <- max(df_qPCR$Ct16S, na.rm = TRUE)
df_qPCR$Ct16S[is.na(df_qPCR$Ct16S)] <- max16S
df_qPCR$Ct16S <- as.numeric (df_qPCR$Ct16S)

#Remove empty rowspty
df_qPCR <- df_qPCR %>% filter(Strains != "")

#Remove empty rowspty
df_qPCR_init <- df_qPCR %>% filter(Strains != "")
```

## 2) Statistics Raw Data

### a) Mono culture based on BFSC comparison

```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}
#Just monoculture
df_qPCR_mono <- subset(df_qPCR_init, Expe2 %in% c("mono"))

# Function to convert p-values to stars
pval_to_stars <- function(p) {
  ifelse(p <= 0.001, "***",
  ifelse(p <= 0.01,  "**",
  ifelse(p <= 0.05,  "*", "ns")))
}


# Overall Comp values (control)
Comp_CtITS <- df_qPCR_mono$CtITS[df_qPCR_mono$Strains == "BFSC"]
Comp_Ct16S <- df_qPCR_mono$Ct16S[df_qPCR_mono$Strains == "BFSC"]

# Get all unique experiments
experiments <- unique(df_qPCR_mono$Expe2)

# Initialize empty results lists
results_ITS <- list()
results_16S <- list()

# Loop over each experiment
for (exp in experiments) {
  
  df_exp <- df_qPCR_mono[df_qPCR_mono$Expe2 == exp, ]
  
  # Strains to compare (exclude Comp)
  other_strains <- setdiff(unique(df_exp$Strains), "Comp")
  
  for (strain in other_strains) {
    
    # CtITS
    strain_vals_ITS <- df_exp$CtITS[df_exp$Strains == strain]
    test_ITS <- wilcox.test(strain_vals_ITS, Comp_CtITS, alternative = "greater")
    
    results_ITS[[paste(exp, strain, sep="_")]] <- data.frame(
      Expe2 = exp,
      Strain = strain,
      W = test_ITS$statistic,
      p_value = test_ITS$p.value,
      n_strain = length(strain_vals_ITS),
      n_Comp = length(Comp_CtITS),
      stars = pval_to_stars(test_ITS$p.value)
    )
    
    # Ct16S
    strain_vals_16S <- df_exp$Ct16S[df_exp$Strains == strain]
    test_16S <- wilcox.test(strain_vals_16S, Comp_Ct16S, alternative = "greater")
    
    results_16S[[paste(exp, strain, sep="_")]] <- data.frame(
      Expe2 = exp,
      Strain = strain,
      W = test_16S$statistic,
      p_value = test_16S$p.value,
      n_strain = length(strain_vals_16S),
      n_Comp = length(Comp_Ct16S),
      stars = pval_to_stars(test_16S$p.value)
    )
    
  }
}

# Combine results into dataframes
results_ITS_MONO <- do.call(rbind, results_ITS)
results_16S_MONO <- do.call(rbind, results_16S)

#Adjust p-value
results_ITS_MONO$p_adj <- p.adjust(results_ITS_MONO$p_value, method = "BH")
results_16S_MONO$p_adj <- p.adjust(results_16S_MONO$p_value, method = "BH")

#Add stars
results_ITS_MONO$stars_adj <- pval_to_stars(results_ITS_MONO$p_adj)
results_16S_MONO$stars_adj <- pval_to_stars(results_16S_MONO$p_adj)

# View results
results_ITS_MONO
results_16S_MONO
```

### b) Pairwise culture based on BFSC comparison

```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}
#Just monoculture
df_qPCR_pair <- subset(df_qPCR_init, Expe2 %in% c("Pairwise"))

# Function to convert p-values to stars
pval_to_stars <- function(p) {
  ifelse(p <= 0.001, "***",
  ifelse(p <= 0.01,  "**",
  ifelse(p <= 0.05,  "*", "ns")))
}

# Overall Comp values (control)
Comp_CtITS <- df_qPCR_pair$CtITS[df_qPCR_pair$Strains == "BFSC"]
Comp_Ct16S <- df_qPCR_pair$Ct16S[df_qPCR_pair$Strains == "BFSC"]

# Get all unique experiments
experiments <- unique(df_qPCR_pair$Expe2)

# Initialize empty results lists
results_ITS <- list()
results_16S <- list()

# Loop over each experiment
for (exp in experiments) {
  
  df_exp <- df_qPCR_pair[df_qPCR_pair$Expe2 == exp, ]
  
  # Strains to compare (exclude Comp)
  other_strains <- setdiff(unique(df_exp$Strains), "Comp")
  
  for (strain in other_strains) {
    
    # CtITS
    strain_vals_ITS <- df_exp$CtITS[df_exp$Strains == strain]
    test_ITS <- wilcox.test(strain_vals_ITS, Comp_CtITS, alternative = "greater")
    
    results_ITS[[paste(exp, strain, sep="_")]] <- data.frame(
      Expe2 = exp,
      Strain = strain,
      W = test_ITS$statistic,
      p_value = test_ITS$p.value,
      n_strain = length(strain_vals_ITS),
      n_Comp = length(Comp_CtITS),
      stars = pval_to_stars(test_ITS$p.value)
    )
    
    # Ct16S
    strain_vals_16S <- df_exp$Ct16S[df_exp$Strains == strain]
    test_16S <- wilcox.test(strain_vals_16S, Comp_Ct16S, alternative = "greater")
    
    results_16S[[paste(exp, strain, sep="_")]] <- data.frame(
      Expe2 = exp,
      Strain = strain,
      W = test_16S$statistic,
      p_value = test_16S$p.value,
      n_strain = length(strain_vals_16S),
      n_Comp = length(Comp_Ct16S),
      stars = pval_to_stars(test_16S$p.value)
    )
    
  }
}

# Combine results into dataframes
results_ITS_PAIR <- do.call(rbind, results_ITS)
results_16S_PAIR <- do.call(rbind, results_16S)

#Adjust p-value
results_ITS_PAIR$p_adj <- p.adjust(results_ITS_PAIR$p_value, method = "BH")
results_16S_PAIR$p_adj <- p.adjust(results_16S_PAIR$p_value, method = "BH")

#Add stars
results_ITS_PAIR$stars_adj <- pval_to_stars(results_ITS_PAIR$p_adj)
results_16S_PAIR$stars_adj <- pval_to_stars(results_16S_PAIR$p_adj)

# View results
results_ITS_PAIR
results_16S_PAIR
```

### c) Pseudomonas culture based on F16+R329 comparison

```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}
#Just monoculture
df_qPCR_pseudo <- subset(df_qPCR_init, Expe2 %in% c("Pseudomonas"))

# Function to convert p-values to stars
pval_to_stars <- function(p) {
  ifelse(p <= 0.001, "***",
  ifelse(p <= 0.01,  "**",
  ifelse(p <= 0.05,  "*", "ns")))
}


# Overall Comp values (control)
Comp_CtITS <- df_qPCR_pseudo$CtITS[df_qPCR_pseudo$Strains == "R329_2"]
Comp_Ct16S <- df_qPCR_pseudo$Ct16S[df_qPCR_pseudo$Strains == "R329_2"]

# Get all unique experiments
experiments <- unique(df_qPCR_pseudo$Expe2)

# Initialize empty results lists
results_ITS <- list()
results_16S <- list()

# Loop over each experiment
for (exp in experiments) {
  
  df_exp <- df_qPCR_pseudo[df_qPCR_pseudo$Expe2 == exp, ]
  
  # Strains to compare (exclude Comp)
  other_strains <- setdiff(unique(df_exp$Strains), "Comp")
  
  for (strain in other_strains) {
    
    # CtITS
    strain_vals_ITS <- df_exp$CtITS[df_exp$Strains == strain]
    test_ITS <- wilcox.test(strain_vals_ITS, Comp_CtITS, alternative = "greater")
    
    results_ITS[[paste(exp, strain, sep="_")]] <- data.frame(
      Expe2 = exp,
      Strain = strain,
      W = test_ITS$statistic,
      p_value = test_ITS$p.value,
      n_strain = length(strain_vals_ITS),
      n_Comp = length(Comp_CtITS),
      stars = pval_to_stars(test_ITS$p.value)
    )
    
    # Ct16S
    strain_vals_16S <- df_exp$Ct16S[df_exp$Strains == strain]
    test_16S <- wilcox.test(strain_vals_16S, Comp_Ct16S, alternative = "greater")
    
    results_16S[[paste(exp, strain, sep="_")]] <- data.frame(
      Expe2 = exp,
      Strain = strain,
      W = test_16S$statistic,
      p_value = test_16S$p.value,
      n_strain = length(strain_vals_16S),
      n_Comp = length(Comp_Ct16S),
      stars = pval_to_stars(test_16S$p.value)
    )
    
  }
}

# Combine results into dataframes
results_ITS_PSEUDO <- do.call(rbind, results_ITS)
results_16S_PSEUDO <- do.call(rbind, results_16S)

#Adjust p-value
results_ITS_PSEUDO$p_adj <- p.adjust(results_ITS_PSEUDO$p_value, method = "BH")
results_16S_PSEUDO$p_adj <- p.adjust(results_16S_PSEUDO$p_value, method = "BH")

#Add stars
results_ITS_PSEUDO$stars_adj <- pval_to_stars(results_ITS_PSEUDO$p_adj)
results_16S_PSEUDO$stars_adj <- pval_to_stars(results_16S_PSEUDO$p_adj)

# View results
results_ITS_PSEUDO
results_16S_PSEUDO
```


## 2) Plot Mono + SynCom

Normalize data per F16 values for ITS, and for bacteria by no-growing 16SrRNA 

### a) Normalize data
```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}
#Keep for latter plot
df_qPCR <- df_qPCR_init

#Plot 1, mono + pairwise
df_qPCR <- subset(df_qPCR, Expe2 %in% c("mono", "Pairwise"))

#For ITS
# Compute median CtITS for FSC strains in Expe == "Mono", F16 never show significant growth.
median_F16_mono <- median(df_qPCR$CtITS[df_qPCR$Strains == c("F16")], na.rm = TRUE)

# Add this median as a new column to all rows
df_qPCR <- df_qPCR %>%
  mutate(Median_CtITS_F16_Mono = median_F16_mono)

# Compute overall median Ct16S for the selected strains per Expe (based on signficiance for OD600 values P <0.001)
median_F16 <- median(df_qPCR$Ct16S[df_qPCR$Expe == "F16" & df_qPCR$Strains %in% c( "R09","R16", "R147",  "R420", "R431", "R565", "R568", "R695")], na.rm = TRUE)
median_mono <- median(df_qPCR$Ct16S[df_qPCR$Expe == "mono" & df_qPCR$Strains %in% c("R09","R16", "R147",  "R431", "R565", "R568", "R695")], na.rm = TRUE)

# Create a new column with the median for each Expe
df_qPCR <- df_qPCR %>%
  mutate(Median_Ct16S = case_when(
    Expe == "F16" ~ median_mono,
    Expe == "mono" ~ median_mono,
    TRUE ~ NA_real_  # Ensure other conditions get NA
  ))

#Normalize data (strains which has not grown based on OD600, where used to calculate the CT value for absence of growth, data were normalized accordingly)
#ITS
df_qPCR <- df_qPCR %>%
  mutate(CtITS_norm = ifelse(CtITS > Median_CtITS_F16_Mono, Median_CtITS_F16_Mono, CtITS))


#16SrRNA
df_qPCR <- df_qPCR %>%
  mutate(Ct16S_norm = ifelse(Ct16S > median_mono, median_mono, Ct16S))
```


### b) Plot
```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}
# Replace specific values in Expe column with "Communities"
df_qPCR <- df_qPCR %>%
  mutate(Expe = ifelse(Strains %in% c("BFSC", "BSC", "FSC", "F16_BSC"), "Communities", Expe))

# Calculate medians using raw -Ct values (not transformed), grouping by Strains and Expe (now with Communities)
medians <- df_qPCR %>%
  group_by(Strains, Expe, Expe2, Size, Kingdom) %>%
  summarise(
    med_x = median(-CtITS_norm, na.rm = TRUE),
    med_y = median(-Ct16S_norm, na.rm = TRUE),
    .groups = "drop"
  )

# Create jittered median coordinates to spread median dots and labels
set.seed(123)  # reproducibility
medians_jittered <- medians %>%
  mutate(
    med_x_jitter = med_x + runif(n(), -0.15, 0.15),
    med_y_jitter = med_y + runif(n(), -0.15, 0.15)
  )

# Change Expe label names in both dataframes to keep consistent
df_qPCR <- df_qPCR %>%
  mutate(
    Expe = case_when(
      Expe == "F16" ~ "Pairwise F16",
      Expe == "mono" ~ "Monoculture",
      TRUE ~ Expe
    )
  )

medians <- medians %>%
  mutate(
    Expe = case_when(
      Expe == "F16" ~ "Pairwise F16",
      Expe == "mono" ~ "Monoculture",
      TRUE ~ Expe
    )
  )

medians_jittered <- medians_jittered %>%
  mutate(
    Expe = case_when(
      Expe == "F16" ~ "Pairwise F16",
      Expe == "mono" ~ "Monoculture",
      TRUE ~ Expe
    )
  )


# Custom colors
Kingdom_colors <- c("Bacteria" = "#2780F5", "Fungi" = "#FFD580", "Bacteria.Fungi" = "darkgray", "NA" = "black")

# Reorder factor levels
#df_qPCR$Expe <- factor(df_qPCR$Expe, levels = c("Monoculture","Pairwise F16", "Communities"))
#medians$Expe <- factor(medians$Expe, levels = c("Monoculture","Pairwise F16", "Communities"))
#medians_jittered$Expe <- factor(medians_jittered$Expe, levels = c("Monoculture","Pairwise F16", "Communities"))

# Remove NA in Kingdom
df_qPCR <- df_qPCR %>% filter(!is.na(Kingdom))
medians <- medians %>% filter(!is.na(Kingdom))
medians_jittered <- medians_jittered %>% filter(!is.na(Kingdom))

#Change label
Culture_names <- c(
  'mono' = "Monoculture",
  'Pairwise' = "Pairwise (B+F16)",
  'Pseudomonas' = "Pseudomonas only")

#Plot
PlotScatterRaw <- ggplot(df_qPCR, aes(x = -CtITS_norm, y = -Ct16S_norm, size = Size)) +
  # big white outline (halo)
  geom_point(aes(color= Kingdom), size = 4, color = "white", alpha = 1,
             position = position_jitter(width = 0.15, height = 0.15)) +
  # actual jittered points
  geom_point(aes(color= Kingdom, shape = Rep), size = 2, alpha = 0.2,
             position = position_jitter(width = 0.15, height = 0.15)) +
  
  # median dots with outline
  geom_point(data = medians_jittered,
             aes(x = med_x_jitter, y = med_y_jitter, size = Size, fill = Kingdom),
            stroke = 1, shape = 21, color = "black") +
  # labels
geom_text_repel(
  data = subset(
    medians_jittered,
    Size == "SynCom" | 
      med_y_jitter > -26 | 
      med_x_jitter > -26.2 | 
      Strains %in% c("R569_2m", "R9_2m")
  ),
  aes(x = med_x_jitter, y = med_y_jitter, label = Strains),
  inherit.aes = FALSE,
  size = 2.5,
  max.overlaps = Inf,
  segment.color = "grey50",
  segment.size = 0.3) +
  labs(x = "Fungi abundance (-Ct ITS_norm)",
       y = "Bacteria abundance (-Ct 16S_norm)",
       color = "Experiment",
       shape = "Experiment") +
facet_grid( Expe2 ~ ., labeller = as_labeller(Culture_names)) +
  scale_color_manual(values = Kingdom_colors) +
  scale_fill_manual(values = Kingdom_colors) +
  theme(strip.background = element_rect(color = "black", fill = "white", size = 1, linetype = "solid"),
        strip.text.x = element_text(size = 12, color = "black"),
        strip.text.y = element_text(size = 12, color = "black"),
        axis.text.x = element_text(angle = 90, hjust = 1)) + 
  geom_hline(yintercept = -median_mono, color = "black", linetype = "dashed", alpha = 0.2) +
  geom_vline(xintercept = -median_F16_mono, color = "black", linetype = "dashed", alpha = 0.2) +
  base_theme


PlotScatterRaw
```
## 3) Plot Pseudomonas

Normalize data per F16 values for ITS, and for bacteria by no-growing 16SrRNA 

### a) Normalize data
```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}
#Plot 1, mono + pairwise
df_qPCR2 <- subset(df_qPCR_init, !Expe2 %in% c("mono", "Pairwise", "SynCom"))

#For ITS
# Compute median CtITS for FSC strains in Expe == "Mono", F16 never show significant growth.
median_F16_mono <- median(df_qPCR$CtITS[df_qPCR$Strains == c("F16")], na.rm = TRUE)

# Add this median as a new column to all rows
df_qPCR2 <- df_qPCR2 %>%
  mutate(Median_CtITS_F16_Mono = median_F16_mono)

# Compute overall median Ct16S for the selected strains per Expe (based on signficiance for OD600 values P <0.001)
median_mono <- median(df_qPCR2$Ct16S[df_qPCR2$Expe2 == "Pseudomonas" & df_qPCR2$Strains %in% c("R9_2m")], na.rm = TRUE)

# Create a new column with the median for each Expe
df_qPCR2 <- df_qPCR2 %>%
  mutate(Median_Ct16S = case_when(
    Expe == "F16" ~ median_mono,
    Expe == "mono" ~ median_mono,
    TRUE ~ NA_real_  # Ensure other conditions get NA
  ))

#Normalize data (strains which has not grown based on OD600, where used to calculate the CT value for absence of growth, data were normalized accordingly)
#ITS
df_qPCR2 <- df_qPCR2 %>%
  mutate(CtITS_norm = ifelse(CtITS > Median_CtITS_F16_Mono, Median_CtITS_F16_Mono, CtITS))

#16SrRNA
df_qPCR2 <- df_qPCR2 %>%
  mutate(Ct16S_norm = ifelse(Ct16S > median_mono, median_mono, Ct16S))
```


### b) Plot
```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}
# Replace specific values in Expe column with "Communities"
df_qPCR2 <- df_qPCR2 %>%
  mutate(Expe = ifelse(Strains %in% c("BFSC", "BSC", "FSC", "F16_BSC"), "Communities", Expe))

# Calculate medians using raw -Ct values (not transformed), grouping by Strains and Expe (now with Communities)
medians <- df_qPCR2 %>%
  group_by(Strains, Expe, Expe2, Size, Kingdom) %>%
  summarise(
    med_x = median(-CtITS_norm, na.rm = TRUE),
    med_y = median(-Ct16S_norm, na.rm = TRUE),
    .groups = "drop"
  )

# Create jittered median coordinates to spread median dots and labels
set.seed(123)  # reproducibility
medians_jittered <- medians %>%
  mutate(
    med_x_jitter = med_x + runif(n(), -0.15, 0.15),
    med_y_jitter = med_y + runif(n(), -0.15, 0.15)
  )

# Change Expe label names in both dataframes to keep consistent
df_qPCR2 <- df_qPCR2 %>%
  mutate(
    Expe = case_when(
      Expe == "F16" ~ "Pairwise F16",
      Expe == "mono" ~ "Monoculture",
      TRUE ~ Expe
    )
  )

medians <- medians %>%
  mutate(
    Expe = case_when(
      Expe == "F16" ~ "Pairwise F16",
      Expe == "mono" ~ "Monoculture",
      TRUE ~ Expe
    )
  )

medians_jittered <- medians_jittered %>%
  mutate(
    Expe = case_when(
      Expe == "F16" ~ "Pairwise F16",
      Expe == "mono" ~ "Monoculture",
      TRUE ~ Expe
    )
  )


# Custom colors
Kingdom_colors <- c("Bacteria" = "#2780F5", "Fungi" = "#FFD580", "Bacteria.Fungi" = "darkgray", "NA" = "black")

# Reorder factor levels
#df_qPCR2$Expe <- factor(df_qPCR2$Expe, levels = c("Monoculture","Pairwise F16", "Communities"))
#medians$Expe <- factor(medians$Expe, levels = c("Monoculture","Pairwise F16", "Communities"))
#medians_jittered$Expe <- factor(medians_jittered$Expe, levels = c("Monoculture","Pairwise F16", "Communities"))

# Remove NA in Kingdom
df_qPCR2 <- df_qPCR2 %>% filter(!is.na(Kingdom))
medians <- medians %>% filter(!is.na(Kingdom))
medians_jittered <- medians_jittered %>% filter(!is.na(Kingdom))

#Change label
Culture_names <- c(
  'mono' = "Monoculture",
  'Pairwise' = "Pairwise (B+F16)",
  'Pseudomonas' = "Pseudomonas only")

#Plot
PlotScatterRaw <- ggplot(df_qPCR2, aes(x = -CtITS_norm, y = -Ct16S_norm, size = Size)) +
  # big white outline (halo)
  geom_point(aes(color= Kingdom), size = 4, color = "white", alpha = 1,
             position = position_jitter(width = 0.15, height = 0.15)) +
  # actual jittered points
  geom_point(aes(color= Kingdom, shape = Rep), size = 2, alpha = 0.2,
             position = position_jitter(width = 0.15, height = 0.15)) +
  
  # median dots with outline
  geom_point(data = medians_jittered,
             aes(x = med_x_jitter, y = med_y_jitter, size = Size, fill = Kingdom),
            stroke = 1, shape = 21, color = "black") +
  # labels
geom_text_repel(
  data = subset(
    medians_jittered,
    Size == "SynCom" | 
      med_y_jitter > -26 | 
      med_x_jitter > -26.2 | 
      Strains %in% c("R569_2m", "R9_2m")
  ),
  aes(x = med_x_jitter, y = med_y_jitter, label = Strains),
  inherit.aes = FALSE,
  size = 2.5,
  max.overlaps = Inf,
  segment.color = "grey50",
  segment.size = 0.3) +
  labs(x = "Fungi abundance (-Ct ITS_norm)",
       y = "Bacteria abundance (-Ct 16S_norm)",
       color = "Experiment",
       shape = "Experiment") +
facet_grid( Expe2 ~ ., labeller = as_labeller(Culture_names)) +
  scale_color_manual(values = Kingdom_colors) +
  scale_fill_manual(values = Kingdom_colors) +
  theme(strip.background = element_rect(color = "black", fill = "white", size = 1, linetype = "solid"),
        strip.text.x = element_text(size = 12, color = "black"),
        strip.text.y = element_text(size = 12, color = "black"),
        axis.text.x = element_text(angle = 90, hjust = 1)) + 
  geom_hline(yintercept = -median_mono, color = "black", linetype = "dashed", alpha = 0.2) +
  geom_vline(xintercept = -median_F16_mono, color = "black", linetype = "dashed", alpha = 0.2) +
  base_theme


PlotScatterRaw
```

# B] Analysis growth Fluorescens/OD600

## 1) With jitter
```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}
df <- read.table(file="Fluorescens-SynCom-3OMG.txt",sep="\t",dec=".",header=TRUE)

# Calculate medians grouping by relevant factors (adjust if necessary)
medians <- df %>%
  group_by(Strains, Expe, Expe2, Size, Kingdom) %>%
  summarise(
    med_x = median(Fluorescens, na.rm = TRUE),
    med_y = median(OD600, na.rm = TRUE),
    .groups = "drop"
  )

# Create jittered median coordinates
set.seed(123)
medians_jittered <- medians %>%
  mutate(
    med_x_jitter = med_x + runif(n(), -0.05, 0.05),
    med_y_jitter = med_y + runif(n(), -0.05, 0.05)
  )

# Update Expe labels if you want consistency
df <- df %>%
  mutate(
    Expe = case_when(
      Expe == "F16" ~ "Pairwise F16",
      Expe == "mono" ~ "Monoculture",
      TRUE ~ Expe
    )
  )

medians <- medians %>%
  mutate(
    Expe = case_when(
      Expe == "F16" ~ "Pairwise F16",
      Expe == "mono" ~ "Monoculture",
      TRUE ~ Expe
    )
  )

medians_jittered <- medians_jittered %>%
  mutate(
    Expe = case_when(
      Expe == "F16" ~ "Pairwise F16",
      Expe == "mono" ~ "Monoculture",
      TRUE ~ Expe
    )
  )

# Merge medians with main data
df <- left_join(df, medians, by = c("Strains", "Expe2", "Size", "Kingdom"))

# Join jittered median coords
df <- df %>%
  left_join(
    medians_jittered %>%
      dplyr::select(Strains, Expe2, med_x_jitter, med_y_jitter, Size, Kingdom),
    by = c("Strains", "Expe2", "Size", "Kingdom")
  )

# Custom colors
Kingdom_colors <- c("Bacteria" = "#2780F5", "Fungi" = "#FFD580", "Bacteria.Fungi" = "darkgray", "NA" = "black")

# Remove NA in Kingdom
df <- df %>% filter(!is.na(Kingdom))
medians <- medians %>% filter(!is.na(Kingdom))
medians_jittered <- medians_jittered %>% filter(!is.na(Kingdom))

#Change label
Culture_names <- c(
  'mono' = "Monoculture",
  'Pairwise' = "Pairwise (B+F16)",
  'Pseudomonas' = "Pseudomonas only",
  'SynCom' = "SynCom")

#Caclualte median MgCl2 for lines
median_table_MgCl2 <- df %>%
  filter(Strains == "MgCl2") %>%
  summarise(
    median_OD600 = median(OD600, na.rm = TRUE),
    median_FLUO  = median(Fluorescens,  na.rm = TRUE)
  )

#Highlith signifciantly growing strains based on OD600
# Vector of strains to highlight
highlight_strains <- c("BFSC", "BSC", "FSC", "R329", "R318", "R651")

# Plot using ggrepel for smart label placement
PlotScatterOD <- ggplot(df, aes(x = Fluorescens, y = OD600, color = Kingdom, size = Size)) +
  # raw points
  geom_point(size = 2, alpha = 0.2) +
  # median points
  geom_point(data = medians, aes(x = med_x, y = med_y, size = Size),
             fill = "white", stroke = 1) +
  # median points with outline
  geom_point(data = medians, aes(x = med_x, y = med_y, size = Size, fill = Kingdom),
             shape = 21, color = "black", stroke = 1) +
  # labels with automatic placement and connecting lines
  geom_text_repel(
  data = subset(medians, Strains %in% highlight_strains),
  aes(x = med_x, y = med_y, label = Strains),
  inherit.aes = FALSE,
  size = 2.5,
  max.overlaps = Inf,
  segment.color = "grey50",
  segment.size = 0.3
) +
  labs(
    x = "Fungal growth (Fluorescens)",
    y = "Microbial growth (OD600)",
    color = "Kingdom",
    shape = "Experiment"
  ) +
facet_grid( ~ Expe2, labeller = as_labeller(Culture_names)) +
  scale_color_manual(values = Kingdom_colors) +
  scale_fill_manual(values = Kingdom_colors) +
   theme(strip.background = element_rect(color = "black", fill = "white", size = 1, linetype = "solid"),
    strip.text.x = element_text(size = 12, color = "black"),
    strip.text.y = element_text(size = 12, color = "black"),
    axis.text.x = element_text(angle = 90, hjust = 1)
  ) +
  base_theme +
 geom_hline(yintercept = 0.0835, color = "black", linetype = "dashed", alpha = 0.2) +
 geom_vline(xintercept = 132.5, color = "black", linetype = "dashed", alpha = 0.2)
 

PlotScatterOD
```


## 2) Statistics on raw data

```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}
# Function to convert p-values to stars
pval_to_stars <- function(p) {
  ifelse(p <= 0.001, "***",
  ifelse(p <= 0.01,  "**",
  ifelse(p <= 0.05,  "*", "ns")))
}

#Load table
df <- read.table(file="Fluorescens-SynCom-3OMG.txt",sep="\t",dec=".",header=TRUE)

#Remove Controls
df <- subset(df, !Strains %in%  c("Control") & Expe2 %in% c("mono", "Pairwise"))

#Remove NA strains
df <- df[!is.na(df$Strain), ]
df <- df[!is.na(df$Expe2), ]

#Calculate MgCl2 control
mgcl2_CtFLUO <- df$Fluorescens[df$Strains == "MgCl2"]
mgcl2_CtFLUO <- mgcl2_CtFLUO[!is.na(mgcl2_CtFLUO)]
  
mgcl2_CtOD <- df$OD600[df$Strains == "MgCl2"]
mgcl2_CtOD <- mgcl2_CtOD[!is.na(mgcl2_CtOD)]
  
# Get all unique experiments
experiments <- unique(df$Expe2)

# Initialize empty results lists
results_FLUO <- list()
results_OD <- list()

# Loop over each experiment
for (exp in experiments) {
  
  df_exp <- df[df$Expe2 == exp, ]

  # Strains to compare (exclude MgCl2)
  other_strains <- setdiff(unique(df_exp$Strains), "MgCl2")
  
  for (strain in other_strains) {
    
    # Fluorescens
    strain_vals_FLUO <- df_exp$Fluorescens[df_exp$Strains == strain]
    test_FLUO <- wilcox.test(strain_vals_FLUO, mgcl2_CtFLUO, alternative = "two.sided")
    
    results_FLUO[[paste(exp, strain, sep="_")]] <- data.frame(
      Expe2 = exp,
      Strain = strain,
      W = test_FLUO$statistic,
      p_value = test_FLUO$p.value,
      n_strain = length(strain_vals_FLUO),
      n_MgCl2 = length(mgcl2_CtFLUO),
      stars = pval_to_stars(test_FLUO$p.value)
    )
    
    # OD600
    strain_vals_OD <- df_exp$OD600[df_exp$Strains == strain]
    test_OD <- wilcox.test(strain_vals_OD, mgcl2_CtOD, alternative = "two.sided")
    
    results_OD[[paste(exp, strain, sep="_")]] <- data.frame(
      Expe2 = exp,
      Strain = strain,
      W = test_OD$statistic,
      p_value = test_OD$p.value,
      n_strain = length(strain_vals_OD),
      n_MgCl2 = length(mgcl2_CtOD),
      stars = pval_to_stars(test_OD$p.value)
    )
    
  }
}

# Combine results into dataframes
results_FLUO_df <- do.call(rbind, results_FLUO)
results_OD_df <- do.call(rbind, results_OD)

#Adjust p-value
results_FLUO_df$p_adj <- p.adjust(results_FLUO_df$p_value, method = "BH")
results_OD_df$p_adj <- p.adjust(results_OD_df$p_value, method = "BH")

#Add stars
results_FLUO_df$stars_adj <- pval_to_stars(results_FLUO_df$p_adj)
results_OD_df$stars_adj <- pval_to_stars(results_OD_df$p_adj)

# View results
results_FLUO_df
results_OD_df
```
### a) Plot Mono / OD600
```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}
# Load table
df <- read.table(file="Fluorescens-SynCom-3OMG.txt",
                 sep="\t", dec=".", header=TRUE)

#Subset NA
df <- subset(df, !is.na(Strains))

# Prepare table
df_mono <- subset(df, Expe2 %in% c("mono"))

#Normalization, substract MgCl2
df_mono <- df_mono %>%
  # Step 1: calculate median OD600 of MgCl2 per Expe2
  group_by(Expe2) %>%
  mutate(median_MgCl2_OD600 = median(OD600[Strains == "MgCl2"], na.rm = TRUE)) %>%
  # Step 2: subtract median from OD600
  ungroup() %>%
  mutate(OD600 = OD600 - median_MgCl2_OD600)

#Prepare stats table 
results_OD_df$Strains <- results_OD_df$Strain
results_OD_df <- subset(results_OD_df, Expe2 %in% c("mono"))

#Preapre y position for stats labels
library(dplyr)
label_positions <- df_mono %>%
  group_by(Expe2, Strains) %>%
  summarise(y_pos = max(OD600, na.rm = TRUE) * 1.05)

results_plot_df <- left_join(results_OD_df, label_positions,
                             by = c("Expe2", "Strains"))

#Calcualte median MgCl2 for line
median_MgCl2_df <- df_mono %>%
  filter(Strains == "MgCl2") %>%
  group_by(Expe2) %>%
  summarise(median_OD = median(OD600, na.rm = TRUE))

library(dplyr)

# Calculate number of replicates per strain and experiment
n_reps_df <- df_mono %>%
  group_by(Strains) %>%
  summarise(n = n(), .groups = "drop") %>%
  # Position at the bottom of the plot
  mutate(y_pos_n = min(df_mono$OD600, na.rm = TRUE) * 0.95)  # slightly below min OD600


#Reorder plot
df_mono$Strains <- factor(df_mono$Strains, levels = c("MgCl2" ,"BFSC", "F16_BSC",  "BSC","FSC" , "R09", "R16","R147","R318","R329","R420","R431", "R565","R568","R651","R695", "F16", "F147","F212","F243"))

# Boxplot with OD600 on Y-axis
PlotBoxOD_mono <- ggplot(df_mono, aes(x = Strains, y = OD600, fill = Kingdom)) +
  
  geom_boxplot(alpha = 1, outlier.shape = NA) +
  
  # optional: add raw data points
  geom_jitter(width = 0.2, alpha = 1, size = 1) +
  
  labs(
    x = "",
    y = "Microbial growth (OD600)",
    fill = "Kingdom"
  ) +
  facet_grid(~ Expe2, labeller = as_labeller(Culture_names)) +
  
  scale_fill_manual(values = Kingdom_colors) +
  
  theme(
    strip.background = element_rect(color = "black", fill = "white", size = 1),
    strip.text.x = element_text(size = 12, color = "black"),
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  geom_hline(data = median_MgCl2_df,
             aes(yintercept = median_OD),
             color = "black",
             linetype = "dashed",
             alpha = 0.4)  +
  base_theme +
  geom_text(data = results_plot_df,
            aes(x = Strains, y = y_pos, label = stars_adj),
            inherit.aes = FALSE,
            size = 4) +
geom_text(data = n_reps_df,
            aes(x = Strains, y = y_pos_n, label = paste0("n=", n)),
            inherit.aes = FALSE,
            size = 2,
            color = "black") 

PlotBoxOD_mono

#Get median
median_OD_df <- df_mono %>%
  group_by(Expe2, Strains) %>%
  summarise(median_OD600 = median(OD600, na.rm = TRUE), .groups = "drop")

median_OD_df
```

### b) Plot Pairwise / Fluoresence
```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}
# Load table
df <- read.table(file="Fluorescens-SynCom-3OMG.txt",
                 sep="\t", dec=".", header=TRUE)

#Subset NA
df <- subset(df, !is.na(Strains))

# Prepare table
df_pair <- subset(df, Expe2 %in% c("Pairwise"))

#Normalization, substract MgCl2
df_pair <- df_pair %>%
  # Step 1: calculate median OD600 of MgCl2 per Expe2
  group_by(Expe2) %>%
  mutate(median_MgCl2_Fluorescens = median(Fluorescens[Strains == "MgCl2"], na.rm = TRUE)) %>%
  # Step 2: subtract median from Fluorescens
  ungroup() %>%
  mutate(Fluorescens = Fluorescens - median_MgCl2_Fluorescens)

#Prepare stats table 
results_FLUO_df$Strains <- results_FLUO_df$Strain
results_FLUO_df <- subset(results_FLUO_df, Expe2 %in% c("Pairwise"))

#Preapre y position for stats labels
library(dplyr)
label_positions <- df_pair %>%
  group_by(Expe2, Strains) %>%
  summarise(y_pos = max(Fluorescens, na.rm = TRUE) * 1.05)

results_plot_df <- left_join(results_FLUO_df, label_positions,
                             by = c("Expe2", "Strains"))

#Calcualte median MgCl2 for line
median_MgCl2_df <- df_pair %>%
  filter(Strains == "MgCl2") %>%
  group_by(Expe2) %>%
  summarise(median_Fluo = median(Fluorescens, na.rm = TRUE))

# Calculate number of replicates per strain and experiment
n_reps_df <- df_pair %>%
  group_by(Strains) %>%
  summarise(n = n(), .groups = "drop") %>%
  # Position at the bottom of the plot
  mutate(y_pos_n = min(df_pair$Fluorescens, na.rm = TRUE) * 0.95)  # slightly below min OD600

#Reorder plot
df_pair$Strains <- factor(df_pair$Strains, levels = c("MgCl2" ,"BFSC", "F16_BSC",  "BSC","FSC" , "R09", "R16","R147","R318","R329","R420","R431", "R565","R568","R651","R695", "F16", "F147","F212","F243"))

# Boxplot with Fluorescens on Y-axis
PlotBoxfluo_pair <- ggplot(df_pair, aes(x = Strains, y = Fluorescens, fill = Kingdom)) +
  geom_boxplot(alpha = 1, outlier.shape = NA) +
  # optional: add raw data points
  geom_jitter(width = 1, alpha = 0.3, size = 1) +
  
  labs(
    x = "",
    y = "Microbial growth (Fluorescens)",
    fill = "Kingdom"
  ) +
  facet_grid(~ Expe2, labeller = as_labeller(Culture_names)) +
  scale_fill_manual(values = Kingdom_colors) +
  theme(
    strip.background = element_rect(color = "black", fill = "white", size = 1),
    strip.text.x = element_text(size = 12, color = "black"),
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  geom_hline(data = median_MgCl2_df,
             aes(yintercept = median_Fluo),
             color = "black",
             linetype = "dashed",
             alpha = 0.4)  +
  base_theme +
  geom_text(data = results_plot_df,
            aes(x = Strains, y = y_pos, label = stars_adj),
            inherit.aes = FALSE,
            size = 4) +
geom_text(data = n_reps_df,
            aes(x = Strains, y = y_pos_n, label = paste0("n=", n)),
            inherit.aes = FALSE,
            size = 2,
            color = "black") 

PlotBoxfluo_pair
```

### c) Plot just Pseudomonas
```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}
#Prepare table
df_Pseudomonas <- subset(df, Expe2 %in% c("Pseudomonas"))
medians_Pseudeomonas <- subset(medians, Expe2 %in% c("Pseudomonas"))

# Plot using ggrepel for smart label placement
PlotScatterOD_Pseudomonas <- ggplot(df_Pseudomonas, aes(x = Fluorescens, y = OD600, color = Kingdom, size = Size)) +
  # raw points
  geom_point(size = 2, alpha = 0.2) +
  # median points
  geom_point(data = medians_Pseudeomonas, aes(x = med_x, y = med_y, size = Size),
             fill = "white", stroke = 1) +
  # median points with outline
  geom_point(data = medians_Pseudeomonas, aes(x = med_x, y = med_y, size = Size, fill = Kingdom),
             shape = 21, color = "black", stroke = 1) +
  # labels with automatic placement and connecting lines
  geom_text_repel(
  data = subset(medians_Pseudeomonas, med_y > 0.10 | med_x > 150),
  aes(x = med_x, y = med_y, label = Strains),
  inherit.aes = FALSE,
  size = 2.5,
  max.overlaps = Inf,
  segment.color = "grey50",
  segment.size = 0.3) +
  labs(
    x = "Fungal growth (Fluorescens)",
    y = "Microbial growth (OD600)",
    color = "Kingdom",
    shape = "Experiment"
  ) +
facet_grid( ~ Expe2, labeller = as_labeller(Culture_names)) +
  scale_color_manual(values = Kingdom_colors) +
  scale_fill_manual(values = Kingdom_colors) +
   theme(strip.background = element_rect(color = "black", fill = "white", size = 1, linetype = "solid"),
    strip.text.x = element_text(size = 12, color = "black"),
    strip.text.y = element_text(size = 12, color = "black"),
    axis.text.x = element_text(angle = 90, hjust = 1)
  ) +
  geom_hline(yintercept = 0.10, color = "black", linetype = "dashed", alpha = 0.2) +
  geom_vline(xintercept = 150, color = "black", linetype = "dashed", alpha = 0.2) +
  base_theme

PlotScatterOD_Pseudomonas
```
# C] Identity pqqDH

## 1) Prepare data
```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}
alignment <- readAAStringSet("Pseudo_pqqDH.al.fasta")

IdMatrix <- DistanceMatrix(alignment, includeTerminalGaps=FALSE)
percent_identity <- data.frame(100*(1-IdMatrix))

#Prepare table
percent_identity$Strains2 <- colnames(percent_identity)
percent_identity$Strains2 <- percent_identity$Strains2 |> 
  sub("_gene.*", "", x = _) |> 
  sub("^Root", "R", x = _)

# Compute median CtITS_norm per Strain
median_per_strain <- df_qPCR2 %>%
  group_by(Strains) %>%
  summarise(median_CtITS_norm = median(CtITS_norm, na.rm = TRUE),
            .groups = "drop")

# Compute median Ct16S_norm per Strain
median_Ct16S_per_strain <- df_qPCR2 %>%
  group_by(Strains) %>%
  summarise(median_Ct16S_norm = median(Ct16S_norm, na.rm = TRUE),
            .groups = "drop")

# Combine ITS and 16S per strain
strain_stats <- median_per_strain %>%
  left_join(median_Ct16S_per_strain, by = "Strains")

# Separate Strains into Strains2 and Expe2
strain_stats <- strain_stats %>%
  separate(Strains, into = c("Strains2", "Expe2"), sep = "_", remove = FALSE)

# Merge with percent_identity metadata
df_merged <- left_join(strain_stats, percent_identity, by = "Strains2")
```
## 2) Plot

```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}
library(dplyr)
library(ggplot2)
library(ggrepel)

df_expe2 <- df_merged %>% 
  filter(Expe2 == "2")

glm_expe2 <- glm(
  Root329_gene_5511 ~ I(-median_CtITS_norm),
  data = df_expe2,
  family = gaussian()
)

glm_summary_expe2 <- summary(glm_expe2)

p_value_expe2 <- coef(glm_summary_expe2)[2, 4]
r2_expe2 <- 1 - (glm_expe2$deviance / glm_expe2$null.deviance)

glm_label_expe2 <- data.frame(
  x_pos = min(-df_expe2$median_CtITS_norm + 2, na.rm = TRUE),
  y_pos = max(df_expe2$Root329_gene_5511, na.rm = TRUE),
  label = paste0("R² = ", round(r2_expe2, 2),
                 ", P = ", signif(p_value_expe2, 2))
)

Plot_corr_ITS <- ggplot(df_expe2, 
                        aes(x = -median_CtITS_norm, 
                            y = Root329_gene_5511)) +
  geom_point(color = "darkgray", size = 2, alpha = 0.8) +
  geom_smooth(method = "glm",
              method.args = list(family = "gaussian"),
              se = FALSE,
              color = "black",
              linewidth = 1) +
  labs(
    x = "Fungal growth (median_CtITS_norm)",
    y = "pqqDH Identity to R329 (%)",
    title = ""
  ) +
  geom_text(
    data = glm_label_expe2,
    aes(x = x_pos, y = y_pos, label = label),
    hjust = 1, vjust = 1,
    size = 4,
    color = "black",
    inherit.aes = FALSE
  ) +
  geom_text_repel(
    aes(label = Strains),
    size = 3,
    max.overlaps = 20
  ) +
  theme(legend.position = "right") +
  base_theme


df_expe2m <- df_merged %>% 
  filter(Expe2 == "2m")

glm_expe2m <- glm(
  Root329_gene_5511 ~ I(-median_Ct16S_norm),
  data = df_expe2m,
  family = gaussian()
)

glm_summary_expe2m <- summary(glm_expe2m)

p_value_expe2m <- coef(glm_summary_expe2m)[2, 4]
r2_expe2m <- 1 - (glm_expe2m$deviance / glm_expe2m$null.deviance)

glm_label_expe2m <- data.frame(
  x_pos = min(-df_expe2m$median_Ct16S_norm + 2, na.rm = TRUE),
  y_pos = max(df_expe2m$Root329_gene_5511, na.rm = TRUE),
  label = paste0("R² = ", round(r2_expe2m, 2),
                 ", P = ", signif(p_value_expe2m, 2))
)

Plot_corr_16S <- ggplot(df_expe2m, 
                        aes(x = -median_Ct16S_norm, 
                            y = Root329_gene_5511)) +
  geom_point(color = "#2780F5", size = 2, alpha = 0.8) +
  geom_smooth(method = "glm",
              method.args = list(family = "gaussian"),
              se = FALSE,
              color = "black",
              linewidth = 1) +
  labs(
    x = "median_Ct16S_norm",
    y = "Root329_gene_5511",
    title = "Expe2 = 2m"
  ) +
  geom_text(
    data = glm_label_expe2m,
    aes(x = x_pos, y = y_pos, label = label),
    hjust = 1, vjust = 1,
    size = 4,
    color = "black",
    inherit.aes = FALSE
  ) +
  geom_text_repel(
    aes(label = Strains),
    size = 3,
    max.overlaps = 20
  ) +
  theme(legend.position = "right") +
  base_theme

Plot_corr_ITS
Plot_corr_16S
```

## 3) Tree

```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}
# Read the tree
tree <- read.tree("Pseudo_pqqDH.al.fasta.treefile")

# Replace "TipName" with the exact name of the tip you want to root on
tree_rooted <- root(tree, outgroup = "Root1280_gene_4128", resolve.root = TRUE)

# Plot tree
p_tree <- ggtree(tree_rooted, layout = "rectangular") +
  geom_tiplab(size = 3) +  # align tips, slight offset
  theme_tree2() +                                     # aligns tips with x-axis
  ggtitle("Pseudo_pqqDH Phylogenetic Tree") +
  theme(plot.title = element_text(hjust = 0.5))      # center title

p_tree
```

# D] Final plots

```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}

Final_plot_jitter <- plot_grid(PlotScatterRaw, PlotScatterOD, rel_heights = c(2,2), rel_widths = c(1,1), ncol=2, nrow=1, align = 'hv')
Final_plot_jitter

Final_plot_supp_S10 <- plot_grid(PlotBoxOD_mono, PlotBoxfluo_pair, rel_heights = c(2,2), rel_widths = c(1,1), ncol=2, nrow=1, align = 'hv')
Final_plot_supp_S10
```

---
END
---