---
title: "Decoy sugars hijack fungal carbon sensing and prevent plant disease"
author: "Guillaume Chesneau & Dorcas Lin"
date: "26-05-2026"
output: html_document
---

# Before you start

This script has been made on R 4.2.2 (and Rstudio 2023.09.1+494)

```{r R version}
R.version
```

```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}

library(rstatix)  # For statistical analysis
library(ggpubr)   # For creating publication-ready plots using ggplot2
library(dunn.test)
library(ggthemes)
library(ggtext)
library(tidyr)
library(markdown)
library(ggplot2)
library(dplyr)
library(broom)
library(cowplot)
library(purrr)
library(FSA)
library(multcompView)
library(rcompanion)

color_palette <- c(
  "Water" = "#949494",
  "3OMG"  = "#7FC29B"
)
theme_set(theme_bw())

# Prepare base theme
base_theme <- theme_bw() +
  theme(
    strip.background = element_rect(color = "white", fill = "white"),
    strip.text.x = element_markdown(size = 10, color = "black", hjust = 0.5, vjust = 0.5),
    legend.title = element_text(size = 11),
    axis.title.y = element_text(size =11),
    legend.text = element_text(size = 10),
    axis.text.x = element_text(size = 10, color = "black", angle = 90, hjust = 1),
    panel.border = element_rect(color = "black", fill = NA))

#Set the levels of the experiment
Microbes_levels<-c("MF","BFSC","BSC","FSC","F16","F44","F64","F96","F100","F117","F147","F212","F243","Pu","Rs","Vd","Bs","AMF")

Microbes_Primers_levels<-c("BFSC - ITS","BFSC - Primer_16S","BSC - Primer_16S","FSC - ITS", "MF" ,"F16 - ITS", "F117 - ITS","Vd - ITS", "F96 - ITS", "F147 - ITS", "F212 - ITS", "F44 - ITS", "F243 - ITS", "F100 - ITS", "F64 - ITS", "Pu - ITS1", "Rs - RS",  "AMF - ITS", "Bs - ITS")


Kingdom_levels <- c("MiniSynCom", "Fungi","Oomycetes","Bacteria")

Kingdom <- tibble(Microbes = Microbes_levels) %>%
  mutate(Kingdom = dplyr::case_when(
    Microbes %in% c("BFSC","BSC","FSC") ~ "MiniSynCom",
    Microbes %in% c("F16","F44","F64","F96","F100","F117","F147","F212","F243","Vd","Bs","AMF") ~ "Fungi",
    Microbes == "Pu" ~ "Oomycetes",
    Microbes == "Rs" ~ "Bacteria",
    TRUE ~ NA_character_
  ))

Treatment_levels<-c ("Water","3OMG")
labels <- c(
  'MF' = "Microbe free",
  'BFSC - ITS'="Bacterial and Fungal<br>SynCom - *ITS*",
  'BFSC - Primer_16S'="Bacterial and Fungal<br>SynCom - *16S*",
  'BSC - Primer_16S'="Bacterial<br>SynCom - *16S*",
  'FSC - ITS' = "Fungal<br>SynCom - *ITS*",
  'F16 - ITS' = "*Plecto.*<br>*cucumerina - ITS*",
  'F44 - ITS' = "*Fusarium*<br>*tricinctum - ITS*",
  'F64 - ITS' = "*Alternaria*<br>*alternata - ITS*",
  'F96 - ITS' = "*Rhizoctonia*<br>*solani - ITS*",
  'F100 - ITS' = "*Boeremia*<br>*exigua - ITS*",
  'F117 - ITS' = "*Plecto.*<br>*plurivora - ITS*",
  'F147 - ITS' = "*Dactylo.*<br>*macrodidyma - ITS*",
  'F212 - ITS' = "*Fusarium*<br>*oxysporum - ITS*",
  'F243 - ITS' = "*Dendryphion*<br>*nanum - ITS*",
  'Pu - ITS1' = "*Globispor.*<br>*ultimum - ITS1*",
  'Rs - RS' = "*Ralstonia*<br>*solanacear. - UGD-up*",
  'Bs - ITS' = "*Bipolaris*<br>*sorokiniana - ITS*",
  'NoAMF' = "Microbe free",
  'Vd - ITS' = "*Verticilium*<br> *dahliae - ITS*",
  'AMF - ITS' = "*Rhizophagus*<br>*irregularis - ITS*",
  'Fungi' = "Fungi",
  'Oomycetes' = "Oomycetes",
  'Bacteria' = "Bacteria",
  'Arabidopsis_thaliana' = "Arabidopsis thaliana"
)

labels2 <- c(
  'MF' = "Microbe free",
  'BFSC'="Bacterial and Fungal<br>SynCom - *ITS*",
  'BFSC'="Bacterial and Fungal<br>SynCom - *16S*",
  'BSC'="Bacterial<br>SynCom - *16S*",
  'FSC' = "Fungal<br>SynCom - *ITS*",
  'F16' = "*Plectosphaerella*<br>*cucumerina*",
  'F44' = "*Fusarium*<br>*tricinctum*",
  'F64' = "*Alternaria*<br>*alternata*",
  'F96' = "*Rhizoctonia*<br>*solani*",
  'F100' = "*Boeremia*<br>*exigua*",
  'F117' = "*Plectosphaerella*<br>*plurivora*",
  'F147' = "*Dactylonectria*<br>*macrodidyma*",
  'F212' = "*Fusarium*<br>*oxysporum*",
  'F243' = "*Dendryphion*<br>*nanum*",
  'Pu' = "*Globisporangium*<br>*ultimum*",
  'Rs' = "*Ralstonia*<br>*solanacearum*",
  'Bs' = "*Bipolaris*<br>*sorokiniana*",
  'Bc' = "*Botrytis*<br>*cinerea*",
  'NoAMF' = "Microbe free",
  'Vd' = "*Verticilium*<br> *dahliae*",
  'AMF' = "*Rhizophagus*<br>*irregularis*",
  'Fungi' = "Fungi",
  'Oomycetes' = "Oomycetes",
  'Bacteria' = "Bacteria",
  'Flowpots' = "Flowpots",
  'MS_Agar' = "MS Agar"
)
```

# A] Colonization

## 1. Prepare Data

```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}
#Open dataset
df <- read.table(file="Df_total_Colonization.txt",sep="\t",dec = ".",header=TRUE, row.names=1)

#Remove Botrytis
df <- subset(df, !Microbes %in% c("Bc"))
```

## 2. Subset Plant / Matrix

```{r}
#Subset the data to separate Soil and Root
Matrix_data2 <- subset(df, Element %in% c("Matrix"))
Plant_data2 <- subset(df, Element %in% c("Plant"))
```

## 3.  Arabidopsis

### a) - Colonization Matrixes
```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}
#Subset At
Matrix_data <- subset(Matrix_data2, Plant %in% c("Arabidopsis_thaliana"))
Matrix_data <- subset(Matrix_data, !Microbes %in% c("Vd", "Bs", "AMF"))

# Calculate adjusted CT
Matrix_data$CT[is.na(Matrix_data$CT)]<-40

# Remove Glucose
Matrix_data <- subset(Matrix_data, !Treatment %in% c("Glucose"))

# Clean soil data: remove invalid combinations
Matrix_data_cleaned <- Matrix_data %>%
  filter(
    (Primers == "Primer_16S" & Microbes == c("BSC", "BFSC")) |
    (Primers == "RS" & Microbes == c("Rs")) |
    (Primers == "ITS1" & Microbes == c("Pu")) |
    (Primers == "ITS" & !(Microbes %in% c("BSC", "Rs", "Pu", "MF")))
  )

# Create concat variable
Matrix_data_cleaned$concat <- paste0(Matrix_data_cleaned$Microbes, "/", Matrix_data_cleaned$Primers,"/",  Matrix_data_cleaned$Plant, "/", Matrix_data_cleaned$System)

# Stats: Wilcoxon test by group
get_signif_results <- function(data, x_start = 1, x_end = 2) {
  
  # Optional: see counts
  data %>%
    count(concat, Treatment) %>%
    print()
  
  # Nest data by concat
  data %>%
    group_by(concat) %>%
    nest() %>%
    mutate(
      test_result = map(data, ~ tryCatch(
        t.test(CT ~ Treatment, data = .x),
        error = function(e) NA_real_
      )),
      p_value = map_dbl(test_result, ~ if(is.list(.x)) .x$p.value else NA_real_),
      y_pos = map_dbl(data, ~ max(.x$CT, na.rm = TRUE) + 0.5),
      x_start = x_start,
      x_end = x_end,
      significance = case_when(
        is.na(p_value) ~ "ns",
        p_value < 0.001 ~ "***",
        p_value < 0.01 ~ "**",
        p_value < 0.05 ~ "*",
        TRUE ~ "ns"
      )) %>%
    select(concat, p_value, y_pos, x_start, x_end, significance) %>%
    unnest(cols = c())
}

#Run Stats
Stats_Matrix <- get_signif_results(Matrix_data_cleaned)

#Add Group Treatment
Stats_Matrix$group1 <- c("Water")
Stats_Matrix$group2 <- c("3OMG")

#Remake columns 
Stats_Matrix <- Stats_Matrix %>%
  separate(
    concat,
    into = c("Microbes", "Primers", "Plant", "System"),
    sep = "/",
    remove = FALSE  # keeps the original concat column
  )

# Merge stats with soil data
Matrix_data_cleaned_Stats <- Matrix_data_cleaned %>%
  left_join(Stats_Matrix, by = c("Microbes", "Primers", "Plant", "System"))

# Prepare the stats table for plotting
Matrix_data_cleaned_Stats <- Matrix_data_cleaned_Stats %>%
  mutate(
    # Only show significant p-values (< 0.05), rounded to 3 decimals
    label_text = ifelse(
      p_value < 0.05,
      paste0("P = ", round(p_value,5)),
      NA_character_
    )
  )

#Remove MiniSynCom
Matrix_data_cleaned_Stats <- subset(Matrix_data_cleaned_Stats, !Experiment %in% c("MiniSC"))

#Order
Matrix_data_cleaned_Stats$Treatment <- factor(Matrix_data_cleaned_Stats$Treatment, levels = c ("Water", "Glucose", "3OMG"))

#Number of samples
rep_counts <- Matrix_data_cleaned_Stats %>%
  group_by(System, Microbes, Treatment) %>%
  summarise(n = n(), .groups = "drop")

# Reorder the 'System' and 'Microbes' factors
Matrix_data_cleaned_Stats$Microbes <- factor(Matrix_data_cleaned_Stats$Microbes, levels = c( "MF" ,"F16", "F117", "F96", "F147", "F212", "F44", "F243", "F100", "F64", "Pu", "Rs"))  # desired order

rep_counts$Microbes <- factor(rep_counts$Microbes, levels = c( "MF" ,"F16", "F117", "F96", "F147", "F212", "F44", "F243", "F100", "F64", "Pu", "Rs"))  # desired order

# Plot
Plot_Matrix_at <- ggplot(Matrix_data_cleaned_Stats, aes(x = Treatment, y = -CT, fill = Treatment)) + 
  geom_boxplot(alpha = 1, outlier.shape = NA) +
  geom_jitter(alpha = 0.6, width = 0.2) +
  ylab(expression("Matrix Colonization (-Ct)")) +
  xlab("") +
  facet_grid(System ~Microbes , space = "free", scales = "free", labeller = as_labeller(labels2)) +
    scale_fill_manual(values = color_palette) +
  labs(shape = "Biological Replicates") +
  base_theme +
# Add significant p-values
  stat_pvalue_manual(
    ok <- Matrix_data_cleaned_Stats %>%
  filter(Treatment != "3OMG") %>%                       # 🔹 remove 3OMG
  distinct(System, Microbes, Treatment, label_text,     # 🔹 avoid duplicates
           p_value, y_pos, .keep_all = TRUE),
    label = "label_text",
    y.position = -10,
    tip.length = 0.02,
    hide.ns = TRUE,
    size = 3
  )+
  # Add replicate counts
  geom_text(
    data = rep_counts,
    aes(x = Treatment, y = -40, 
        label = paste0("n = ", n)),
    inherit.aes = FALSE,
    size = 3
  )  + theme(
    strip.background = element_rect(color = "black", fill = "white"))

Plot_Matrix_at

#Save The table for HeatMap
Matrix_data_cleaned_Stats_At <- Matrix_data_cleaned_Stats
```

### b) - Colonization Plant

```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}
#Subset At
Plant_data <- subset(Plant_data2, Plant %in% c("Arabidopsis_thaliana"))
Plant_data <- subset(Plant_data, !Microbes %in% c("Vd", "Bs", "AMF"))

#Put non detected Microbial primers as a value of 40
Plant_data$UBQ10[is.na(Plant_data$UBQ10)] <- 40
Plant_data$CT[is.na(Plant_data$CT)] <- 40

#Remove when plants are not detected
Plant_data <- subset(Plant_data, !UBQ10 %in% c(40))


# Remove Glucose
Plant_data <- subset(Plant_data, !Treatment %in% c("Glucose"))

# Clean soil data: remove invalid combinations
Plant_data_cleaned <- Plant_data %>%
  filter(
    (Primers == "Primer_16S" & Microbes == c("BSC", "BFSC")) |
    (Primers == "RS" & Microbes == c("Rs")) |
    (Primers == "ITS1" & Microbes == c("Pu")) |
    (Primers == "ITS" & !(Microbes %in% c("BSC", "Rs", "Pu", "MF")))
  )

# Calculate the ratio for each row (2^Dekta CT)
Plant_data_cleaned$LogDeltaCT <- log(2^(-(Plant_data_cleaned$CT - Plant_data_cleaned$UBQ10)))
Plant_data_cleaned$DeltaCT <- 2^(-(Plant_data_cleaned$CT - Plant_data_cleaned$UBQ10))

  
# Create concat variable
Plant_data_cleaned$concat <- paste0(Plant_data_cleaned$Microbes, "/", Plant_data_cleaned$Primers,"/",  Plant_data_cleaned$Plant, "/", Plant_data_cleaned$System)


# Stats: Wilcoxon test by group
get_signif_results <- function(data, x_start = 1, x_end = 2) {
  
  # Optional: see counts
  data %>%
    count(concat, Treatment) %>%
    print()
  
  # Nest data by concat
  data %>%
    group_by(concat) %>%
    nest() %>%
    mutate(
      test_result = map(data, ~ tryCatch(
        t.test(LogDeltaCT ~ Treatment, data = .x),
        error = function(e) NA_real_
      )),
      p_value = map_dbl(test_result, ~ if(is.list(.x)) .x$p.value else NA_real_),
      y_pos = map_dbl(data, ~ max(.x$DeltaCT, na.rm = TRUE) + 0.5),
      x_start = x_start,
      x_end = x_end,
      significance = case_when(
        is.na(p_value) ~ "ns",
        p_value < 0.001 ~ "***",
        p_value < 0.01 ~ "**",
        p_value < 0.05 ~ "*",
        TRUE ~ "ns"
      )) %>%
    select(concat, p_value, y_pos, x_start, x_end, significance) %>%
    unnest(cols = c())
}

#Run Stats
Stats_Plant <- get_signif_results(Plant_data_cleaned)

#Add Group Treatment
Stats_Plant$group1 <- c("Water")
Stats_Plant$group2 <- c("3OMG")

#Remake columns 
Stats_Plant <- Stats_Plant %>%
  separate(
    concat,
    into = c("Microbes", "Primers", "Plant", "System"),
    sep = "/",
    remove = FALSE  # keeps the original concat column
  )



# Merge stats with soil data
Plant_data_cleaned_Stats <- Plant_data_cleaned %>%
  left_join(Stats_Plant, by = c("Microbes", "Primers", "Plant", "System"))

# Prepare the stats table for plotting
Plant_data_cleaned_Stats <- Plant_data_cleaned_Stats %>%
  mutate(
    # Only show significant p-values (< 0.05), rounded to 3 decimals
    label_text = ifelse(
      p_value < 0.05,
      paste0("P = ", round(p_value,5)),
      NA_character_
    )
  )



#Remove MiniSynCom
Plant_data_cleaned_Stats <- subset(Plant_data_cleaned_Stats, !Experiment %in% c("MiniSC"))

#Order
Plant_data_cleaned_Stats$Treatment <- factor(Plant_data_cleaned_Stats$Treatment, levels = c ("Water", "Glucose", "3OMG"))

#Number of samples
rep_counts <- Plant_data_cleaned_Stats %>%
  group_by(System, Microbes, Treatment) %>%
  summarise(n = n(), .groups = "drop")

# Reorder the 'System' and 'Microbes' factors
Plant_data_cleaned_Stats$Microbes <- factor(Plant_data_cleaned_Stats$Microbes, levels = c( "MF" ,"F16", "F117", "F96", "F147", "F212", "F44", "F243", "F100", "F64", "Pu", "Rs"))  # desired order

rep_counts$Microbes <- factor(rep_counts$Microbes, levels = c( "MF" ,"F16", "F117", "F96", "F147", "F212", "F44", "F243", "F100", "F64", "Pu", "Rs"))  # desired order

# Plot
Plot_PLant_at <- ggplot(Plant_data_cleaned_Stats, aes(x = Treatment, y = LogDeltaCT, fill = Treatment)) + 
  geom_boxplot(alpha = 1, outlier.shape = NA) +
  geom_jitter(alpha = 0.6, width = 0.2) +
  ylab(expression("Root Colonization " ~ 2^{Delta*CT} ~ italic(AtUBQ10))) +
  xlab("") +
  facet_grid(System~Microbes, space = "free", scales = "free", labeller = as_labeller(labels2)) +
    scale_fill_manual(values = color_palette) +
  labs(shape = "Biological Replicates") +
  base_theme +
  # Add significant p-values
  stat_pvalue_manual(
    ok <- Plant_data_cleaned_Stats %>%
  filter(Treatment != "3OMG") %>%                       # 🔹 remove 3OMG
  distinct(System, Microbes, Treatment, label_text,     # 🔹 avoid duplicates
           p_value, y_pos, .keep_all = TRUE),
    label = "label_text",
    y.position = 12,
    tip.length = 0.02,
    hide.ns = TRUE,
    size = 3
  )+
  # Add replicate counts
  geom_text(
    data = rep_counts,
    aes(x = Treatment, y = min(Plant_data_cleaned_Stats$LogDeltaCT, na.rm = TRUE) - 0.5, 
        label = paste0("n = ", n)),
    inherit.aes = FALSE,
    size = 3
  ) + theme(
    strip.background = element_rect(color = "black", fill = "white"))

Plot_PLant_at

#Save The table for HeatMap
Plant_data_cleaned_Stats_At <- Plant_data_cleaned_Stats
```
## 4.  Tomato

### a) - Colonization Matrixes
```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}
#Subset At
Matrix_data <- subset(Matrix_data2, Plant %in% c("Micro_Tomato"))

# Calculate adjusted CT
Matrix_data$CT[is.na(Matrix_data$CT)]<-40

# Remove Glucose
Matrix_data <- subset(Matrix_data, !Treatment %in% c("Glucose"))

# Clean soil data: remove invalid combinations
Matrix_data_cleaned <- Matrix_data %>%
  filter(
    (Primers == "Primer_16S" & Microbes == c("BSC", "BFSC")) |
    (Primers == "RS" & Microbes == c("Rs")) |
    (Primers == "ITS1" & Microbes == c("Pu")) |
    (Primers == "ITS" & !(Microbes %in% c("BSC", "Rs", "Pu", "MF")))
  )

# Create concat variable
Matrix_data_cleaned$concat <- paste0(Matrix_data_cleaned$Microbes, "/", Matrix_data_cleaned$Primers,"/",  Matrix_data_cleaned$Plant, "/", Matrix_data_cleaned$System)


# Stats: Wilcoxon test by group
get_signif_results <- function(data, x_start = 1, x_end = 2) {
  
  # Optional: see counts
  data %>%
    count(concat, Treatment) %>%
    print()
  
  # Nest data by concat
  data %>%
    group_by(concat) %>%
    nest() %>%
    mutate(
      test_result = map(data, ~ tryCatch(
        t.test(CT ~ Treatment, data = .x),
        error = function(e) NA_real_
      )),
      p_value = map_dbl(test_result, ~ if(is.list(.x)) .x$p.value else NA_real_),
      y_pos = map_dbl(data, ~ max(.x$CT, na.rm = TRUE) + 0.5),
      x_start = x_start,
      x_end = x_end,
      significance = case_when(
        is.na(p_value) ~ "ns",
        p_value < 0.001 ~ "***",
        p_value < 0.01 ~ "**",
        p_value < 0.05 ~ "*",
        TRUE ~ "ns"
      )) %>%
    select(concat, p_value, y_pos, x_start, x_end, significance) %>%
    unnest(cols = c())
}

#Run Stats
Stats_Matrix <- get_signif_results(Matrix_data_cleaned)

#Add Group Treatment
Stats_Matrix$group1 <- c("Water")
Stats_Matrix$group2 <- c("3OMG")

#Remake columns 
Stats_Matrix <- Stats_Matrix %>%
  separate(
    concat,
    into = c("Microbes", "Primers", "Plant", "System"),
    sep = "/",
    remove = FALSE  # keeps the original concat column
  )



# Merge stats with soil data
Matrix_data_cleaned_Stats <- Matrix_data_cleaned %>%
  left_join(Stats_Matrix, by = c("Microbes", "Primers", "Plant", "System"))

# Prepare the stats table for plotting
Matrix_data_cleaned_Stats <- Matrix_data_cleaned_Stats %>%
  mutate(
    # Only show significant p-values (< 0.05), rounded to 3 decimals
    label_text = ifelse(
      p_value < 0.05,
      paste0("P = ", round(p_value,5)),
      NA_character_
    )
  )

#Remove MiniSynCom
Matrix_data_cleaned_Stats <- subset(Matrix_data_cleaned_Stats, !Experiment %in% c("MiniSC"))

#Order
Matrix_data_cleaned_Stats$Treatment <- factor(Matrix_data_cleaned_Stats$Treatment, levels = c ("Water", "Glucose", "3OMG"))

#Number of samples
rep_counts <- Matrix_data_cleaned_Stats %>%
  group_by(System, Microbes, Treatment) %>%
  summarise(n = n(), .groups = "drop")

# Reorder the 'System' and 'Microbes' factors
Matrix_data_cleaned_Stats$Microbes <- factor(Matrix_data_cleaned_Stats$Microbes, levels = c("MF" ,"F16", "Vd", "F96", "AMF"))  # desired order

rep_counts$Microbes <- factor(rep_counts$Microbes, levels = c("MF" ,"F16", "Vd", "F96", "AMF"))  # desired order


# Plot
Plot_Matrix_T <- ggplot(Matrix_data_cleaned_Stats, aes(x = Treatment, y = -CT, fill = Treatment)) + 
  geom_boxplot(alpha = 1, outlier.shape = NA) +
  geom_jitter(alpha = 0.6, width = 0.2) +
  ylab(expression("Matrix Colonization (-Ct)")) +
  xlab("") +
  facet_grid(System ~Microbes , space = "free", scales = "free", labeller = as_labeller(labels2)) +
    scale_fill_manual(values = color_palette) +
  labs(shape = "Biological Replicates") +
  base_theme +
  # Add significant p-values
  stat_pvalue_manual(
    ok <- Matrix_data_cleaned_Stats %>%
  filter(Treatment != "3OMG") %>%                       # 🔹 remove 3OMG
  distinct(System, Microbes, Treatment, label_text,     # 🔹 avoid duplicates
           p_value, y_pos, .keep_all = TRUE),
    label = "label_text",
    y.position = -10,
    tip.length = 0.02,
    hide.ns = TRUE,
    size = 3
  )+
  # Add replicate counts
  geom_text(
    data = rep_counts,
    aes(x = Treatment, y = -40, 
        label = paste0("n = ", n)),
    inherit.aes = FALSE,
    size = 3
  ) + theme(
    strip.background = element_rect(color = "black", fill = "white"))
Plot_Matrix_T

#Save The table for HeatMap
Matrix_data_cleaned_Stats_MT <- Matrix_data_cleaned_Stats
```


### b) - Colonization Plant

```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}
#Subset At
Plant_data <- subset(Plant_data2, Plant %in% c("Micro_Tomato"))

#Put non detected Microbial primers as a value of 40
Plant_data$UBQ10[is.na(Plant_data$UBQ10)] <- 40
Plant_data$CT[is.na(Plant_data$CT)] <- 40

#Remove when plants are not detected
Plant_data <- subset(Plant_data, !UBQ10 %in% c(40))


# Remove Glucose
Plant_data <- subset(Plant_data, !Treatment %in% c("Glucose"))

# Clean soil data: remove invalid combinations
Plant_data_cleaned <- Plant_data %>%
  filter(
    (Primers == "Primer_16S" & Microbes == c("BSC", "BFSC")) |
    (Primers == "RS" & Microbes == c("Rs")) |
    (Primers == "ITS1" & Microbes == c("Pu")) |
    (Primers == "ITS" & !(Microbes %in% c("BSC", "Rs", "Pu", "MF")))
  )

# Calculate the ratio for each row (2^Dekta CT)
Plant_data_cleaned$LogDeltaCT <- log(2^(-(Plant_data_cleaned$CT - Plant_data_cleaned$UBQ10)))
Plant_data_cleaned$DeltaCT <- 2^(-(Plant_data_cleaned$CT - Plant_data_cleaned$UBQ10))

  
# Create concat variable
Plant_data_cleaned$concat <- paste0(Plant_data_cleaned$Microbes, "/", Plant_data_cleaned$Primers,"/",  Plant_data_cleaned$Plant, "/", Plant_data_cleaned$System)


# Stats: Wilcoxon test by group

get_signif_results <- function(data, x_start = 1, x_end = 2) {
  
  # Optional: see counts
  data %>%
    count(concat, Treatment) %>%
    print()
  
  # Nest data by concat
  data %>%
    group_by(concat) %>%
    nest() %>%
    mutate(
      test_result = map(data, ~ tryCatch(
        t.test(LogDeltaCT ~ Treatment, data = .x),
        error = function(e) NA_real_
      )),
      p_value = map_dbl(test_result, ~ if(is.list(.x)) .x$p.value else NA_real_),
      y_pos = map_dbl(data, ~ max(.x$DeltaCT, na.rm = TRUE) + 0.5),
      x_start = x_start,
      x_end = x_end,
      significance = case_when(
        is.na(p_value) ~ "ns",
        p_value < 0.001 ~ "***",
        p_value < 0.01 ~ "**",
        p_value < 0.05 ~ "*",
        TRUE ~ "ns"
      )) %>%
    select(concat, p_value, y_pos, x_start, x_end, significance) %>%
    unnest(cols = c())
}

#Run Stats
Stats_Plant <- get_signif_results(Plant_data_cleaned)

#Add Group Treatment
Stats_Plant$group1 <- c("Water")
Stats_Plant$group2 <- c("3OMG")

#Remake columns 
Stats_Plant <- Stats_Plant %>%
  separate(
    concat,
    into = c("Microbes", "Primers", "Plant", "System"),
    sep = "/",
    remove = FALSE  # keeps the original concat column
  )



# Merge stats with soil data
Plant_data_cleaned_Stats <- Plant_data_cleaned %>%
  left_join(Stats_Plant, by = c("Microbes", "Primers", "Plant", "System"))

# Prepare the stats table for plotting
Plant_data_cleaned_Stats <- Plant_data_cleaned_Stats %>%
  mutate(
    # Only show significant p-values (< 0.05), rounded to 3 decimals
    label_text = ifelse(
      p_value < 0.05,
      paste0("P = ", round(p_value,5)),
      NA_character_
    )
  )



#Remove MiniSynCom
Plant_data_cleaned_Stats <- subset(Plant_data_cleaned_Stats, !Experiment %in% c("MiniSC"))

#Order
Plant_data_cleaned_Stats$Treatment <- factor(Plant_data_cleaned_Stats$Treatment, levels = c ("Water", "Glucose", "3OMG"))

#Number of samples
rep_counts <- Plant_data_cleaned_Stats %>%
  group_by(System, Microbes, Treatment) %>%
  summarise(n = n(), .groups = "drop")

# Reorder the 'System' and 'Microbes' factors
Plant_data_cleaned_Stats$Microbes <- factor(Plant_data_cleaned_Stats$Microbes, levels = c("MF" ,"F16", "Vd", "F96", "AMF"))  # desired order

rep_counts$Microbes <- factor(rep_counts$Microbes, levels = c("MF" ,"F16", "Vd", "F96", "AMF"))  # desired order

# Plot
Plot_PLant_T <- ggplot(Plant_data_cleaned_Stats, aes(x = Treatment, y = LogDeltaCT, fill = Treatment)) + 
  geom_boxplot(alpha = 1, outlier.shape = NA) +
  geom_jitter(alpha = 0.6, width = 0.2) +
  ylab(expression("Root Colonization " ~ 2^{Delta*CT} ~ italic(AtUBQ10))) +
  xlab("") +
  facet_grid(System~Microbes, space = "free", scales = "free", labeller = as_labeller(labels2)) +
    scale_fill_manual(values = color_palette) +
  labs(shape = "Biological Replicates") +
  base_theme +
   # Add significant p-values
  stat_pvalue_manual(
    ok <- Plant_data_cleaned_Stats %>%
  filter(Treatment != "3OMG") %>%                       # 🔹 remove 3OMG
  distinct(System, Microbes, Treatment, label_text,     # 🔹 avoid duplicates
           p_value, y_pos, .keep_all = TRUE),
    label = "label_text",
    y.position = 6,
    tip.length = 0.02,
    hide.ns = TRUE,
    size = 3
  )+
  # Add replicate counts
  geom_text(
    data = rep_counts,
    aes(x = Treatment, y = min(Plant_data_cleaned_Stats$LogDeltaCT, na.rm = TRUE) - 0.5, 
        label = paste0("n = ", n)),
    inherit.aes = FALSE,
    size = 3
  ) + theme(
    strip.background = element_rect(color = "black", fill = "white"))

Plot_PLant_T

#Save The table for HeatMap
Plant_data_cleaned_Stats_MT <- Plant_data_cleaned_Stats

#MAke plot but Plecrto alone
Plant_data_cleaned_Stats_F16 <- subset(Plant_data_cleaned_Stats, Microbes %in% c("F16"))
Plant_data_cleaned_Stats_F16 <- subset(Plant_data_cleaned_Stats, Microbes %in% c("F16"))
rep_counts_F16 <- subset(rep_counts, Microbes %in% c("F16"))

#Plot
Plot_PLant_F16 <- ggplot(Plant_data_cleaned_Stats_F16, 
                       aes(x = Treatment, y = LogDeltaCT, fill = Treatment)) + 
  geom_boxplot(alpha = 1, outlier.shape = NA) +
  geom_jitter(alpha = 0.6, width = 0.2) +
  ylab(expression("Root Colonization " ~ 2^{Delta*CT} ~ italic(AtUBQ10))) +
  xlab("") +
  facet_grid(System ~ Microbes, 
             space = "free", 
             scales = "free_y",
             labeller = as_labeller(labels2)) +
  scale_fill_manual(values = color_palette) +
  labs(shape = "Biological Replicates") +
  base_theme +
  stat_pvalue_manual(
    ok <- Plant_data_cleaned_Stats_F16 %>%
      filter(Treatment != "3OMG") %>%
      distinct(System, Microbes, Treatment, label_text,
               p_value, y_pos, .keep_all = TRUE),
    label = "label_text",
    y.position = 6,
    tip.length = 0.02,
    hide.ns = TRUE,
    size = 3
  ) +
  geom_text(
    data = rep_counts_F16,
    aes(x = Treatment, 
        y = min(Plant_data_cleaned_Stats_F16$LogDeltaCT, na.rm = TRUE) - 0.5,
        label = paste0("n = ", n)),
    inherit.aes = FALSE,
    size = 3
  ) +
  theme(
    strip.background = element_rect(color = "black", fill = "white")
  ) 

Plot_PLant_F16
```

## 5.  Barley

### a) - Colonization Matrixes
```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}
#Subset At
Matrix_data <- subset(Matrix_data2, Plant %in% c("Hordeum_vulgare"))

# Calculate adjusted CT
Matrix_data$CT[is.na(Matrix_data$CT)]<-40

# Remove Glucose
Matrix_data <- subset(Matrix_data, !Treatment %in% c("Glucose"))

# Clean soil data: remove invalid combinations
Matrix_data_cleaned <- Matrix_data %>%
  filter(
    (Primers == "Primer_16S" & Microbes == c("BSC", "BFSC")) |
    (Primers == "RS" & Microbes == c("Rs")) |
    (Primers == "ITS1" & Microbes == c("Pu")) |
    (Primers == "ITS" & !(Microbes %in% c("BSC", "Rs", "Pu", "MF")))
  )

# Create concat variable
Matrix_data_cleaned$concat <- paste0(Matrix_data_cleaned$Microbes, "/", Matrix_data_cleaned$Primers,"/",  Matrix_data_cleaned$Plant, "/", Matrix_data_cleaned$System)

# Stats: Wilcoxon test by group

get_signif_results <- function(data, x_start = 1, x_end = 2) {
  
  # Optional: see counts
  data %>%
    count(concat, Treatment) %>%
    print()
  
  # Nest data by concat
  data %>%
    group_by(concat) %>%
    nest() %>%
    mutate(
      test_result = map(data, ~ tryCatch(
        t.test(CT ~ Treatment, data = .x),
        error = function(e) NA_real_
      )),
      p_value = map_dbl(test_result, ~ if(is.list(.x)) .x$p.value else NA_real_),
      y_pos = map_dbl(data, ~ max(.x$CT, na.rm = TRUE) + 0.5),
      x_start = x_start,
      x_end = x_end,
      significance = case_when(
        is.na(p_value) ~ "ns",
        p_value < 0.001 ~ "***",
        p_value < 0.01 ~ "**",
        p_value < 0.05 ~ "*",
        TRUE ~ "ns"
      )) %>%
    select(concat, p_value, y_pos, x_start, x_end, significance) %>%
    unnest(cols = c())
}

#Run Stats
Stats_Matrix <- get_signif_results(Matrix_data_cleaned)

#Add Group Treatment
Stats_Matrix$group1 <- c("Water")
Stats_Matrix$group2 <- c("3OMG")

#Remake columns 
Stats_Matrix <- Stats_Matrix %>%
  separate(
    concat,
    into = c("Microbes", "Primers", "Plant", "System"),
    sep = "/",
    remove = FALSE  # keeps the original concat column
  )

# Merge stats with soil data
Matrix_data_cleaned_Stats <- Matrix_data_cleaned %>%
  left_join(Stats_Matrix, by = c("Microbes", "Primers", "Plant", "System"))

# Prepare the stats table for plotting
Matrix_data_cleaned_Stats <- Matrix_data_cleaned_Stats %>%
  mutate(
    # Only show significant p-values (< 0.05), rounded to 3 decimals
    label_text = ifelse(
      p_value < 0.05,
      paste0("P = ", round(p_value,5)),
      NA_character_
    )
  )

#Remove MiniSynCom
Matrix_data_cleaned_Stats <- subset(Matrix_data_cleaned_Stats, !Experiment %in% c("MiniSC"))

#Order
Matrix_data_cleaned_Stats$Treatment <- factor(Matrix_data_cleaned_Stats$Treatment, levels = c ("Water", "Glucose", "3OMG"))

#Number of samples
rep_counts <- Matrix_data_cleaned_Stats %>%
  group_by(System, Microbes, Treatment) %>%
  summarise(n = n(), .groups = "drop")

# Reorder the 'System' and 'Microbes' factors
Matrix_data_cleaned_Stats$Microbes <- factor(Matrix_data_cleaned_Stats$Microbes, levels = c("MF" ,"F16", "AMF", "Bs"))  # desired order

rep_counts$Microbes <- factor(rep_counts$Microbes, levels = c("MF" ,"F16", "AMF", "Bs"))  # desired order

# Plot
Plot_Matrix_B <- ggplot(Matrix_data_cleaned_Stats, aes(x = Treatment, y = -CT, fill = Treatment)) + 
  geom_boxplot(alpha = 1, outlier.shape = NA) +
  geom_jitter(alpha = 0.6, width = 0.2) +
  ylab(expression("Matrix Colonization (-Ct)")) +
  xlab("") +
  facet_grid(System ~Microbes , space = "free", scales = "free", labeller = as_labeller(labels2)) +
    scale_fill_manual(values = color_palette) +
  labs(shape = "Biological Replicates") +
  base_theme +
  # Add significant p-values
  stat_pvalue_manual(
    ok <- Matrix_data_cleaned_Stats %>%
  filter(Treatment != "3OMG") %>%                       # 🔹 remove 3OMG
  distinct(System, Microbes, Treatment, label_text,     # 🔹 avoid duplicates
           p_value, y_pos, .keep_all = TRUE),
    label = "label_text",
    y.position = -10,
    tip.length = 0.02,
    hide.ns = TRUE,
    size = 3
  )+
  # Add replicate counts
  geom_text(
    data = rep_counts,
    aes(x = Treatment, y = -40, 
        label = paste0("n = ", n)),
    inherit.aes = FALSE,
    size = 3
  ) + theme(
    strip.background = element_rect(color = "black", fill = "white"))
Plot_Matrix_B

#Save The table for HeatMap
Matrix_data_cleaned_Stats_B <- Matrix_data_cleaned_Stats
```

### b) - Colonization Plant

```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}
#Subset At
Plant_data <- subset(Plant_data2, Plant %in% c("Hordeum_vulgare"))

#Put non detected Microbial primers as a value of 40
Plant_data$UBQ10[is.na(Plant_data$UBQ10)] <- 40
Plant_data$CT[is.na(Plant_data$CT)] <- 40

#Remove when plants are not detected
Plant_data <- subset(Plant_data, !UBQ10 %in% c(40))

# Remove Glucose
Plant_data <- subset(Plant_data, !Treatment %in% c("Glucose"))

# Clean soil data: remove invalid combinations
Plant_data_cleaned <- Plant_data %>%
  filter(
    (Primers == "Primer_16S" & Microbes == c("BSC", "BFSC")) |
    (Primers == "RS" & Microbes == c("Rs")) |
    (Primers == "ITS1" & Microbes == c("Pu")) |
    (Primers == "ITS" & !(Microbes %in% c("BSC", "Rs", "Pu", "MF")))
  )

# Calculate the ratio for each row (2^Dekta CT)
Plant_data_cleaned$LogDeltaCT <- log(2^(-(Plant_data_cleaned$CT - Plant_data_cleaned$UBQ10)))
Plant_data_cleaned$DeltaCT <- 2^(-(Plant_data_cleaned$CT - Plant_data_cleaned$UBQ10))

# Create concat variable
Plant_data_cleaned$concat <- paste0(Plant_data_cleaned$Microbes, "/", Plant_data_cleaned$Primers,"/",  Plant_data_cleaned$Plant, "/", Plant_data_cleaned$System)

# Stats: Wilcoxon test by group
get_signif_results <- function(data, x_start = 1, x_end = 2) {
  
  # Optional: see counts
  data %>%
    count(concat, Treatment) %>%
    print()
  
  # Nest data by concat
  data %>%
    group_by(concat) %>%
    nest() %>%
    mutate(
      test_result = map(data, ~ tryCatch(
        t.test(LogDeltaCT ~ Treatment, data = .x),
        error = function(e) NA_real_
      )),
      p_value = map_dbl(test_result, ~ if(is.list(.x)) .x$p.value else NA_real_),
      y_pos = map_dbl(data, ~ max(.x$DeltaCT, na.rm = TRUE) + 0.5),
      x_start = x_start,
      x_end = x_end,
      significance = case_when(
        is.na(p_value) ~ "ns",
        p_value < 0.001 ~ "***",
        p_value < 0.01 ~ "**",
        p_value < 0.05 ~ "*",
        TRUE ~ "ns"
      )) %>%
    select(concat, p_value, y_pos, x_start, x_end, significance) %>%
    unnest(cols = c())
}

#Run Stats
Stats_Plant <- get_signif_results(Plant_data_cleaned)

#Add Group Treatment
Stats_Plant$group1 <- c("Water")
Stats_Plant$group2 <- c("3OMG")

#Remake columns 
Stats_Plant <- Stats_Plant %>%
  separate(
    concat,
    into = c("Microbes", "Primers", "Plant", "System"),
    sep = "/",
    remove = FALSE  # keeps the original concat column
  )

# Merge stats with soil data
Plant_data_cleaned_Stats <- Plant_data_cleaned %>%
  left_join(Stats_Plant, by = c("Microbes", "Primers", "Plant", "System"))

# Prepare the stats table for plotting
Plant_data_cleaned_Stats <- Plant_data_cleaned_Stats %>%
  mutate(
    # Only show significant p-values (< 0.05), rounded to 3 decimals
    label_text = ifelse(
      p_value < 0.05,
      paste0("P = ", round(p_value,5)),
      NA_character_
    )
  )

#Remove MiniSynCom
Plant_data_cleaned_Stats <- subset(Plant_data_cleaned_Stats, !Experiment %in% c("MiniSC"))

#Order
Plant_data_cleaned_Stats$Treatment <- factor(Plant_data_cleaned_Stats$Treatment, levels = c ("Water", "Glucose", "3OMG"))

#Number of samples
rep_counts <- Plant_data_cleaned_Stats %>%
  group_by(System, Microbes, Treatment) %>%
  summarise(n = n(), .groups = "drop")

# Reorder the 'System' and 'Microbes' factors
Plant_data_cleaned_Stats$Microbes <- factor(Plant_data_cleaned_Stats$Microbes, levels = c("MF" ,"F16", "AMF", "Bs"))  # desired order

rep_counts$Microbes <- factor(rep_counts$Microbes, levels = c("MF" ,"F16", "AMF", "Bs"))  # desired order

# Plot
Plot_PLant_B <- ggplot(Plant_data_cleaned_Stats, aes(x = Treatment, y = LogDeltaCT, fill = Treatment)) + 
  geom_boxplot(alpha = 1, outlier.shape = NA) +
  geom_jitter(alpha = 0.6, width = 0.2) +
  ylab(expression("Root Colonization " ~ 2^{Delta*CT} ~ italic(AtUBQ10))) +
  xlab("") +
  facet_grid(System~Microbes, space = "free", scales = "free", labeller = as_labeller(labels2)) +
    scale_fill_manual(values = color_palette) +
  labs(shape = "Biological Replicates") +
  base_theme +
  # Add significant p-values
  stat_pvalue_manual(
    ok <- Plant_data_cleaned_Stats %>%
  filter(Treatment != "3OMG") %>%                       # 🔹 remove 3OMG
  distinct(System, Microbes, Treatment, label_text,     # 🔹 avoid duplicates
           p_value, y_pos, .keep_all = TRUE),
    label = "label_text",
    y.position = 12,
    tip.length = 0.02,
    hide.ns = TRUE,
    size = 3
  )+
  # Add replicate counts
  geom_text(
    data = rep_counts,
    aes(x = Treatment, y = min(Plant_data_cleaned_Stats$LogDeltaCT, na.rm = TRUE) - 0.5, 
        label = paste0("n = ", n)),
    inherit.aes = FALSE,
    size = 3
  ) + theme(
    strip.background = element_rect(color = "black", fill = "white"))
Plot_PLant_B

#Save The table for HeatMap
Plant_data_cleaned_Stats_B <- Plant_data_cleaned_Stats

#Get median
median <- Plant_data_cleaned_Stats %>%
  group_by(System, Microbes, Treatment) %>%
  summarise(
    median_area = median(DeltaCT, na.rm = TRUE),
    .groups = "drop"
  )
median
```


## 4 - Final Plot
```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}
# Stack plots vertically with matching widths
final_plot_At <- plot_grid(
  Plot_PLant_at,
  Plot_Matrix_at,
  ncol = 1,
   align = "hv",                 # align horizontally
  axis = "tb",                 # align top and bottom axes
 scale = c(1, 1)   # scale individual plots
)
final_plot_At

# Stack plots vertically with matching widths
final_plot_At <- plot_grid(
  Plot_PLant_at,
  Plot_Matrix_at,
  ncol = 1,
   align = "hv",                 # align horizontally
  axis = "tb",                 # align top and bottom axes
 scale = c(1, 1)   # scale individual plots
)
final_plot_At

# Stack plots vertically with matching widths
final_plot_MT_B <- plot_grid(
  Plot_PLant_T,
  Plot_PLant_B,
  Plot_Matrix_T,
  Plot_Matrix_B,
  ncol = 2,
   align = "hv",                 # align horizontally
  axis = "tb",                 # align top and bottom axes
 scale = c(1, 1,1,1)   # scale individual plots
)
final_plot_MT_B
```
# B] HeatMap Colonization

## 1 - Combine tables
```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}
#Add missing columns
Matrix_data_cleaned_Stats_At$DeltaCT <- NA
Matrix_data_cleaned_Stats_MT$DeltaCT <- NA
Matrix_data_cleaned_Stats_B$DeltaCT <- NA
Matrix_data_cleaned_Stats_At$LogDeltaCT <- NA
Matrix_data_cleaned_Stats_MT$LogDeltaCT <- NA
Matrix_data_cleaned_Stats_B$LogDeltaCT <- NA

Full_table_bind <- rbind(Matrix_data_cleaned_Stats_At, Plant_data_cleaned_Stats_At, Matrix_data_cleaned_Stats_MT, Plant_data_cleaned_Stats_MT, Matrix_data_cleaned_Stats_B, Plant_data_cleaned_Stats_B)
```

## 1 - Arabidopsis

### a) - Colonization Matrixes

```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}

#Data for matrix
Full_table_bind

#Data for matrix
Matrix_data_cleaned_Stats_At <- subset(Full_table_bind, Plant %in% c("Arabidopsis_thaliana"))
Matrix_data_cleaned_Stats_At <- subset(Matrix_data_cleaned_Stats_At, !Microbes %in% c("Vd", "Bs", "AMF"))

#Add column Microba + primer
Matrix_data_cleaned_Stats_At$Microbes_Primers <- paste0(Matrix_data_cleaned_Stats_At$Microbes, "_", Matrix_data_cleaned_Stats_At$Primers)

# Add a column indicating significance
Matrix_data_cleaned_Stats_At <- Matrix_data_cleaned_Stats_At %>%
  mutate(Significant = ifelse(p_value < 0.05, TRUE, FALSE))

#Subset
Matrix_data_cleaned_Stats_At_Agar <- subset(Matrix_data_cleaned_Stats_At, System %in% c("MS_Agar"))
Matrix_data_cleaned_Stats_At_Plant <- subset(Matrix_data_cleaned_Stats_At, !System %in% c("MS_Agar"))

# Reorder the 'System' and 'Microbes' factors
Matrix_data_cleaned_Stats_At_Agar$Microbes <- factor(Matrix_data_cleaned_Stats_At_Agar$Microbes, levels = c( "MF" ,"F16", "F117", "F96", "F147", "F212", "F44", "F243", "F100", "F64", "Pu", "Rs"))  # desired order

# Heatmap with conditional dark outline
plot_Matrix_Agar_at <- ggplot(Matrix_data_cleaned_Stats_At_Agar, 
               aes(x = Treatment, y = System, fill = -CT)) +
  geom_tile(aes(color = ifelse(Significant, "black", "white")), size = 0.8) +
  scale_color_identity() +  # use the colors as-is
  scale_fill_gradientn(
  colors = c("#ffffff", "#fee0d2", "#fc9272", "#de2d26", "darkred"),
  values = c(0, 0.3, 0.6, 0.8, 1),
    na.value = "grey90",
    name = "Colonization\n(adjCT)"
  ) +
  facet_grid(~Microbes, scales = "free_y", labeller = as_labeller(labels2)) +
  labs(
    x = "Treatment",
    y = "System",
    title = "Plant Colonization Matrix Agar"
  ) +
  base_theme + theme(
    strip.background = element_rect(color = "black", fill = "white"))

plot_Matrix_Agar_at

# Reorder the 'System' and 'Microbes' factors
Matrix_data_cleaned_Stats_At_Plant$Microbes <- factor(Matrix_data_cleaned_Stats_At_Plant$Microbes, levels = c( "MF" ,"F16", "F117", "F96", "F147", "F212", "F44", "F243", "F100", "F64", "Pu", "Rs"))  # desired order

# Assume your data frame is called df
plot_Matrix_FP_At <- ggplot(Matrix_data_cleaned_Stats_At_Plant, 
               aes(x = Treatment, y = System, fill = -CT)) +
  geom_tile(aes(color = ifelse(Significant, "black", "white")), size = 0.8) +
  scale_color_identity() +  # use the colors as-is
  scale_fill_gradientn(
  colors = c("#ffffff", "#fee0d2", "#fc9272", "#de2d26", "darkred"),
  values = c(0, 0.3, 0.6, 0.8, 1),
    na.value = "grey90",
    name = "Colonization\n(adjCT)"
  ) +
  facet_grid(~ Microbes, scales = "free_y", labeller = as_labeller(labels2)) +
  labs(
    x = "Treatment",
    y = "Matrix",
    title = "Plant Colonization Matrix FP (Arabidopsis thaliana)"
  )+
  base_theme + theme(
    strip.background = element_rect(color = "black", fill = "white"))

plot_Matrix_FP_At
```

### b) - Colonization Plants

```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}
#Data for matrix
Full_table_bind

#Data for matrix
Plant_data_cleaned_Stats_At <- subset(Full_table_bind, Plant %in% c("Arabidopsis_thaliana"))


#Add column Microba + primer
Plant_data_cleaned_Stats_At$Microbes_Primers <- paste0(Plant_data_cleaned_Stats_At$Microbes, "_", Plant_data_cleaned_Stats_At$Primers)

# Add a column indicating significance
Plant_data_cleaned_Stats_At <- Plant_data_cleaned_Stats_At %>%
  mutate(Significant = ifelse(p_value < 0.05, TRUE, FALSE))

#Subset
Plant_data_cleaned_Stats_At_Agar <- subset(Plant_data_cleaned_Stats_At, System %in% c("MS_Agar"))

Plant_data_cleaned_Stats_At_Plant <- subset(Plant_data_cleaned_Stats_At, !System %in% c("MS_Agar"))

# Reorder the 'System' and 'Microbes' factors
Plant_data_cleaned_Stats_At_Agar$Microbes <- factor(Plant_data_cleaned_Stats_At_Agar$Microbes, levels = c( "MF" ,"F16", "F117", "F96", "F147", "F212", "F44", "F243", "F100", "F64", "Pu", "Rs"))  # desired order


# Heatmap with conditional dark outline
plot_Plant_Agar_At <- ggplot(Plant_data_cleaned_Stats_At_Agar, 
               aes(x = Treatment, y = System, fill = LogDeltaCT)) +
  geom_tile(aes(color = ifelse(Significant, "black", "white")), size = 0.8) +
  scale_color_identity() +  # use the colors as-is
   scale_fill_gradientn(
  colors = c("#ffffff", "#fee0d2", "#fc9272", "#de2d26", "darkred"),
  values = c(0, 0.3, 0.6, 0.8, 1),
  name = "Colonization\n(adjCT)",
  na.value = "grey90"
) +
  facet_grid(~ Microbes, scales = "free_y", labeller = as_labeller(labels2)) +
  labs(
    x = "Treatment",
    y = "System",
    title = "Plant Colonization Agar (Arabidopsis thaliana)"
  )+
  base_theme + theme(
    strip.background = element_rect(color = "black", fill = "white"))

plot_Plant_Agar_At

# Reorder the 'System' and 'Microbes' factors
Plant_data_cleaned_Stats_At_Plant$Microbes <- factor(Plant_data_cleaned_Stats_At_Plant$Microbes, levels = c( "MF" ,"F16", "F117", "F96", "F147", "F212", "F44", "F243", "F100", "F64", "Pu", "Rs"))  # desired order



# Assume your data frame is called df
plot_Plant_FP_At <- ggplot(Plant_data_cleaned_Stats_At_Plant, 
               aes(x = Treatment, y = System, fill = LogDeltaCT)) +
  geom_tile(aes(color = ifelse(Significant, "black", "white")), size = 0.8) +
  scale_color_identity() +  # use the colors as-is
  scale_fill_gradientn(
  colors = c("#ffffff", "#fee0d2", "#fc9272", "#de2d26", "darkred"),
  values = c(0, 0.3, 0.6, 0.8, 1),
  name = "Colonization\n(adjCT)",
  na.value = "grey90"
)+
  facet_grid(~ Microbes, scales = "free_y", labeller = as_labeller(labels2)) +
  labs(
    x = "Treatment",
    y = "System",
    title = "Plant Colonization Flowpots (Arabidopsis thaliana)"
  )+
  base_theme + theme(
    strip.background = element_rect(color = "black", fill = "white"))

plot_Plant_FP_At
```

## 2 - Tomato

### a) - Colonization Matrixes

```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}

#Data for matrix
Full_table_bind

#Data for matrix
Matrix_data_cleaned_Stats_MT <- subset(Full_table_bind, Plant %in% c("Micro_Tomato"))

#Add column Microba + primer
Matrix_data_cleaned_Stats_MT$Microbes_Primers <- paste0(Matrix_data_cleaned_Stats_MT$Microbes, "_", Matrix_data_cleaned_Stats_MT$Primers)

# Add a column indicating significance
Matrix_data_cleaned_Stats_MT <- Matrix_data_cleaned_Stats_MT %>%
  mutate(Significant = ifelse(p_value < 0.05, TRUE, FALSE))

#Subset
Matrix_data_cleaned_Stats_MT_Plant <- subset(Matrix_data_cleaned_Stats_MT, !System %in% c("MS_Agar"))

# Reorder the 'System' and 'Microbes' factors
Matrix_data_cleaned_Stats_MT_Plant$Microbes <- factor(Matrix_data_cleaned_Stats_MT_Plant$Microbes, levels = c("MF" ,"F16", "Vd", "F96", "AMF"))  # desired order

# Assume your data frame is called df
plot_Matrix_FP_MT <- ggplot(Matrix_data_cleaned_Stats_MT_Plant, 
               aes(x = Treatment, y = System, fill = -CT)) +
  geom_tile(aes(color = ifelse(Significant, "black", "white")), size = 0.8) +
  scale_color_identity() +  # use the colors as-is
  scale_fill_gradientn(
  colors = c("#7FC29B", "#dff5eb", "#f7f7f7", "#000000"),
    values = c(0, 0.5,  0.8, 1),
    na.value = "grey90",
    name = "Colonization\n(adjCT)"
  ) +
  facet_grid(~ Microbes, scales = "free_y", labeller = as_labeller(labels2)) +
  labs(
    x = "Treatment",
    y = "System",
    title = "Plant Colonization Matrix FP (Micro tomato)"
  )+
  base_theme + theme(
    strip.background = element_rect(color = "black", fill = "white"))

plot_Matrix_FP_MT
```

### b) - Colonization Plants

```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}
#Data for matrix
Full_table_bind

#Data for matrix
Plant_data_cleaned_Stats_MT <- subset(Full_table_bind, Plant %in% c("Micro_Tomato"))


#Add column Microba + primer
Plant_data_cleaned_Stats_MT$Microbes_Primers <- paste0(Plant_data_cleaned_Stats_MT$Microbes, "_", Plant_data_cleaned_Stats_MT$Primers)

# Add a column indicating significance
Plant_data_cleaned_Stats_MT <- Plant_data_cleaned_Stats_MT %>%
  mutate(Significant = ifelse(p_value < 0.05, TRUE, FALSE))

#Subset
Plant_data_cleaned_Stats_MT_Plant <- subset(Plant_data_cleaned_Stats_MT, !System %in% c("MS_Agar"))

# Plot
plot_Plant_FP_MT <- ggplot(Plant_data_cleaned_Stats_MT_Plant, 
               aes(x = Treatment, y = System, fill = LogDeltaCT)) +
  geom_tile(aes(color = ifelse(Significant, "black", "white")), size = 0.8) +
  scale_color_identity() +  # use the colors as-is
  scale_fill_gradientn(
    colors = c("#7FC29B", "#dff5eb", "#f7f7f7", "#bfbfbf", "#7f7f7f", "#000000"),
    values = c(0, 0.2, 0,4, 0.6,  0.8, 1),
    na.value = "grey90",
    name = "Colonization\n(adjCT)"
  ) +
  facet_grid(~ Microbes, scales = "free_y", labeller = as_labeller(labels2)) +
  labs(
    x = "Treatment",
    y = "System",
    title = "Plant Colonization Flowpots (Micro tomato)"
  )+
  base_theme + theme(
    strip.background = element_rect(color = "black", fill = "white"))

plot_Plant_FP_MT
  
```

## 3 - Barley

### a) - Colonization Matrixes

```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}

#Data for matrix
Full_table_bind

#Data for matrix
Matrix_data_cleaned_Stats_HV <- subset(Full_table_bind, Plant %in% c("Hordeum_vulgare"))

#Add column Microba + primer
Matrix_data_cleaned_Stats_HV$Microbes_Primers <- paste0(Matrix_data_cleaned_Stats_HV$Microbes, "_", Matrix_data_cleaned_Stats_HV$Primers)

# Add a column indicating significance
Matrix_data_cleaned_Stats_HV <- Matrix_data_cleaned_Stats_HV %>%
  mutate(Significant = ifelse(p_value < 0.05, TRUE, FALSE))

#Subset
Matrix_data_cleaned_Stats_HV_Plant <- subset(Matrix_data_cleaned_Stats_HV, !System %in% c("MS_Agar"))

# Reorder the 'System' and 'Microbes' factors
Matrix_data_cleaned_Stats_HV_Plant$Microbes <- factor(Matrix_data_cleaned_Stats_HV_Plant$Microbes, levels = c("MF" ,"F16", "AMF", "Bs")) # desired order

# Assume your data frame is called df
plot_Matrix_FP_HV <- ggplot(Matrix_data_cleaned_Stats_HV_Plant, 
               aes(x = Treatment, y = System, fill = -CT)) +
  geom_tile(aes(color = ifelse(Significant, "black", "white")), size = 0.8) +
  scale_color_identity() +  # use the colors as-is
  scale_fill_gradientn(
  colors = c("#7FC29B", "#dff5eb", "#f7f7f7", "#000000"),
    values = c(0, 0.5,  0.8, 1),
    na.value = "grey90",
    name = "Colonization\n(adjCT)"
  ) +
  facet_grid(~Microbes, scales = "free_y", labeller = as_labeller(labels2)) +
  labs(
    x = "Treatment",
    y = "System",
    title = "Plant Colonization Matrix FP (Micro tomato)"
  )+
  base_theme + theme(
    strip.background = element_rect(color = "black", fill = "white"))

plot_Matrix_FP_HV
```

### b) - Colonization Plants

```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}
#Data for matrix
Full_table_bind

#Data for matrix
Plant_data_cleaned_Stats_HV <- subset(Full_table_bind, Plant %in% c("Hordeum_vulgare"))


#Add column Microba + primer
Plant_data_cleaned_Stats_HV$Microbes_Primers <- paste0(Plant_data_cleaned_Stats_HV$Microbes, "_", Plant_data_cleaned_Stats_HV$Primers)

# Add a column indicating significance
Plant_data_cleaned_Stats_HV <- Plant_data_cleaned_Stats_HV %>%
  mutate(Significant = ifelse(p_value < 0.05, TRUE, FALSE))

#Subset
Plant_data_cleaned_Stats_HV_Plant <- subset(Plant_data_cleaned_Stats_HV, !System %in% c("MS_Agar"))

# Plot
plot_Plant_FP_HV <- ggplot(Plant_data_cleaned_Stats_HV_Plant, 
               aes(x = Treatment, y = System, fill = LogDeltaCT)) +
  geom_tile(aes(color = ifelse(Significant, "black", "white")), size = 0.8) +
  scale_color_identity() +  # use the colors as-is
  scale_fill_gradientn(
    colors = c("#7FC29B", "#dff5eb", "#f7f7f7", "#bfbfbf", "#7f7f7f", "#000000"),
    values = c(0, 0.2, 0,4, 0.6,  0.8, 1),
    na.value = "grey90",
    name = "Colonization\n(adjCT)"
  ) +
  facet_grid(~ Microbes, scales = "free_y", labeller = as_labeller(labels2)) +
  labs(
    x = "Treatment",
    y = "System",
    title = "Plant Colonization Flowpots (Micro tomato)"
  )+
  base_theme + theme(
    strip.background = element_rect(color = "black", fill = "white"))

plot_Plant_FP_HV
```

## 4 - All plots Heatmaps

```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}
#A. thaliana
# Plot Arabido Full
Plot_Arabido_Full <- plot_grid(
  plot_Plant_FP_At,
  plot_Matrix_FP_At,
    plot_Plant_Agar_At,
  plot_Matrix_Agar_at,
  ncol = 1,
   align = "hv",                 # align horizontally
  axis = "tb",                 # align top and bottom axes
    rel_widths = c(1,0.1))  
Plot_Arabido_Full


#M. tomato
plot_Matrix_FP_MT
plot_Plant_FP_MT

#Barley
plot_Matrix_FP_HV
plot_Plant_FP_HV
```

# C] Correlation Map

## 1 - Prepare data
```{r echo=FALSE, fig.show='hide', message=FALSE, results="hide"}
#Load data PPI
    raw_data_PPI<-read.csv(file = "Total_PlantPhenotype.csv", header = TRUE, sep=",")
    
    raw_data_PPI<-raw_data_PPI%>%
      dplyr::left_join(Kingdom,by = "Microbes")%>%
      filter(Treatment!="Glucose")
    
    raw_data_PPI$Weight_length[is.na(raw_data_PPI$Weight_length)]<-0
    raw_data_PPI$Weight_length_per_plant<-raw_data_PPI$Weight_length/raw_data_PPI$Nb
    
    #Remove NA
    raw_data_PPI$Weight_length_per_plant[is.na(raw_data_PPI$Weight_length_per_plant)] <- 0
    
    #Calculate GR
    raw_data_PPI$GR <- raw_data_PPI$Nb / max(raw_data_PPI$Nb, na.rm = TRUE)
        
    #Calculate PPI (Plant Performance Index)   
    raw_data_PPI$PPI <- ifelse(
  raw_data_PPI$System == "Flowpots",
  raw_data_PPI$GR * raw_data_PPI$Weight_length_per_plant,   # Flowpots
  raw_data_PPI$Weight_length_per_plant                       # Other systems
)

raw_data_PPI$PPI[is.na(raw_data_PPI$PPI)] <- 0
        

    #Calcualte the median PPI
      raw_data_PPI_summary <- raw_data_PPI %>%
        group_by(Plant, System, Experiment, Bio_rep, Microbes, Treatment, Element, Kingdom) %>%
        summarise(
          Median_PPI = median(PPI, na.rm = TRUE)
        ) %>%
        ungroup()
      
      #For merging need to change column System MS_agar into MS_Agar
raw_data_PPI_summary$System[raw_data_PPI_summary$System == "MS_agar"] <- "MS_Agar"


#Load data CT
    raw_data_Ct <- Full_table_bind
    
     #Calcualte the median LogDeltaCT
      raw_data_Ct_summary <- raw_data_Ct %>%
        group_by(Plant, System, Experiment, Bio_rep, Microbes, Treatment, Element) %>%
        summarise(
          Median_Ct = median(CT, na.rm = TRUE),
          Median_LogDeltaCT = median(LogDeltaCT, na.rm = TRUE)
        ) %>%
        ungroup()
      
    # Merge the two tables on conditions
Data_full <- merge(
  raw_data_PPI_summary[, c("Plant", "System", "Bio_rep", 
                           "Microbes", "Treatment","Kingdom", "Median_PPI")],
  raw_data_Ct_summary[, c("Plant", "System", "Bio_rep", 
                          "Microbes", "Treatment", "Element", "Median_Ct", "Median_LogDeltaCT")],
  by = c("Plant", "System", "Bio_rep", 
         "Microbes", "Treatment"),
  all.x = TRUE
)

#Remove MF
Data_full <- subset(Data_full, !Microbes %in% c("MF", "BSC", "FSC", "BFSC"))
```

## 2 - Arabidopsis 

### a) - Colonization Matrixes
```{r echo=FALSE, fig.show='hide', message=FALSE, results="hide"}
# Keep only Arabidopsis_thaliana, Matrix, Fungi
Data_full_At_Matrix <- subset(
  Data_full,
  Plant %in% c("Arabidopsis_thaliana") &
    Element %in% c("Matrix") &
    Kingdom %in% c("Fungi")
)

# Compute correlation stats per Plant *and* System
cor_stats <- Data_full_At_Matrix %>%
  group_by(Plant, System) %>%
  summarise(
    r = cor(-Median_Ct, Median_PPI, use = "complete.obs"),
    r2 = r^2,
    p_value = cor.test(-Median_Ct, Median_PPI)$p.value,
    .groups = "drop"
  )

# Compute x-position per facet for correlation labels
cor_stats_labels <- cor_stats %>%
  group_by(System, Plant) %>%
  summarise(
    y_pos = max(Data_full_At_Matrix$Median_PPI[Data_full_At_Matrix$System == System &
                                             Data_full_At_Matrix$Plant == Plant], na.rm = TRUE),
    x_pos = -max(Data_full_At_Matrix$Median_Ct[Data_full_At_Matrix$System == System &
                                                     Data_full_At_Matrix$Plant == Plant], na.rm = TRUE),
    label = paste0("\nR² = ", round(r2, 2), ", P = ", signif(p_value, 2)),
    .groups = "drop"
  )

Plot_corr_Matrix_at <- ggplot(Data_full_At_Matrix, aes(x = -Median_Ct, y = Median_PPI, color = Treatment)) +
  geom_point(size = 2, alpha = 0.8) +
  geom_smooth(method = "lm", se = FALSE, color = "black", linewidth = 0.5) +
 facet_wrap(~ System,
  scales = "free",
  labeller = labeller(
    System = c(
      "Flowpots" = "System : FlowPot",
      "MS_Agar"  = "System : Agar 1/2 MS"))) + 
  labs(
   x = expression(Matrix~Colonization~(-CT)),
    y = "Median PPI",
    title = "",
    shape = "System",
    linetype = "System",
    color = "Kingdom"
  ) +
  scale_color_manual(values = color_palette) +
  base_theme +
  theme(
    strip.text = element_text(size = 12),
    plot.title = element_text(hjust = 0.5),
    legend.position = "right"
  ) +
  geom_text(
    data = cor_stats_labels,
    aes(x = x_pos, y = y_pos, label = label),
    hjust = 0,
    vjust = 1,
    size = 2.5,
    color = "black",
    inherit.aes = FALSE,
    show.legend = FALSE
  )

Plot_corr_Matrix_at

#Get median
median <- Data_full_At_Matrix %>%
  group_by(System, Microbes, Treatment) %>%
  summarise(
    median_area = median(Median_PPI, na.rm = TRUE),
    .groups = "drop"
  )
median
```

### b) - Colonization Plants
```{r echo=FALSE, fig.show='hide', message=FALSE, results="hide"}
#Keep only At, Matrix
Data_full_At_Root <- subset(Data_full, Plant %in% c("Arabidopsis_thaliana") & !Element %in% c("Matrix") & Kingdom %in% c("Fungi"))

#test normality
normality <- Data_full_At_Root %>%
  select(System, Median_LogDeltaCT, Median_PPI) %>%
  gather(Variable, Value, -System) %>% 
  group_by(System, Variable) %>%
  shapiro_test(Value) 
normality # Good exept for Ct flowpots

#qqplot
qqplots <- qqplots <- Data_full_At_Root %>%
  group_by(System) %>%
  summarise(
    qq_FP = list(ggqqplot(Median_LogDeltaCT, title = paste0("QQ: ", System, " - FP"))),
    qq_CT = list(ggqqplot(Median_PPI,   title = paste0("QQ: ", System, " - CT")))
  ) #Really good

# Compute correlation stats per Plant *and* System
cor_stats <- Data_full_At_Root %>%
  group_by(Plant, System) %>%
  summarise(
    r = cor(-Median_LogDeltaCT, Median_PPI, use = "complete.obs"),
    r2 = r^2,
    p_value = cor.test(-Median_LogDeltaCT, Median_PPI)$p.value,
    .groups = "drop"
  )

# Compute x-position per facet for correlation labels
cor_stats_labels <- cor_stats %>%
  group_by(System, Plant) %>%
  summarise(
    y_pos = max(Data_full_At_Root$Median_PPI[Data_full_At_Root$System == System &
                                             Data_full_At_Root$Plant == Plant], na.rm = TRUE),
    x_pos = -max(Data_full_At_Root$Median_LogDeltaCT[Data_full_At_Root$System == System &
                                                     Data_full_At_Root$Plant == Plant], na.rm = TRUE),
    label = paste0("\nR² = ", round(r2, 2), ", P = ", signif(p_value, 2)),
    .groups = "drop"
  )

Plot_corr_PLant_at <- ggplot(Data_full_At_Root, 
                             aes(x = Median_LogDeltaCT, y = Median_PPI)) +
  geom_point(aes(fill = Treatment),
             shape = 21,
             color = "black",
             size = 2,
             alpha = 1) +
  geom_smooth(method = "lm", se = FALSE, color = "black", linewidth = 0.5) +
  facet_wrap(~ System,
             scales = "free",
             labeller = labeller(
               System = c(
                 Flowpots = "System : FlowPot",
                 MS_Agar  = "System : Agar 1/2 MS"
               )
             )) +
  labs(
    x = expression(Root~Colonization~(2^DeltaCT)),
    y = "Median PPI",
    color = "Kingdom"
  ) +
  scale_fill_manual(values = color_palette) +
  base_theme +
  theme(
    strip.text = element_text(size = 12),
    plot.title = element_text(hjust = 0.5),
    legend.position = "right"
  ) +
  base_theme +
  geom_text(
    data = cor_stats_labels,
    aes(x = 3, y = y_pos, label = label),
    hjust = 0,
    vjust = 1,
    size = 2.5,
    color = "black",
    inherit.aes = FALSE,
    show.legend = FALSE
  )

Plot_corr_PLant_at
```

## 3 - Correlation Systems 

### a) Plant Performance
```{r echo=FALSE, fig.show='hide', message=FALSE, results="hide"}
#Prepare tables
Data_full_At_Root_FP <- subset(Data_full_At_Root, System %in% c("Flowpots"))
Data_full_At_Root_Agar <- subset(Data_full_At_Root, System %in% c("MS_Agar"))

#Summarize bio_rep fro FP
Data_full_At_Root_FP_median <- Data_full_At_Root_FP %>%
  group_by(Plant, System, Microbes, Treatment, Kingdom) %>%
  summarise(
    median_PPI_FP = median(Median_PPI, na.rm = TRUE),
    Median_LogDeltaCT_FP  = median(Median_LogDeltaCT, na.rm = TRUE),
    .groups = "drop"
  )

#Merge tables
Merged_At_Data <- Data_full_At_Root_FP_median %>%
  left_join(
    Data_full_At_Root_Agar %>%
      select(Plant, System, Microbes, Treatment, Kingdom, Median_PPI, Median_LogDeltaCT),
    by = c("Plant", "Microbes", "Treatment", "Kingdom")
  )

#Add condition for facetting
Merged_At_Data$Data_type <- c("Plant phenotype")

#test normality
normality <- Merged_At_Data %>%
  select(median_PPI_FP, Median_PPI) %>%
  gather(Variable, Value) %>% 
  group_by(Variable) %>%
  shapiro_test(Value)
normality # Good only for protection Agar

#qqplot
qqplot <- ggqqplot(Merged_At_Data$median_PPI_FP, title = "QQ plot: Median_LogDeltaCT_FP") #Really good
qqplot2 <- ggqqplot(Merged_At_Data$Median_PPI,   title = "QQ plot: Median_LogDeltaCT") #Less good for first values

# Compute correlation stats per Plant *and* System
cor_stats_PPI <- Merged_At_Data %>%
  summarise(
    r = cor(median_PPI_FP, Median_PPI, use = "complete.obs"),
    r2 = r^2,
    p_value = cor.test(median_PPI_FP, Median_PPI)$p.value,
    .groups = "drop"
  )

cor_stats_PPI$Data_type <-  c("Root phenotype")

# Compute x-position per facet for correlation labels
cor_stats_labels <- cor_stats_PPI %>%
  summarise(
    y_pos = max(Merged_At_Data$Median_PPI),
    x_pos = min(Merged_At_Data$median_PPI_FP),
    label = paste0("\nR² = ", round(r2, 2), ", P = ", signif(p_value, 2)),
    .groups = "drop"
  )

#Plot correlation Systems
Plot_corr_PPI <- ggplot(Merged_At_Data, 
                        aes(x = median_PPI_FP, y = Median_PPI, color = Treatment)) +
  geom_point(size = 2, alpha = 1) +
  geom_smooth(method = "lm", se = FALSE, 
              color = "black", linewidth = 1) +
  facet_wrap(~ Data_type, scales = "free") +   # independent axes per panel
  labs(
    x = "Flowpot (Plant Performance Index)",
    y = "Agar 1/2 MS (Root length (mm))",
    title = "",
    shape = "System",
    linetype = "System",
    color = "Treatment"
  ) +
  scale_color_manual(values = color_palette) +
  base_theme +
  theme(
    strip.text = element_text(size = 12),
    plot.title = element_text(hjust = 0.5),
    legend.position = "right"
  ) +
  geom_text(
    data = cor_stats_labels,   # pre-calculated correlation stats
    aes(x = x_pos, y = y_pos, label = label),
    hjust = 0,
    vjust = 1,
    size = 2.5,
    color = "gray30",
    inherit.aes = FALSE,
    show.legend = FALSE
  ) 
Plot_corr_PPI
```

### b) Colonization

```{r echo=FALSE, fig.show='hide', message=FALSE, results="hide"}
#Merge tables
Merged_At_Data <- Data_full_At_Root_FP_median %>%
  left_join(
    Data_full_At_Root_Agar %>%
      select(Plant, System, Microbes, Treatment, Kingdom, Median_PPI, Median_LogDeltaCT),
    by = c("Plant", "Microbes", "Treatment", "Kingdom")
  )

#Add condition for facetting
Merged_At_Data$Data_type <- c("Plant colonization")

#test normality
normality <- Merged_At_Data %>%
  select(Median_LogDeltaCT_FP, Median_LogDeltaCT) %>%
  gather(Variable, Value) %>% 
  group_by(Variable) %>%
  shapiro_test(Value)
normality #Really good

#qqplot
qqplot <- ggqqplot(Merged_At_Data$Median_LogDeltaCT_FP, title = "QQ plot: Median_LogDeltaCT_FP") #Really good
qqplot2 <- ggqqplot(Merged_At_Data$Median_LogDeltaCT,   title = "QQ plot: Median_LogDeltaCT") #Really good

# Compute correlation stats per Plant *and* System
cor_stats_Ct <- Merged_At_Data %>%
  summarise(
    r = cor(Median_LogDeltaCT_FP, Median_LogDeltaCT, use = "complete.obs"),
    r2 = r^2,
    p_value = cor.test(Median_LogDeltaCT_FP, Median_LogDeltaCT)$p.value,
    .groups = "drop"
  )

cor_stats_Ct$Data_type <-  c("Root colonization")

# Compute x-position per facet for correlation labels
cor_stats_labels <- cor_stats_Ct %>%
  summarise(
    y_pos = max(Merged_At_Data$Median_LogDeltaCT),
    x_pos = min(Merged_At_Data$Median_LogDeltaCT_FP),
    label = paste0("\nR² = ", round(r2, 2), ", P = ", signif(p_value, 2)),
    .groups = "drop"
  )

#Plot correlation Systems
Plot_corr_Ct <- ggplot(Merged_At_Data, 
                        aes(x = Median_LogDeltaCT_FP, y = Median_LogDeltaCT, color = Treatment)) +
  geom_point(size = 2, alpha = 1) +
  geom_smooth(method = "lm", se = FALSE, 
              color = "black", linewidth = 1) +
  facet_wrap(~ Data_type, scales = "free") +   # independent axes per panel
  labs(
    x = expression(Flowpot~(2^{Delta*CT}~italic(AtUBQ10))),
    y = expression(Agar~(2^{Delta*CT}~italic(AtUBQ10))),
    title = "",
    shape = "System",
    linetype = "System",
    color = "Treatment"
  ) +
  scale_color_manual(values = color_palette) +
  base_theme +
  theme(
    strip.text = element_text(size = 12),
    plot.title = element_text(hjust = 0.5),
    legend.position = "right"
  ) +
  geom_text(
    data = cor_stats_labels,   # pre-calculated correlation stats
    aes(x = x_pos, y = y_pos, label = label),
    hjust = 0,
    vjust = 1,
    size = 2.5,
    color = "gray30",
    inherit.aes = FALSE,
    show.legend = FALSE
  ) 
Plot_corr_Ct

#write.csv(Merged_At_Data, "Colonization_Roots.csv")
```

## 3 - Tomato 

### a) - Colonization Matrixes
```{r echo=FALSE, fig.show='hide', message=FALSE, results="hide"}
# Keep only Arabidopsis_thaliana, Matrix, Fungi
Data_full_MT_Matrix <- subset(
  Data_full,
  Plant %in% c("Micro_Tomato") &
    Element %in% c("Matrix") &
    Kingdom %in% c("Fungi")
)

# Compute correlation stats per Plant *and* System
cor_stats <- Data_full_MT_Matrix %>%
  group_by(Plant, System) %>%
  summarise(
    r = cor(-Median_Ct, Median_PPI, use = "complete.obs"),
    r2 = r^2,
    p_value = cor.test(-Median_Ct, Median_PPI)$p.value,
    .groups = "drop"
  )

# Compute x-position per facet for correlation labels
cor_stats_labels <- cor_stats %>%
  group_by(System, Plant) %>%
  summarise(
    y_pos = max(Data_full_MT_Matrix$Median_PPI[Data_full_MT_Matrix$System == System &
                                             Data_full_MT_Matrix$Plant == Plant], na.rm = TRUE),
    x_pos = -max(Data_full_MT_Matrix$Median_Ct[Data_full_MT_Matrix$System == System &
                                                     Data_full_MT_Matrix$Plant == Plant], na.rm = TRUE),
    label = paste0("System: ", System, "\nR² = ", round(r2, 2), ", P = ", signif(p_value, 2)),
    .groups = "drop"
  )

Plot_corr_Matrix_MT <- ggplot(Data_full_MT_Matrix, aes(x = -Median_Ct, y = Median_PPI, shape = System, color = Kingdom)) +
  geom_point(size = 2, alpha = 0.8) +
  geom_smooth(method = "lm", se = FALSE, aes(linetype = System), color = "black", linewidth = 1) +
  facet_wrap(~System, scales = "free", labeller = labeller(System = c("Flowpots" = "System : Flowpots"))) +  # fully independent x & y axes per panel
  labs(
    x = "Median Ct",
    y = "Median PPI",
    title = "",
    shape = "System",
    linetype = "System",
    color = "Kingdom"
  ) +
  scale_color_manual(values = c("Fungi" = "#F9B244")) +
  base_theme +
  theme(
    strip.text = element_text(size = 12),
    plot.title = element_text(hjust = 0.5),
    legend.position = "right"
  ) +
  geom_text(
    data = cor_stats_labels,
    aes(x = x_pos, y = y_pos, label = label),
    hjust = 0,
    vjust = 1,
    size = 2.5,
    color = "gray30",
    inherit.aes = FALSE,
    show.legend = FALSE
  )

Plot_corr_Matrix_MT
```

### b) - Colonization Plants
```{r echo=FALSE, fig.show='hide', message=FALSE, results="hide"}
#Keep only At, Matrix
Data_full_MT_Root <- subset(Data_full, Plant %in% c("Micro_Tomato") & !Element %in% c("Matrix") & Kingdom %in% c("Fungi"))

# Compute correlation stats per Plant *and* System
cor_stats <- Data_full_MT_Root %>%
  group_by(Plant, System) %>%
  summarise(
    r = cor(Median_LogDeltaCT, Median_PPI, use = "complete.obs"),
    r2 = r^2,
    p_value = cor.test(Median_LogDeltaCT, Median_PPI)$p.value,
    .groups = "drop"
  )

# Compute x-position per facet for correlation labels
cor_stats_labels <- cor_stats %>%
  group_by(System, Plant) %>%
  summarise(
    y_pos = max(Data_full_MT_Root$Median_PPI[Data_full_MT_Root$System == System &
                                             Data_full_MT_Root$Plant == Plant], na.rm = TRUE),
    x_pos = min(Data_full_MT_Root$Median_LogDeltaCT[Data_full_MT_Root$System == System &
                                                     Data_full_MT_Root$Plant == Plant], na.rm = TRUE),
    label = paste0("System: ", System, "\nR² = ", round(r2, 2), ", P = ", signif(p_value, 2)),
    .groups = "drop"
  )

Plot_corr_Plant_MT <- ggplot(Data_full_MT_Root, aes(x = Median_LogDeltaCT, y = Median_PPI, shape = System, color = Kingdom)) +
  geom_point(size = 2, alpha = 0.8) +
  geom_smooth(method = "lm", se = FALSE, aes(linetype = System), color = "black", linewidth = 1) +
  facet_wrap(~ System, scales = "free", labeller = labeller(System = c("Flowpots" = "System : Flowpots"))) +  # fully independent x & y axes per panel
  labs(
    x = expression(Root~Colonization~(2^{Delta*CT}~italic(AtUBQ10))),
    y = "Median PPI",
    title = "",
    shape = "System",
    linetype = "System",
    color = "Kingdom"
  ) +
  scale_color_manual(values = c("Fungi" = "#F9B244")) +
  base_theme +
  theme(
    strip.text = element_text(size = 12),
    plot.title = element_text(hjust = 0.5),
    legend.position = "right"
  ) +
  geom_text(
    data = cor_stats_labels,
    aes(x = x_pos, y = y_pos, label = label),
    hjust = 0,
    vjust = 1,
    size = 2.5,
    color = "gray30",
    inherit.aes = FALSE,
    show.legend = FALSE
  )

Plot_corr_Plant_MT
```

## 4 - Final Plot Root
```{r, echo=FALSE, message=FALSE, results="hide", fig.show='hide'}
#Plot Colonization
Plot_PLant_at
Plot_PLant_T
Plot_PLant_B

#Plot Correlation
Plot_corr_Plant_at
Plot_corr_PPI
Plot_corr_Ct

final_plot_corr <- plot_grid(
  Plot_corr_Plant_at,
  Plot_corr_PPI,
  Plot_corr_Ct,
  ncol = 3,
  rel_heights = c(1),
  rel_widths = c(1,0.6, 0.6),
   align = "hv",
  axis = "tb")

final_plot_corr
```

---
END
---


